---
title: "Quantifying comprehensive marine vessel emissions using data fusion"
author: "emLab, UCSB and GFW"
date: last-modified
date-format: "[Last Updated on] MMMM DD, YYYY"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    fig-pos: H
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
cache: true
---

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE, cache = FALSE}
#| label: document-setup

# Set default chunk options
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  dev = "png",
  fig.path = "../figures/"
)

# Load librariers
library(ggplot2)
library(sf)

analysis_start_year <- 2016
analysis_end_year <- 2024
```

```{r}
#| label: setup-graphics
# Use Okabe-Ito colorblind-friendly colors
all_data_source_color_palette <- c(
  "Marine vessels (AIS + S1)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[5]],
  "AIS + S1" = paletteer::paletteer_d("colorblindr::OkabeIto")[[5]],
  "Terrestrial transportation (EDGAR)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[6]],
  "Terrestrial transportation  (IEA)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[1]],
  "AIS (broadcasting)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[3]],
  "S1 (non-broadcasting)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[
    7
  ]],
  "EDGAR" = paletteer::paletteer_d("colorblindr::OkabeIto")[[8]],
  "EIA" = paletteer::paletteer_d("colorblindr::OkabeIto")[[4]]
)

# Set consistent theme for all plots
theme_plot <- function() {
  theme_minimal() %+replace%
    theme(
      axis.title.x = element_text(face = "bold"),
      axis.title.y = element_text(angle = 90, face = "bold", vjust = 3),
      strip.text.x = element_text(angle = 0, face = "bold"),
      strip.text.y = element_text(angle = 0, face = "bold")
    )
}

# Set map projection for all map figures
map_projection <- "+proj=eqearth"

# Create bounding box of globe, for making black map backgrounds
lats <- c(90:-90, -90:90, 90)
longs <- c(rep(c(180, -180), each = 181), 180)
world_bbox_sf <-
  list(cbind(longs, lats)) %>%
  sf::st_polygon() %>%
  sf::st_sfc(
    # create sf geometry list column
    crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
  ) %>%
  sf::st_sf() %>%
  sf::st_transform(map_projection)

land <- rnaturalearth::ne_countries(returnclass = "sf") |>
  dplyr::select(geometry)
```

```{r}
#| label: load-data
targets::tar_load(c(
  "n_unique_vessels",
  "n_ais_messages",
  "annual_edgar_emissions",
  "annual_emissions_all_pollutants",
  "total_spatial_emissions_by_pollutant_2024",
  "annual_ais_to_dark_activity_extrapolation",
  "total_monthly_emissions_by_pollutant",
  "annual_ais_co2_emissions_by_vessel_type"
))
```

```{r}
#| label: wrangle-edgar-data
annual_edgar_tranportation_emissions <- annual_edgar_emissions |>
  dplyr::select(
    -c(
      IPCC_annex,
      C_group_IM24_sh,
      Country_code_A3,
      Name,
      ipcc_code_2006_for_standard_report,
      Substance,
      fossil_bio
    )
  ) |>
  dplyr::rename(sector = ipcc_code_2006_for_standard_report_name) |>
  tidyr::pivot_longer(
    cols = -c(sector),
    names_to = "year",
    values_to = "emissions_co2_kt"
  ) |>
  dplyr::mutate(
    year = stringr::str_remove_all(year, "Y_") |>
      as.numeric(),
    emissions_co2_mt = emissions_co2_kt * 1e3
  ) |>
  dplyr::select(-emissions_co2_kt) |>
  dplyr::mutate(
    land_based_tranpsportation = ifelse(
      sector %in%
        c(
          "Civil Aviation",
          "Road Transportation no resuspension",
          "Other Transportation",
          "Railways"
        ),
      TRUE,
      FALSE
    )
  ) |>
  # Only need years that correspond to our dataset
  dplyr::filter(year >= analysis_start_year, year <= analysis_end_year)

# Fraction of global 2023 CO2 emissions that came from transportation sectors
fraction_co2_from_transportation <- annual_edgar_tranportation_emissions |>
  dplyr::mutate(
    transportation = ifelse(
      land_based_tranpsportation | sector == "Water-borne Navigation",
      TRUE,
      FALSE
    )
  ) |>
  dplyr::filter(year == 2023) |>
  dplyr::group_by(transportation) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(
    names_from = "transportation",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::mutate(fraction_from_transportation = `TRUE` / (`TRUE` + `FALSE`)) |>
  dplyr::pull(fraction_from_transportation) |>
  scales::percent()
```

```{r}
#| label: wrangle-gfw-time-series-emissions-data

annual_emissions_all_pollutants <- annual_emissions_all_pollutants |>
  tidyr::pivot_longer(
    -c(year, fishing),
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    fleet = ifelse(
      stringr::str_detect(pollutant, "dark"),
      "Non-broadcasting",
      "AIS-broadcasting"
    )
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_mt") |>
      stringr::str_remove_all("_dark")
  )

annual_emissions_all_pollutants_with_total <- annual_emissions_all_pollutants |>
  # Add total, that aggregates across broadcasting and non-broadcasting
  dplyr::bind_rows(
    annual_emissions_all_pollutants |>
      dplyr::group_by(year, fishing, pollutant) |>
      dplyr::summarize(emissions_mt = sum(emissions_mt)) |>
      dplyr::ungroup() |>
      dplyr::mutate(fleet = "Total")
  )

# Aggregate emissions across fishing/non-fishing
annual_emissions <- annual_emissions_all_pollutants_with_total |>
  dplyr::group_by(year, fleet, pollutant) |>
  dplyr::summarise(emissions_mt = sum(emissions_mt)) |>
  dplyr::ungroup()

# Calculate the total percent change in emissions by fleet
total_percent_change_by_fleet <- annual_emissions |>
  dplyr::filter(year %in% c(2016, 2024), fleet != "Total") |>
  tidyr::pivot_wider(names_from = year, values_from = emissions_mt) |>
  dplyr::arrange(pollutant) |>
  dplyr::group_by(pollutant) |>
  dplyr::mutate(percent_total_2024 = `2024` / sum(`2024`)) |>
  dplyr::ungroup() %>%
  split(.[, c("pollutant")]) %>%
  purrr::map_df(janitor::adorn_totals, fill = NA) |>
  tidyr::fill(pollutant, .direction = "down") |>
  # janitor::adorn_totals("row") |>
  dplyr::mutate(percent_total_2024 = scales::percent(percent_total_2024)) |>
  dplyr::mutate(percent_change = 100 * (`2024` - `2016`) / `2016`) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS (broadcasting)",
      TRUE ~ "S1 (non-broadcasting)"
    )
  )
```

```{r}
#| label: wrangle-gfw-spatials-emissions-data
# Total spatial emissions by pollutant for 2024
total_spatial_emissions_by_pollutant_2024 <- total_spatial_emissions_by_pollutant_2024 |>
  tidyr::pivot_longer(
    -c(lat_bin, lon_bin),
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_total_mt") |>
      stringr::str_to_upper() |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  )
```

# In-line manuscript statistics

- The transportation sector is a major source of global greenhouse gas (GHG) and local air pollution emissions, responsible for `r fraction_co2_from_transportation` of all CO$_{2}$ emissions globally in 2023 

- For over `r scales::comma(signif(n_unique_vessels$n_unique_vessels,2))` AIS-broadcasting vessels, we quantify point-estimate emissions for each vessel using precise latitude/longitude/timestamp information for over `r scales::comma(signif(n_ais_messages$n_billion_ais_messages,3))` billion individual AIS messages

# Figures

## Figure 1

### Figure 1, panel A

```{r}
# For figures, order pollutants first for GHG pollutants,
# then local pollutants
pollutant_factor_levels <- c(
  "CO[2]",
  "CH[4]",
  "N[2]*O",
  "CO",
  "NO[X]",
  "SO[X]",
  "PM[2.5]",
  "PM[10]",
  "VOCs"
)
pollutant_map_data_co2 <- total_spatial_emissions_by_pollutant_2024 |>
  dplyr::filter(pollutant %in% c("CO2"))

make_pollutant_map <- function(pollutant_map_data, scale_label) {
  ggplot() +
    geom_sf(data = world_bbox_sf, fill = "black") +
    ggspatial::geom_spatial_tile(
      data = pollutant_map_data,
      aes(
        x = lon_bin + 0.5,
        y = lat_bin + 0.5,
        color = emissions_mt,
        fill = emissions_mt
      )
    ) +
    geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
    scale_fill_viridis_c(
      option = "mako",
      name = scale_label,
      breaks = c(0.0001, 10, 10000, 10e6),
      labels = c("0.0001", "10", "10k", "10M"),
      trans = "log10"
    ) +
    scale_color_viridis_c(
      option = "mako",
      name = scale_label,
      breaks = c(0.0001, 10, 10000, 10e6),
      labels = c("0.0001", "10", "10k", "10M"),
      trans = "log10"
    ) +
    labs(x = "", y = "") +
    theme_void() +
    guides(
      fill = guide_colorbar(
        title.position = "top",
        title.hjust = 0.5,
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    scale_x_continuous(expand = c(0, 1)) +
    scale_y_continuous(expand = c(0, 1))
}

co2_map <- make_pollutant_map(
  pollutant_map_data_co2,
  scale_label = expression(2024 ~ CO[2] ~ emissions ~ (MT))
) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.height = unit(0.75, "cm"),
    legend.key.width = unit(2, "cm")
  )
```

### Figure 1, panel B

```{r}
# Calculate annual terrestrial transportation emissions from EDGAR
land_based_transportation_emissions_edgar <- annual_edgar_tranportation_emissions |>
  dplyr::filter(land_based_tranpsportation) |>
  dplyr::group_by(year) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::mutate(data_source = "Terrestrial transportation (EDGAR)")


all_tranportation_data <- land_based_transportation_emissions_edgar |>
  dplyr::bind_rows(
    annual_emissions |>
      dplyr::filter(pollutant == "co2", fleet == "Total") |>
      dplyr::rename(emissions_co2_mt = emissions_mt) |>
      dplyr::mutate(data_source = "Marine vessels (AIS + S1)") |>
      dplyr::select(-c(fleet, pollutant))
  )


# Get 2016 values for time-series normalization
all_tranportation_data_starting <- all_tranportation_data |>
  dplyr::filter(year == 2016) |>
  dplyr::mutate(emissions_co2_mt_starting = emissions_co2_mt) |>
  dplyr::select(-c(year, emissions_co2_mt))

all_tranportation_data_normalized <- all_tranportation_data |>
  dplyr::inner_join(all_tranportation_data_starting, by = "data_source") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_co2_mt /
      emissions_co2_mt_starting -
      1
  )

emissions_comparison_figure_normalized <- all_tranportation_data_normalized |>
  dplyr::filter(
    data_source %in%
      c("Terrestrial transportation (EDGAR)", "Marine vessels (AIS + S1)")
  ) |>
  ggplot(aes(x = year, y = emissions_co2_mt_normalized, color = data_source)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(linewidth = 1.1) +
  scale_x_continuous(
    breaks = unique(all_tranportation_data_normalized$year),
    limits = c(2016, 2023)
  ) +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(-0.15, 0.25),
    breaks = seq(-.15, .25, .05)
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of ~ CO[2] ~ emissions,
      from ~ 2016 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.25, 0.875),
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", color = "gray")
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c(
      "Marine vessels (AIS + S1)",
      "Terrestrial transportation (EDGAR)"
    )
  )
```

### Figure 1, combined

```{r}
#| label: fig-co2-emissions-map-normalized-sector-time-series
#| #| fig-cap: (A) Spatial map of CO2 2024 emissions, aggregated across AIS-broadcasting vessels and non-broadcasting vessels, and shown at a 1x1 degree spatial resolution; (B) Change in annual CO$_{2}$ emissions from a 2016 baseline, by sector
#| fig-width: 7
#| fig-height: 7
cowplot::plot_grid(
  co2_map,
  emissions_comparison_figure_normalized,
  ncol = 1,
  labels = c("A", "B"),
  rel_heights = c(0.55, 0.45)
)
```

## Figure 2 

### Figure 2, panel A

```{r}
total_annual_emissions_by_fleet_fig <- annual_emissions |>
  dplyr::filter(pollutant == "co2") |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS (broadcasting)",
      TRUE ~ "S1 (non-broadcasting)"
    )
  ) |>
  ggplot(aes(x = year, y = emissions_mt, color = fleet)) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "B", scale = 1e-9),
    limits = c(0, 1.5e9),
    breaks = seq(0, 1.5e9, 0.25e9)
  ) +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(2016, 2024)
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS + S1", "AIS (broadcasting)", "S1 (non-broadcasting)")
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(Annual ~ CO[2] ~ emissions ~ (MT))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.75, 0.4),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  )
```

### Figure 2, panel B

```{r}
# Get 2016 values for time-series normalization
annual_co2_emissions_starting <- annual_emissions |>
  dplyr::filter(pollutant == "co2", year == 2016) |>
  dplyr::mutate(emissions_co2_mt_starting = emissions_mt) |>
  dplyr::select(-c(year, emissions_mt))

normalized_annual_emissions_by_fleet_fig <- annual_emissions |>
  dplyr::filter(pollutant == "co2") |>
  dplyr::left_join(annual_co2_emissions_starting, by = "fleet") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_mt / emissions_co2_mt_starting - 1
  ) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS (broadcasting)",
      TRUE ~ "S1 (non-broadcasting)"
    )
  ) |>
  ggplot(aes(x = year, y = emissions_co2_mt_normalized, color = fleet)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(-0.2, 0.35),
    breaks = seq(-.2, .35, .05)
  ) +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(2016, 2024)
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS (broadcasting)", "AIS + S1", "S1 (non-broadcasting)")
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of ~ CO[2] ~ emissions,
      from ~ 2016 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.25, 0.85),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  )
```


### Figure 2 combined

```{r}
#| label: fig-total-annual-emissions-by-data-source
#| fig-cap: (A) Annual marine vessel CO2 emissions, with each line colored based on the data source. (B) Relative change in annual marine vessel CO2 emissions from a 2016 baseline, with each line colored based on the data source.
#| fig-width: 6
#| fig-height: 6.5
#|
cowplot::plot_grid(
  total_annual_emissions_by_fleet_fig,
  normalized_annual_emissions_by_fleet_fig,
  ncol = 1,
  align = "v",
  axis = "lr",
  labels = c("A", "B")
)
```

## Figure 3

```{r}
#| label: build-decomposition-model

# For aggregate annual extrapolation dataset, calculated total emissions, kw-hours,distance, hours
annual_extrapolation_dataset <- annual_ais_to_dark_activity_extrapolation |>
  dplyr::mutate(
    emissions_co2_mt_total = emissions_co2_mt + emissions_co2_mt_dark,
    kw_hours_total = kw_hours + kw_hours_dark,
    distance_nm_total = distance_nm + distance_nm_dark,
    hours_total = hours + hours_dark
  ) |>
  dplyr::select(
    -c(
      emissions_co2_mt,
      emissions_co2_mt_dark,
      kw_hours,
      kw_hours_dark,
      distance_nm,
      distance_nm_dark,
      hours,
      hours_dark,
      avg_speed_knots
    )
  ) |>
  # Calculate average aggregate speed as total distance traveled divided by total hours
  dplyr::mutate(avg_speed_knots = distance_nm_total / hours_total)

# Using annual_extrapolation_dataset, use tidymodels to train a random forest model
# to predict emissions
set.seed(123)

# Set initial split of data
inital_split <- annual_extrapolation_dataset |>
  rsample::initial_split(prop = 0.8)

training_data <- rsample::training(inital_split)
testing_data <- rsample::testing(inital_split)

# Specify recipe for model
model_recipe <- recipes::recipe(
  emissions_co2_mt_total ~
    fishing + length_size_class_percentile + kw_hours_total + avg_speed_knots,
  data = head(rsample::training(inital_split))
) |>
  recipes::step_num2factor(
    length_size_class_percentile,
    levels = as.character(unique(
      annual_extrapolation_dataset$length_size_class_percentile
    ))
  )

# Specify model type
model_type <- parsnip::rand_forest(
  trees = 1000,
  # Default of ranger is squareroot of number of model features. We have 4 features, so use a value of 2.
  mtry = 2,
  # Default of ranger is 5 for regression
  min_n = 5
) |>
  parsnip::set_engine("ranger") |>
  parsnip::set_mode("regression")

# Initialize model workflow
model_workflow <- workflows::workflow() |>
  workflows::add_recipe(model_recipe) |>
  workflows::add_model(model_type)

# Fit model on training data
workflow_fit <- parsnip::fit(model_workflow, training_data)

# Make predictions on testing data, and augment this to testing data frame
worflow_predictions <- parsnip::augment(workflow_fit, testing_data)


# Calculate performance metrics

# Multiple regression metrics
multi_metric <- yardstick::metric_set(
  yardstick::rsq,
  yardstick::mape,
  yardstick::rsq_trad
)

performance_metrics <- multi_metric(
  data = worflow_predictions,
  truth = emissions_co2_mt_total,
  estimate = .pred
) |>
  dplyr::select(-.estimator)

# Final model on all data
# Fit final model on all data
workflow_fit_final <- parsnip::fit(model_workflow, annual_extrapolation_dataset)
```

```{r}
#| label: tbl-decomposition-model-performance
#| tbl-cap: Performance metrics for random forest model that predicts annual CO2 emissions, by vessel size class and type, using hours and average speed.

performance_metrics |>
  knitr::kable(digits = 3)
```

```{r}
#| label: predict-using-decomposition-model
# Find initial/starting 2016 values of kw-hours and speed, by fishing and length_size_class_percentile, for decomposition
starting_values <- annual_extrapolation_dataset |>
  dplyr::filter(year == 2016) |>
  dplyr::select(
    fishing,
    length_size_class_percentile,
    kw_hours_total_starting = kw_hours_total,
    hours_total_starting = hours_total,
    avg_speed_knots_starting = avg_speed_knots
  )

# Add starting values to predictions dataset
predictions_dataset <- annual_extrapolation_dataset |>
  dplyr::left_join(
    starting_values,
    by = c("fishing", "length_size_class_percentile")
  )

final_model_fit <- workflows::extract_fit_parsnip(workflow_fit_final)

# Now run predictions for 4 different scenarios
final_prediction_dataset <- predictions_dataset |>
  # Add model predictions using observed kw-hours and observed speed
  dplyr::bind_cols(
    observed_kw_hours_observed_speed = predict(
      final_model_fit,
      predictions_dataset
    )$.pred
  ) |>
  # Add model predictions using observed kw-hours and constant speed (from 2016)
  dplyr::bind_cols(
    observed_kw_hours_constant_speed = predict(
      final_model_fit,
      predictions_dataset |>
        dplyr::select(-avg_speed_knots) |>
        dplyr::rename(avg_speed_knots = avg_speed_knots_starting)
    )$.pred
  ) |>
  # Add model predictions using constant kw-hours (from 2016) and observed speed
  dplyr::bind_cols(
    constant_kw_hours_observed_speed = predict(
      final_model_fit,
      predictions_dataset |>
        dplyr::select(-kw_hours_total) |>
        dplyr::rename(kw_hours_total = kw_hours_total_starting)
    )$.pred
  ) |>
  # Add model predictions using constant kw-hours (from 2016) and constant speed (from 2016)
  dplyr::bind_cols(
    constant_kw_hours_constant_speed = predict(
      final_model_fit,
      predictions_dataset |>
        dplyr::select(-avg_speed_knots) |>
        dplyr::rename(avg_speed_knots = avg_speed_knots_starting) |>
        dplyr::select(-kw_hours_total) |>
        dplyr::rename(kw_hours_total = kw_hours_total_starting)
    )$.pred
  )
```

## Figure 3

```{r}
#| label: fig-hours-speed-decomposition
#| fig-cap: Estimated decomposition of change in total CO2 emissions from 2016 baseline (aggregating across both broadcasting and non-broadcasting vessels), using the random forest model outputs of four different scenarios that decompose the relative impact of changes in total kW-hours and changes in average speed/
#| fig-width: 7.5
#| fig-height: 3
# Aggregate prediction values by year (across vessel size classes and type)
final_annual_prediction_dataset <- final_prediction_dataset |>
  dplyr::group_by(year) |>
  dplyr::summarise(across(
    c(
      observed_kw_hours_observed_speed,
      observed_kw_hours_constant_speed,
      constant_kw_hours_observed_speed,
      constant_kw_hours_constant_speed
    ),
    ~ sum(., na.rm = TRUE)
  )) |>
  dplyr::rename(
    `Observed kW-hours, observed speed` = observed_kw_hours_observed_speed,
    `Observed kW-hours, constant speed` = observed_kw_hours_constant_speed,
    `Constant kW-hours, observed speed` = constant_kw_hours_observed_speed,
    `Constant kW-hours, constant speed` = constant_kw_hours_constant_speed
  ) |>
  tidyr::pivot_longer(-year)

# We will normalize all of these predictions to the 2016 starting prediction
starting_value_from_prediction <- final_annual_prediction_dataset |>
  dplyr::filter(year == 2016) |>
  dplyr::rename(starting_value = value) |>
  dplyr::select(-year)

time_series_predicted_decomposition_df <- final_annual_prediction_dataset |>
  dplyr::inner_join(starting_value_from_prediction, by = "name") |>
  dplyr::mutate(normalized_value = value / starting_value - 1)

time_series_predicted_decomposition_df |>
  ggplot(aes(x = year, y = normalized_value, color = name)) +
  geom_line(linewidth = 1.1) +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_color_manual(
    "Prediction scenario",
    values = c(
      "coral3",
      "coral4",
      paletteer::paletteer_d("colorblindr::OkabeIto")[[2]],
      paletteer::paletteer_d("colorblindr::OkabeIto")[[5]]
    )
  ) +
  scale_x_continuous(breaks = unique(final_prediction_dataset$year)) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "",
    y = ""
  ) +
  theme_plot() +
  theme(panel.grid.minor.x = element_blank())
```

# Supplementary figures

```{r}
#| label: fig-pollutant-maps
#| fig-cap: Spatial map of 2024 emissions for each pollutant, aggregated across the AIS-broadcasting fleet and non-broadcasting, and shown at a 1x1 degree spatial resolution
#| fig-width: 8
#| fig-height: 6
pollutant_map_data_all <- total_spatial_emissions_by_pollutant_2024 |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

make_pollutant_map(
  pollutant_map_data_all,
  scale_label = "2024 emissions (MT)"
) +
  facet_wrap(pollutant ~ ., ncol = 1, labeller = label_parsed) +
  facet_wrap(
    pollutant ~ .,
    ncol = 3,
    labeller = label_parsed
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.height = unit(1.5, "cm"),
    legend.key.width = unit(1.5, "cm")
  )
```

```{r}
#| label: fig-co2-emissions-time-series
#| fig-cap: Time series of monthly global CO2 emissions, aggregated across AIS-broadcasting vessels and non-broadcasting vessels.
#| fig-width: 8
#| fig-height: 3

# Total monthly non-spatial emissions by pollutant
total_monthly_emissions_by_pollutant <- total_monthly_emissions_by_pollutant |>
  dplyr::mutate(month = lubridate::ymd(month)) |>
  tidyr::pivot_longer(
    -month,
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_total_mt") |>
      stringr::str_to_upper() |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::filter(
    lubridate::year(month) <= analysis_end_year,
    lubridate::year(month) >= analysis_start_year
  )

pollutant_time_series_data_co2 <- total_monthly_emissions_by_pollutant |>
  dplyr::filter(pollutant == "CO2")

make_pollutant_time_series_figure <- function(
    pollutant_time_series_data,
    y_axis_label) {{ ggplot(
  data = pollutant_time_series_data,
  aes(x = month, y = emissions_mt)
) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::comma) +
  theme_plot() +
  labs(x = "", y = y_axis_label) +
  theme(panel.grid.minor.x = element_blank()) +
  scale_y_continuous(limits = c(0, NA)) }}

make_pollutant_time_series_figure(
  pollutant_time_series_data_co2,
  y_axis_label = expression(Monthly ~ CO[2] ~ emissions ~ (MT))
) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  )
```

```{r}
#| label: fig-pollutant-time-series
#| fig-cap: Time series of monthly global emissions for each pollutant, aggregated across AIS-broadcasting vessels and non-broadcasting vessels.
#| fig-width: 8
#| fig-height: 8

pollutant_time_series_data_all <- total_monthly_emissions_by_pollutant |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

make_pollutant_time_series_figure(
  pollutant_time_series_data_all,
  y_axis_label = "Monthly emissions (MT)"
) +
  facet_wrap(
    pollutant ~ .,
    ncol = 3,
    scales = "free_y",
    labeller = label_parsed
  )
```

```{r}
#| label: fig-co2-emissions-by-vessel-type
#| fig-cap: Total 2024 CO2 emissions by vessel type for AIS-broadcasting vessels (millions metric tonnes).
#| fig-width: 4
#| fig-height: 5
annual_ais_co2_emissions_by_vessel_type |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    fishing = ifelse(
      vessel_class %in%
        c(
          "trawlers",
          "squid_jigger",
          "tuna_purse_seines",
          "set_longlines",
          "pole_and_line",
          "set_gillnets",
          "pots_and_traps",
          "dredge_fishing",
          "other_seines",
          "other_purse_seines",
          "trollers"
        ),
      "Fishing",
      "Non-fishing"
    ) |>
      forcats::fct_rev()
  ) |>
  dplyr::mutate(
    vessel_class = stringr::str_replace_all(
      vessel_class,
      "cargo.",
      "cargo: "
    ) |>
      stringr::str_replace_all("_", " ") |>
      stringr::str_to_sentence()
  ) |>
  ggplot(aes(
    x = emissions_co2_mt,
    y = reorder(vessel_class, emissions_co2_mt)
  )) +
  geom_col(position = position_dodge2(preserve = "single")) +
  ggforce::facet_col(~fishing, space = "free", scales = "free") +
  theme_plot() +
  labs(
    x = "2024 global CO2 emissions (MT)",
    y = ""
  ) +
  scale_x_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  )
```