---
title: "Quantifying comprehensive marine vessel emissions using data fusion"
author: "emLab, UCSB and GFW"
date: last-modified
date-format: "[Last Updated on] MMMM DD, YYYY"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    fig-pos: H
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
cache: true
---

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE, cache = FALSE}
#| label: document-setup

# Set default chunk options
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  dev = "png",
  fig.path = "../figures/"
)

# Load librariers
library(ggplot2)
library(sf)

analysis_start_year <- 2017
analysis_end_year <- 2024
```

```{r}
#| label: load-data
#| cache: FALSE
targets::tar_load(c(
  "n_unique_vessels",
  "n_ais_messages",
  "fraction_vessels_emissions_by_registry_info",
  "n_s1_detections",
  "annual_edgar_emissions",
  "oecd_data",
  "all_performance_metrics",
  "monthly_aggregated_time_series",
  "annual_emissions_all_pollutants",
  "total_spatial_emissions_by_pollutant",
  "total_monthly_emissions_by_pollutant",
  "port_visit_co2_emissions_by_country",
  "trip_co2_emissions_by_from_to_countries",
  "annual_global_emissions_by_receiver_type",
  "annual_global_emissions_by_receiver_type_and_flag",
  "annual_spatial_emissions_by_receiver_type",
  "vessel_size_info",
  "registered_validation_data",
  "mrv_data_validation",
  "trip_emissions_for_mrv_validation",
  "registered_validation_data",
  "all_varimp_data",
  "performance_detections_cls_roc_pr_curves",
  "performance_detections_cls_conf_mat",
  "annual_spatial_co2_emissions_ais_dark_by_fleet",
  "total_spatial_co2_emissions_dark_by_footprint",
  "data_sources",
  "pixels_for_offshore_training_testing_split",
  "total_spatial_co2_emissions_by_ocean",
  "number_s1_imaged_months_by_pixel",
  "lm_other_gases_tidy_fit_stats",
  "s1_time_series",
  "length_size_bin_distributions",
  "ping_level_hours_distribution"
))
```

```{r}
#| label: setup-graphics

# For figures and tables, order pollutants first for GHG pollutants,
# then local pollutants. For figures, we use expressions to show subscripts
pollutant_factor_levels <- c(
  "CO[2]",
  "CH[4]",
  "N[2]*O",
  "CO",
  "NO[X]",
  "SO[X]",
  "PM[2.5]",
  "PM[10]",
  "VOCs"
)

# Set pollutant levels for tables
# For tables, we don't use expressions
pollutant_factor_levels_tables <- pollutant_factor_levels |>
  stringr::str_remove_all("\\[") |>
  stringr::str_remove_all("\\]") |>
  stringr::str_remove_all("\\*]*")

# Use Okabe-Ito colorblind-friendly colors
all_data_source_color_palette <- c(
  "Marine vessels (AIS + S1)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[5]],
  "GFW (AIS + S1)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[5]],
  "AIS + S1" = paletteer::paletteer_d("colorblindr::OkabeIto")[[5]],
  "Other transportation sectors (EDGAR)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[7]],
  "GFW (AIS)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[1]],
  "AIS" = paletteer::paletteer_d("colorblindr::OkabeIto")[[1]],
  "S1" = paletteer::paletteer_d("colorblindr::OkabeIto")[[
    6
  ]],
  "All sectors (EDGAR)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[8]],
  "EDGAR" = paletteer::paletteer_d("colorblindr::OkabeIto")[[8]],
  "OECD" = paletteer::paletteer_d("colorblindr::OkabeIto")[[6]],
  "IMO" = paletteer::paletteer_d("colorblindr::OkabeIto")[[3]],
  "Fishing" = paletteer::paletteer_d("colorblindr::OkabeIto")[[5]],
  "Non-fishing" = paletteer::paletteer_d("colorblindr::OkabeIto")[[2]]
)

# Set consistent theme for all plots
theme_plot <- function() {
  theme_minimal() %+replace%
    theme(
      axis.title.x = element_text(face = "bold", size = 12),
      axis.title.y = element_text(
        angle = 90,
        face = "bold",
        vjust = 3,
        size = 12
      ),
      strip.text.x = element_text(angle = 0, face = "bold", size = 14),
      strip.text.y = element_text(angle = 0, face = "bold", size = 14)
    )
}

# Set map projection for all map figures
map_projection <- "+proj=eqearth"

# Create bounding box of globe, for making black map backgrounds
lats <- c(90:-90, -90:90, 90)
longs <- c(rep(c(180, -180), each = 181), 180)
world_bbox_sf <-
  list(cbind(longs, lats)) %>%
  sf::st_polygon() %>%
  sf::st_sfc(
    # create sf geometry list column
    crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
  ) %>%
  sf::st_sf() %>%
  sf::st_transform(map_projection)

# ESRI World Countries Generalized Shapefile
# Downloaded from https://www.arcgis.com/home/item.html?id=fa510018bdd044b08fc64d2a16bc680a
land <- sf::st_read(
  here::here(
    "data/World_Countries_Generalized_Shapefile/World_Countries_Generalized.shp"
  ),
  quiet = TRUE
) |>
  dplyr::mutate(
    country_iso3 = countrycode::countrycode(ISO, "iso2c", "iso3c")
  )


# Convert lat/lon pixel area to square km
# https://stackoverflow.com/a/1253545
# Need function to convert degrees to radians
# https://stackoverflow.com/q/32370485
deg2rad <- function(deg) {
  (deg * pi) / (180)
}

# Given a latitude and pixel size, what is a pixel's area?
# https://stackover flow.com/a/1253545
lat_to_area <- function(lat, pixel_size) {
  110.574 * (111.320 * cos(deg2rad(lat + pixel_size / 2)))
}

# We use a 1x1 degree pixel size for our spatial emissions data, so set pixel_size to 1 degree
pixel_size_analysis <- 1

# Set map background color as the midpoint value from scico::scico vik, which we use for
map_background_color <- scico::scico(
  n = 1,
  palette = "vik",
  begin = 0.5,
  end = 0.5
)
```

```{r}
#| label: wrangle-edgar-data
annual_edgar_tranportation_emissions <- annual_edgar_emissions |>
  dplyr::select(
    -c(
      IPCC_annex,
      C_group_IM24_sh,
      Country_code_A3,
      Name,
      ipcc_code_2006_for_standard_report,
      Substance,
      fossil_bio
    )
  ) |>
  dplyr::rename(sector = ipcc_code_2006_for_standard_report_name) |>
  tidyr::pivot_longer(
    cols = -c(sector),
    names_to = "year",
    values_to = "emissions_co2_kt"
  ) |>
  dplyr::mutate(
    year = stringr::str_remove_all(year, "Y_") |>
      as.numeric(),
    emissions_co2_mt = emissions_co2_kt * 1e3
  ) |>
  dplyr::select(-emissions_co2_kt) |>
  dplyr::mutate(
    land_based_tranpsportation = ifelse(
      sector %in%
        c(
          "Civil Aviation",
          "Road Transportation no resuspension",
          "Other Transportation",
          "Railways"
        ),
      TRUE,
      FALSE
    )
  ) |>
  # Only need years that correspond to our dataset
  dplyr::filter(year >= analysis_start_year, year <= analysis_end_year)

# Fraction of global 2024 CO2 emissions that came from transportation sectors
fraction_co2_from_transportation <- annual_edgar_tranportation_emissions |>
  dplyr::mutate(
    transportation = ifelse(
      land_based_tranpsportation | sector == "Water-borne Navigation",
      TRUE,
      FALSE
    )
  ) |>
  dplyr::filter(year == 2024) |>
  dplyr::group_by(transportation) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(
    names_from = "transportation",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::mutate(fraction_from_transportation = `TRUE` / (`TRUE` + `FALSE`)) |>
  dplyr::pull(fraction_from_transportation) |>
  scales::percent()
```

```{r}
#| label: wrangle-gfw-time-series-emissions-data

annual_emissions_all_pollutants <- annual_emissions_all_pollutants |>
  tidyr::pivot_longer(
    -c(year, fishing),
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    fleet = ifelse(
      stringr::str_detect(pollutant, "dark"),
      "Non-broadcasting",
      "AIS-broadcasting"
    )
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_mt") |>
      stringr::str_remove_all("_dark")
  )

annual_emissions_all_pollutants_with_total <- annual_emissions_all_pollutants |>
  # Add total, that aggregates across broadcasting and non-broadcasting
  dplyr::bind_rows(
    annual_emissions_all_pollutants |>
      dplyr::group_by(year, fishing, pollutant) |>
      dplyr::summarize(emissions_mt = sum(emissions_mt)) |>
      dplyr::ungroup() |>
      dplyr::mutate(fleet = "Total")
  )

# Aggregate emissions across fishing/non-fishing
annual_emissions <- annual_emissions_all_pollutants_with_total |>
  dplyr::group_by(year, fleet, pollutant) |>
  dplyr::summarise(emissions_mt = sum(emissions_mt)) |>
  dplyr::ungroup()

# Calculate the total percent change in emissions by fleet
total_percent_change_by_fleet <- annual_emissions |>
  dplyr::filter(
    year %in% c(analysis_start_year, analysis_end_year),
    fleet != "Total"
  ) |>
  tidyr::pivot_wider(names_from = year, values_from = emissions_mt) |>
  dplyr::arrange(pollutant) |>
  dplyr::group_by(pollutant) |>
  dplyr::mutate(
    percent_total_2024 = (`2024` / sum(`2024`))
  ) |>
  dplyr::ungroup() %>%
  split(.[, c("pollutant")]) %>%
  purrr::map_df(janitor::adorn_totals, fill = NA) |>
  tidyr::fill(pollutant, .direction = "down") |>
  dplyr::mutate(
    percent_change = ((`2024` - `2017`) /
      `2017`)
  ) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS (broadcasting)",
      TRUE ~ "S1 (non-broadcasting)"
    )
  )
```

```{r}
# For each pollutant, estimate how much emissions are underestimated by just using AIS
# As compared to using AIS + S1
ais_scale_underestimation_2024 <- total_percent_change_by_fleet |>
  dplyr::filter(fleet != "S1 (non-broadcasting)") |>
  dplyr::select(fleet, pollutant, `2024`) |>
  tidyr::pivot_wider(names_from = "fleet", values_from = `2024`) |>
  dplyr::mutate(
    ais_underestimation_2024 = (1 - `AIS (broadcasting)` / `AIS + S1`) |>
      scales::percent(accuracy = 0.01)
  ) |>
  dplyr::rename(
    emissions_mt_2024_ais = `AIS (broadcasting)`,
    emissions_mt_2024_ais_s1 = `AIS + S1`
  )

# For each pollutant, estimate how much the 2017->2024 trend in emissions is underestimated by just using AIS
# As compared to using AIS + S1
ais_trend_overestimation_2017_2024 <- total_percent_change_by_fleet |>
  dplyr::filter(fleet != "S1 (non-broadcasting)") |>
  dplyr::select(fleet, pollutant, percent_change) |>
  tidyr::pivot_wider(names_from = "fleet", values_from = percent_change) |>
  dplyr::mutate(
    trend_overestimation_2017_2024 = (`AIS (broadcasting)` /
      `AIS + S1` -
      1) |>
      scales::percent(accuracy = 0.01)
  ) |>
  dplyr::rename(
    trend_2017_2024_ais = `AIS (broadcasting)`,
    trend_2017_2024_ais_s1 = `AIS + S1`
  )

# Put all this information together
pollutant_ais_underestimation_overestimation_summary <- ais_scale_underestimation_2024 |>
  dplyr::left_join(ais_trend_overestimation_2017_2024, by = "pollutant") |>
  dplyr::select(
    pollutant,
    ais_underestimation_2024,
    trend_overestimation_2017_2024
  )
```

```{r}
#| label: wrangel-edgar-data
# Calculate annual other transportation emissions from EDGAR
land_based_transportation_emissions_edgar <- annual_edgar_tranportation_emissions |>
  dplyr::filter(land_based_tranpsportation) |>
  dplyr::group_by(year) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::mutate(data_source = "Other transportation sectors (EDGAR)")


global_emissions_edgar <- annual_edgar_tranportation_emissions |>
  dplyr::group_by(year) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::mutate(data_source = "All sectors (EDGAR)")

all_edgar_gfw_data <- land_based_transportation_emissions_edgar |>
  dplyr::bind_rows(global_emissions_edgar) |>
  dplyr::bind_rows(
    annual_emissions |>
      dplyr::filter(pollutant == "co2", fleet == "Total") |>
      dplyr::rename(emissions_co2_mt = emissions_mt) |>
      dplyr::mutate(data_source = "Marine vessels (AIS + S1)") |>
      dplyr::select(-c(fleet, pollutant))
  )

# Get analysis_start_year values for time-series normalization
all_edgar_gfw_data_starting <- all_edgar_gfw_data |>
  dplyr::filter(year == analysis_start_year) |>
  dplyr::mutate(emissions_co2_mt_starting = emissions_co2_mt) |>
  dplyr::select(-c(year, emissions_co2_mt))

all_edgar_gfw_data_normalized <- all_edgar_gfw_data |>
  dplyr::inner_join(all_edgar_gfw_data_starting, by = "data_source") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_co2_mt /
      emissions_co2_mt_starting -
      1
  )


percent_transportation_emmissions_from_marine_2024 <- all_edgar_gfw_data |>
  dplyr::filter(year == 2024) |>
  dplyr::select(-year) |>
  tidyr::pivot_wider(
    names_from = "data_source",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::mutate(
    percent = scales::percent(
      `Marine vessels (AIS + S1)` /
        (`Marine vessels (AIS + S1)` + `Other transportation sectors (EDGAR)`)
    )
  ) |>
  dplyr::pull(percent)

percent_increase_from_2017_to_2024 <- all_edgar_gfw_data_normalized |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    emissions_co2_mt_normalized_percent = scales::percent(
      emissions_co2_mt_normalized
    )
  )
```


```{r}
# Prep data for comparing different inventories
marine_based_transportation_emissions_edgar <- annual_edgar_tranportation_emissions |>
  dplyr::filter(sector == "Water-borne Navigation") |>
  dplyr::group_by(year) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::mutate(data_source = "EDGAR")

imo_data <- tibble::tibble(
  year = c(2016, 2017, 2018),
  emissions_co2_mt = c(1.026e9, 1.064e9, 1.056e9)
) |>
  dplyr::mutate(data_source = "IMO")

all_inventory_data <- all_edgar_gfw_data |>
  dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>
  dplyr::mutate(data_source = "GFW (AIS + S1)") |>
  dplyr::bind_rows(marine_based_transportation_emissions_edgar) |>
  dplyr::bind_rows(oecd_data) |>
  dplyr::bind_rows(imo_data) |>
  dplyr::bind_rows(
    annual_emissions |>
      dplyr::filter(pollutant == "co2", fleet == "AIS-broadcasting") |>
      dplyr::select(year, emissions_co2_mt = emissions_mt) |>
      dplyr::mutate(data_source = "GFW (AIS)")
  ) |>
  dplyr::filter(year >= analysis_start_year, year <= analysis_end_year)
```

```{r}
#| label: wrangle-gfw-spatials-emissions-data
# Total spatial emissions by pollutant for analysis_start_year and 2024
total_spatial_emissions_by_pollutant_formatted <- total_spatial_emissions_by_pollutant |>
  tidyr::pivot_longer(
    -c(lat_bin, lon_bin, year),
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_total_mt") |>
      stringr::str_to_upper() |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(
    emissions_mt_per_km2 = emissions_mt /
      lat_to_area(lat_bin, pixel_size_analysis)
  ) |>
  dplyr::select(-emissions_mt) |>
  tidyr::pivot_wider(
    names_from = "year",
    values_from = "emissions_mt_per_km2",
    names_prefix = "emissions_mt_per_km2_"
  )
```

```{r}
# Summarize fraction of all estimate dark emissions that occurred inside the S1 footprint and in months that were imaged; inside the S1 footprint and in months that weren't imaged; and outside the S1 footprint entirely
dark_emissions_by_area <- monthly_aggregated_time_series |>
  dplyr::group_by(inside_footprint, s1_imaged) |>
  dplyr::summarise(emissions_co2_dark_mt = sum(emissions_co2_dark_mt)) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    fraction = emissions_co2_dark_mt / sum(emissions_co2_dark_mt),
    percent_total = ifelse(
      fraction < 0.01,
      scales::percent(
        fraction,
        accuracy = 0.001
      ),
      scales::percent(fraction, accuracy = 0.1)
    )
  )

# Summarize fraction of 2024 CO2 emissions that came from non-broadcasting vessels,
# disaggregated by fishing and non-fishing
dark_emissions_by_fishing <- monthly_aggregated_time_series |>
  dplyr::filter(lubridate::year(time) == 2024) |>
  dplyr::group_by(fishing) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    percent_dark = scales::percent(
      emissions_co2_dark_mt / (emissions_co2_dark_mt + emissions_co2_mt),
      accuracy = 0.1
    )
  )

monthly_fraction_fig <- monthly_aggregated_time_series |>
  dplyr::group_by(fishing, year = lubridate::year(time)) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    fraction = emissions_co2_dark_mt /
      (emissions_co2_dark_mt + emissions_co2_mt)
  ) |>
  ggplot(aes(x = year, y = fraction, color = fishing)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA))
```


```{r}
#| label: social-cost-calculation

# Social cost of carbon
# From Moore et al. 2024: https://www.pnas.org/doi/abs/10.1073/pnas.2410733121
social_cost_carbon_usd_per_mt_co2 <- 283

# Social cost of methane
# From Wang and Teng 2023: https://www.nature.com/articles/s41558-023-01803-4
social_cost_methane_usd_per_mt_ch4 <- 2900

# Social cost of n2o
# From Wang and Teng 2023: https://www.nature.com/articles/s41558-023-01803-4
social_cost_nitrous_oxide_usd_per_mt_n2o <- 49600

social_cost_tibble <- tibble::tibble(
  pollutant = c("co2", "ch4", "n2o"),
  sc_usd_per_mt = c(
    social_cost_carbon_usd_per_mt_co2,
    social_cost_methane_usd_per_mt_ch4,
    social_cost_nitrous_oxide_usd_per_mt_n2o
  )
)

annual_sc_fishing_non_fishing <- annual_emissions_all_pollutants_with_total |>
  dplyr::filter(fleet %in% c("AIS-broadcasting", "Non-broadcasting")) |>
  dplyr::group_by(year, fishing, pollutant) |>
  dplyr::summarise(emissions_mt = sum(emissions_mt)) |>
  dplyr::ungroup() |>
  dplyr::inner_join(social_cost_tibble, by = "pollutant") |>
  dplyr::filter((year == 2021 & !fishing) | (year == 2022 & fishing)) |>
  dplyr::mutate(sc_billion_usd = emissions_mt * sc_usd_per_mt / 1e9) |>
  dplyr::mutate(emissions_mt = round(emissions_mt)) |>
  dplyr::mutate(dplyr::across(
    -c(year, sc_billion_usd),
    ~ prettyNum(., big.mark = ",")
  )) |>
  dplyr::group_split(fishing) |>
  purrr::map_df(~ janitor::adorn_totals(., where = "row")) |>
  dplyr::mutate(fishing = ifelse(fishing == "-", NA_character_, fishing)) |>
  tidyr::fill(fishing, .direction = "down") |>
  dplyr::mutate(
    fishing = ifelse(fishing == "TRUE", "Fishing", "Non-fishing")
  ) |>
  dplyr::mutate(pollutant = stringr::str_to_upper(pollutant))
```

```{r}
# Summarize number of vessels and total CO2 emissions for vessels that are:
# on the IMO registry; on some other registry; and without any registry info
fraction_vessels_emissions_by_registry_info_expanded <- fraction_vessels_emissions_by_registry_info |>
  dplyr::mutate(
    fraction_n_unique_vessels = n_unique_vessels / sum(n_unique_vessels)
  ) |>
  dplyr::mutate(
    fraction_emissions_co2_mt = emissions_co2_mt / sum(emissions_co2_mt)
  ) |>
  janitor::adorn_totals()
```

```{r}
# Prep trip and port stay data

# Divde international trip emissions equally to from and to countries
international_trip_emissions_by_country <- trip_co2_emissions_by_from_to_countries |>
  dplyr::filter(from_country_iso3 != to_country_iso3) |>
  dplyr::mutate(emissions_co2_mt = emissions_co2_mt / 2) |>
  dplyr::select(
    country_iso3 = from_country_iso3,
    emissions_co2_mt,
    vessel_class,
    n_unique_trips
  ) |>
  dplyr::bind_rows(
    trip_co2_emissions_by_from_to_countries |>
      dplyr::filter(from_country_iso3 != to_country_iso3) |>
      dplyr::mutate(
        emissions_co2_mt = emissions_co2_mt / 2
      ) |>
      dplyr::select(
        country_iso3 = to_country_iso3,
        emissions_co2_mt,
        vessel_class
      )
  ) |>
  dplyr::group_by(country_iso3, vessel_class) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt)
  ) |>
  dplyr::ungroup()

domestic_trip_emissions_by_country <- trip_co2_emissions_by_from_to_countries |>
  dplyr::filter(from_country_iso3 == to_country_iso3) |>
  dplyr::select(
    country_iso3 = from_country_iso3,
    emissions_co2_mt,
    vessel_class
  )

all_port_and_trip_emissions <- international_trip_emissions_by_country |>
  dplyr::mutate(type = "International trips") |>
  dplyr::bind_rows(
    domestic_trip_emissions_by_country |>
      dplyr::mutate(type = "Domestic trips")
  ) |>
  dplyr::bind_rows(
    port_visit_co2_emissions_by_country |>
      dplyr::rename(
        country_iso3 = port_country_iso3
      ) |>
      dplyr::mutate(type = "Port stays")
  ) |>
  dplyr::mutate(
    type = forcats::fct_relevel(type, "International trips", "Port stays")
  )

# Top 3 countries in terms of 2024 international trips, domestic trips, and port stays
all_port_and_trip_emissions_top_3 <- all_port_and_trip_emissions |>
  dplyr::group_by(country_iso3, type) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt)) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    country = countrycode::countrycode(
      country_iso3,
      "iso3c",
      "country.name"
    )
  ) |>
  dplyr::group_by(type) |>
  dplyr::slice_max(n = 3, order_by = emissions_co2_mt) |>
  dplyr::ungroup()

emissions_by_vessel_class_df <-
  all_port_and_trip_emissions |>
  dplyr::group_by(vessel_class, type) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt)) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    fishing = ifelse(
      vessel_class %in%
        c(
          "trawlers",
          "squid_jigger",
          "tuna_purse_seines",
          "set_longlines",
          "pole_and_line",
          "set_gillnets",
          "pots_and_traps",
          "dredge_fishing",
          "other_seines",
          "other_purse_seines",
          "trollers",
          "drifting_longlines",
          "driftnets",
          "fish_factory",
          "other_fishing"
        ),
      "Fishing",
      "Non-fishing"
    ) |>
      forcats::fct_rev()
  ) |>
  dplyr::mutate(
    vessel_class = stringr::str_replace_all(
      vessel_class,
      "\\.",
      ": "
    ) |>
      stringr::str_replace_all("_", " ") |>
      stringr::str_to_sentence() |>
      stringr::str_replace_all(
        "chemical oil",
        "chemical or oil"
      )
  )

top_type_by_fishing_non_fishing <- emissions_by_vessel_class_df |>
  dplyr::group_by(fishing, type) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt)) |>
  dplyr::ungroup() |>
  dplyr::group_by(fishing) |>
  dplyr::slice_max(n = 1, order_by = emissions_co2_mt) |>
  dplyr::ungroup()
```

# In-line manuscript statistics

## Abstract

-  We found that vessels emitted approximately `r signif((total_percent_change_by_fleet |> dplyr::filter(fleet == "AIS + S1", pollutant == "co2") |> dplyr::pull('2024'))/1e9,3)` billion metric tonnes of CO$_{2}$ in 2024, accounting for `r percent_transportation_emmissions_from_marine_2024` of global fossil fuel emissions from the transportation sector. 

- Current AIS-only approaches underestimate the magnitude of 2024 emissions by `r pollutant_ais_underestimation_overestimation_summary |>   dplyr::filter(pollutant == "co2") |>   dplyr::pull(ais_underestimation_2024)`. Meanwhile, because AIS coverage has generally been increasing over time, current AIS-only approaches overestimate the increasing emissions trend by `r pollutant_ais_underestimation_overestimation_summary |>   dplyr::filter(pollutant == "co2") |>   dplyr::pull(trend_overestimation_2017_2024)` (from 2017 to 2024).

## Introduction

- Climate change mitigation requires accurate accounting across sectors; while the transportation sector as a whole was responsible for `r fraction_co2_from_transportation` of all global CO$_{2}$ emissions in 2024, fundamental gaps remain in understanding the contribution by marine vessels. 

- For over  `r scales::comma(signif(n_unique_vessels$n_unique_vessels, 2))` AIS-broadcasting vessels, we quantify emissions for
each vessel using precise latitude/longitude/timestamp information for over `r scales::comma(signif(n_ais_messages$n_billion_ais_messages, 3))` billion individual AIS messages.

## Results

-  Using this data fusion approach, we find that total global CO$_{2}$ emissions at sea in 2024 was `r signif((total_percent_change_by_fleet |> dplyr::filter(fleet == "AIS + S1", pollutant == "co2") |> dplyr::pull('2024'))/1e9,3)` billion MT, with an increase of `r scales::percent(total_percent_change_by_fleet |> dplyr::filter(fleet == "AIS + S1", pollutant == "co2") |> dplyr::pull(percent_change))` from 2017 to 2024. If only considering emissions from vessels that broadcast AIS, in line with currently accepted state-of-the-art models, global CO$_{2}$ emissions in 2024 would be only `r signif((total_percent_change_by_fleet |> dplyr::filter(fleet == "AIS (broadcasting)", pollutant == "co2") |> dplyr::pull('2024'))/1e9,3)` billion MT, with a more pronounced increase from 2017 to 2024 of `r scales::percent(total_percent_change_by_fleet |> dplyr::filter(fleet == "AIS (broadcasting)", pollutant == "co2") |> dplyr::pull(percent_change))`.   

- Thus using only AIS data from broadcasting vessels underestimates current global CO$_{2}$ emissions by  `r pollutant_ais_underestimation_overestimation_summary |>   dplyr::filter(pollutant == "co2") |>   dplyr::pull(ais_underestimation_2024)`, and overestimates the increasing trend for 2017--2024 by `r pollutant_ais_underestimation_overestimation_summary |>   dplyr::filter(pollutant == "co2") |>   dplyr::pull(trend_overestimation_2017_2024)`

-   Across the entire `r analysis_start_year` to `r analysis_end_year`
    time series,
    `r dark_emissions_by_area |>   dplyr::filter(inside_footprint & s1_imaged) |>   dplyr::pull(percent_total)`
    of non-broadcasting emissions occurred inside the S1 footprint and
    in pixel-months that were imaged;
    `r dark_emissions_by_area |>   dplyr::filter(inside_footprint & !s1_imaged) |>   dplyr::pull(percent_total)`
    occurred inside the S1 footprint but in pixel-months that weren't
    imaged; and
    `r dark_emissions_by_area |>   dplyr::filter(!inside_footprint) |>   dplyr::pull(percent_total)`
    occurred outside the S1 footprint entirely.

-   In `r analysis_end_year`,
    `r dark_emissions_by_fishing |> dplyr::filter(fishing) |> dplyr::pull(percent_dark)`
    of CO$_{2}$ emissions from fishing vessels were from
    non-broadcasting vessels, and
    `r dark_emissions_by_fishing |> dplyr::filter(!fishing) |> dplyr::pull(percent_dark)`
    of CO$_{2}$ emissions from non-fishing vessels were from
    non-broadcasting vessels. For both types of vessels, this fraction
    of emissions from non-broadcasting vessels has been declining
    steadily since 2020.


- Most emissions for non-fishing vessels occur during `r top_type_by_fishing_non_fishing |> dplyr::filter(fishing == "Non-fishing") |> dplyr::pull(type) |> stringr::str_to_lower()`, while most emissions for fishing vessels occur during `r top_type_by_fishing_non_fishing |> dplyr::filter(fishing == "Fishing") |> dplyr::pull(type) |> stringr::str_to_lower()`.

- `r all_port_and_trip_emissions_top_3 |> dplyr::filter(type == "Domestic trips") |> dplyr::pull(country) |> glue::glue_collapse(", ", last = ", and ")` had the most CO$_{2}$ emissions during within-country domestic trips in 2024. 

- `r all_port_and_trip_emissions_top_3 |> dplyr::filter(type == "International trips") |> dplyr::pull(country) |> glue::glue_collapse(", ", last = ", and ")` had the most CO$_{2}$ emissions during international trips that started or ended in these countries in 2024 (where the emissions from each trip are partitioned equally to its departure and arrival country).

- `r all_port_and_trip_emissions_top_3 |> dplyr::filter(type == "Port stays") |> dplyr::pull(country) |> glue::glue_collapse(", ", last = ", and ")` had the most CO$_{2}$ emissions during port stays.

- Of the roughly `r fraction_vessels_emissions_by_registry_info_expanded |> dplyr::filter(registry_type == "Total") |> dplyr::pull(n_unique_vessels) |> signif(3) |> prettyNum(big.mark=",")` AIS-broadcasting vessels in our analysis, only `r fraction_vessels_emissions_by_registry_info_expanded |> dplyr::filter(registry_type == "imo") |> dplyr::pull(n_unique_vessels) |> signif(3) |> prettyNum(big.mark=",")` (`r fraction_vessels_emissions_by_registry_info_expanded |> dplyr::filter(registry_type == "imo") |> dplyr::pull(fraction_n_unique_vessels) |> scales::percent()`) are listed with the IMO, while `r fraction_vessels_emissions_by_registry_info_expanded |> dplyr::filter(registry_type == "other_registry") |> dplyr::pull(n_unique_vessels) |> signif(3) |> prettyNum(big.mark=",")` (`r fraction_vessels_emissions_by_registry_info_expanded |> dplyr::filter(registry_type == "other_registry") |> dplyr::pull(fraction_n_unique_vessels) |> scales::percent()`) are listed on some other public registry; this leaves over `r fraction_vessels_emissions_by_registry_info_expanded |> dplyr::filter(registry_type == "no_registry") |> dplyr::pull(n_unique_vessels) |> signif(3) |> prettyNum(big.mark=",")` vessels (`r fraction_vessels_emissions_by_registry_info_expanded |> dplyr::filter(registry_type == "no_registry") |> dplyr::pull(fraction_n_unique_vessels) |> scales::percent()`) that have no public registry information and must rely completely on modeled vessel characteristics (main engine power and design speed).

- Because many large vessels are already listed with the IMO, the majority of our total CO$_2$ emissions come from these vessels (`r fraction_vessels_emissions_by_registry_info_expanded |> dplyr::filter(registry_type == "imo") |> dplyr::pull(fraction_emissions_co2_mt) |> scales::percent()`); `r fraction_vessels_emissions_by_registry_info_expanded |> dplyr::filter(registry_type == "other_registry") |> dplyr::pull(fraction_emissions_co2_mt) |> scales::percent()` come from vessels listed on other registries; and `r fraction_vessels_emissions_by_registry_info_expanded |> dplyr::filter(registry_type == "no_registry") |> dplyr::pull(fraction_emissions_co2_mt) |> scales::percent()` come from vessels with no public registry information and which rely entirely on modeled vessel characteristics.

-   Using the latest data available from the EU's Emissions Database for
    Global Atmospheric Research (EDGAR), we estimate that in 2024 marine
    vessels accounted for
    `r percent_transportation_emmissions_from_marine_2024` of all
    transportation-related CO$_{2}$ emissions. Marine vessel CO$_{2}$
    emissions increased by
    `r percent_increase_from_2017_to_2024 |>   dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>   dplyr::pull(emissions_co2_mt_normalized_percent)`
    between `r analysis_start_year` and 2024,
    `r round(   (percent_increase_from_2017_to_2024 |>     dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>     dplyr::pull(emissions_co2_mt_normalized)) /     (percent_increase_from_2017_to_2024 |>       dplyr::filter(data_source == "Other transportation sectors (EDGAR)") |>       dplyr::pull(emissions_co2_mt_normalized)) )`
    times faster than the growth rate of other transportation
    emissions, and `r round(   (percent_increase_from_2017_to_2024 |>     dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>     dplyr::pull(emissions_co2_mt_normalized)) /     (percent_increase_from_2017_to_2024 |>       dplyr::filter(data_source == "All sectors (EDGAR)") |>       dplyr::pull(emissions_co2_mt_normalized)) )` faster than the growth rate of total global emissions across all sectors.


## Discussion

- For non-fishing vessels, the total social cost of CO$_{2}$, CH$_{4}$, and N$_{2}$O emissions in 2021 was `r annual_sc_fishing_non_fishing |> dplyr::filter(fishing == "Non-fishing",year=="Total") |> dplyr::pull(sc_billion_usd) |> round()` billion USD. This represents about `r scales::percent((annual_sc_fishing_non_fishing |> dplyr::filter(fishing == "Non-fishing",year=="Total") |> dplyr::pull(sc_billion_usd))/1e3/6.63)` of the total value of all goods transported by maritime trade (using 6.63 trillion USD as the 2021 value of maritime-traded goods).
    
- For fishing vessels, the total social cost of CO$_{2}$, CH$_{4}$, and N$_{2}$O emissions in 2022 was `r annual_sc_fishing_non_fishing |> dplyr::filter(fishing == "Fishing",year=="Total") |> dplyr::pull(sc_billion_usd) |> round()` billion USD. This represents about `r scales::percent((annual_sc_fishing_non_fishing |> dplyr::filter(fishing == "Fishing",year=="Total") |> dplyr::pull(sc_billion_usd))/159)` of the total value of all wild capture fisheries (using 159 billion USD as the 2022 value of wild capture fisheries).

## Methods

- The time between pings is typically small (the mean value is `r signif(ping_level_hours_distribution$mean_hours,3)` hours and the median value is`r signif(ping_level_hours_distribution$median_hours,3)` hours, or about `r round(ping_level_hours_distribution$mean_hours*60)` minutes and `r round(ping_level_hours_distribution$median_hours*60)` minutes respectively), while the maximum value for this across our dataset is `r round(ping_level_hours_distribution$max_hours)` hours (meaning this is the most amount of time over which we will estimate emissions for a single message).

- In all, we analyze AIS data for over `r scales::comma(signif(n_unique_vessels$n_unique_vessels, 2))` AIS-broadcasting vessels, across  `r scales::comma(signif(n_ais_messages$n_billion_ais_messages, 3))`  billion individual timestamped AIS messages which are assigned to`r signif((trip_co2_emissions_by_from_to_countries |> dplyr::filter(from_country_iso3!=to_country_iso3) |> dplyr::pull(n_unique_trips) |> sum())/1e6,3)` million international trips, `r signif((trip_co2_emissions_by_from_to_countries |> dplyr::filter(from_country_iso3==to_country_iso3) |> dplyr::pull(n_unique_trips) |> sum())/1e6,3)` million domestic trips, and `r signif((port_visit_co2_emissions_by_country |> dplyr::pull(n_unique_port_vists) |> sum())/1e6,3)` million port visits.

- As a result, we used a dataset containing `r signif(n_s1_detections$n_million_s1_detections,3)` million vessel detections from 2017 to 2024, of which `r signif(n_s1_detections$n_million_s1_detections_unmatched,3)` million detections were not matched to a broadcasting vessel ('unmatched'). These detections come from across more than `r prettyNum(floor(n_s1_detections$n_thousand_s1_scenes)*1000,big.mark = ",")` individual S1 scenes.

# Results

## Figure 1

```{r}
# Figure 1 A
total_annual_emissions_by_fleet_fig <- annual_emissions |>
  dplyr::filter(pollutant == "co2") |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS",
      TRUE ~ "S1"
    )
  ) |>
  ggplot(aes(x = year, y = emissions_mt, color = fleet)) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "B", scale = 1e-9),
    limits = c(0, 1.5e9),
    breaks = seq(0, 1.5e9, 0.25e9)
  ) +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(analysis_start_year, analysis_end_year)
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS + S1", "AIS", "S1")
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(Annual ~ CO[2] ~ emissions ~ (MT))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(0.75, 0.4),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  )
```

```{r}
# Figure 1 B
# Get analysis_start_year values for time-series normalization
annual_co2_emissions_starting <- annual_emissions |>
  dplyr::filter(pollutant == "co2", year == analysis_start_year) |>
  dplyr::mutate(emissions_co2_mt_starting = emissions_mt) |>
  dplyr::select(-c(year, emissions_mt))

normalized_annual_emissions_by_fleet_fig <- annual_emissions |>
  dplyr::filter(pollutant == "co2") |>
  dplyr::left_join(annual_co2_emissions_starting, by = "fleet") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_mt / emissions_co2_mt_starting - 1
  ) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS",
      TRUE ~ "S1"
    )
  ) |>
  ggplot(aes(x = year, y = emissions_co2_mt_normalized, color = fleet)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(analysis_start_year, analysis_end_year)
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS", "AIS + S1", "S1")
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of ~ CO[2],
      emissions ~ from ~ 2017 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(0.25, 0.85),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  )
```

```{r}
# Figure 1 C
pollutant_map_data_co2 <- total_spatial_emissions_by_pollutant_formatted |>
  dplyr::filter(pollutant %in% c("CO2")) |>
  dplyr::mutate(
    emissions_mt_per_km2_delta_abs = emissions_mt_per_km2_2024 -
      emissions_mt_per_km2_2017
  ) |>
  dplyr::select(
    lon_bin,
    lat_bin,
    emissions_mt_per_km2_delta_abs,
    emissions_mt_per_km2_2024
  ) |>
  tidyr::pivot_longer(
    cols = c(emissions_mt_per_km2_delta_abs, emissions_mt_per_km2_2024)
  )

co2_map <- pollutant_map_data_co2 |>
  dplyr::mutate(
    name = ifelse(
      name == "emissions_mt_per_km2_delta_abs",
      "Change in emissions from 2017 to 2024",
      "2024 emissions"
    )
  ) |>
  ggplot() +
  geom_sf(
    data = world_bbox_sf,
    fill = map_background_color
  ) +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      color = value,
      fill = value
    )
  ) +
  facet_wrap(name ~ ., ncol = 1) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scico::scale_fill_scico(
    name = expression(CO[2] ~ emissions ~ (MT / km^2)),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = map_background_color,
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  scico::scale_color_scico(
    name = expression(CO[2] ~ emissions ~ (MT / km^2)),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = map_background_color,
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme_void() +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black"
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(2, "cm")
  )
```

```{r}
#| label: fig-emissions-by-data-source-and-maps
#| fig-cap: (A) Annual marine vessel CO$_{2}$ emissions, with each line colored based on the data source. (B) Relative change in annual marine vessel CO$_{2}$ emissions from a 2017 baseline, with each line colored based on the data source. (C) Spatial map of CO$_{2}$ emissions in 2024 (top) and change in CO$_{2}$ emissions between 2017 and 2024 (bottom), aggregated across AIS-broadcasting vessels and non-broadcasting vessels, and shown at a 1x1 degree spatial resolution; emissions are area-normalized for each pixel (metric tonnes per square kilometer, shown on a pseudolog-10 scale).
#| fig-width: 9
#| fig-height: 6

# Figure 1 combined
cowplot::plot_grid(
  cowplot::plot_grid(
    total_annual_emissions_by_fleet_fig,
    normalized_annual_emissions_by_fleet_fig,
    ncol = 1,
    labels = c("A", "B")
  ),
  co2_map,
  ncol = 2,
  labels = c("", "C"),
  rel_widths = c(0.4, 0.6),
  align = "vh"
)
```

## Figure 2

```{r}
# Figure 2A
time_series_data_monthly_emissions_by_fleet <- monthly_aggregated_time_series |>
  dplyr::group_by(time, fishing) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(time = lubridate::ymd(time)) |>
  tidyr::pivot_longer(
    cols = c(emissions_co2_mt, emissions_co2_dark_mt),
    names_to = "fleet",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "emissions_co2_mt" ~ "AIS",
      fleet == "emissions_co2_dark_mt" ~ "S1"
    ),
    fleet = factor(fleet, levels = c("AIS", "S1")),
    fishing = dplyr::case_when(
      fishing == TRUE ~ "Fishing",
      fishing == FALSE ~ "Non-fishing"
    ),
    fishing = factor(fishing, levels = c("Non-fishing", "Fishing"))
  )

make_pollutant_by_fleet_time_series_figure <- function(
  pollutant_time_series_data,
  y_axis_label
) {
  ggplot(
    pollutant_time_series_data,
    aes(x = time, y = emissions_mt, fill = fleet)
  ) +
    geom_area(position = 'stack', alpha = 0.9) +
    geom_line(
      data = subset(pollutant_time_series_data, fleet == "S1"),
      aes(x = time, y = emissions_mt),
      color = "white"
    ) +
    scale_y_continuous(
      labels = scales::comma,
      limits = c(0, NA)
    ) +
    scale_fill_manual(
      values = all_data_source_color_palette,
      breaks = c("AIS", "S1")
    ) +
    scale_x_continuous(
      breaks = seq(
        lubridate::ymd(glue::glue("{analysis_start_year}-01-01")),
        lubridate::ymd(glue::glue("{analysis_end_year+1}-01-01")),
        by = "1 year"
      ),
      labels = seq(analysis_start_year, analysis_end_year + 1)
    ) +
    facet_wrap(~fishing, nrow = 1, scales = "free_y") +
    theme_plot() +
    labs(x = "", y = y_axis_label) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "inside",
      legend.position.inside = c(0.60, 0.90),
      legend.background = element_rect(fill = "white", color = "gray"),
      legend.title = element_blank(),
      axis.title.y = element_text(size = 9),
      axis.text.x = element_text(angle = 25, hjust = 0.7),
      legend.text = element_text(size = 8),
      legend.key.size = unit(0.5, "cm"),
      strip.text.x = element_text(size = 12)
    )
}

time_series_by_fleet_fig_co2 <- make_pollutant_by_fleet_time_series_figure(
  time_series_data_monthly_emissions_by_fleet,
  y_axis_label = expression(Monthly ~ CO[2] ~ emissions ~ (MT))
) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  )
```

```{r}
# Figure 2B
spatial_co2_emissions_ais_dark_by_fleet <- annual_spatial_co2_emissions_ais_dark_by_fleet |>
  dplyr::group_by(lon_bin, lat_bin, fishing) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt, na.rm = TRUE)
  ) |>
  dplyr::ungroup() |>
  tidyr::pivot_longer(
    cols = c(emissions_co2_mt, emissions_co2_dark_mt),
    names_to = "fleet",
    values_to = "emissions_co2_mt"
  ) |>
  dplyr::mutate(
    emissions_mt_per_km2 = emissions_co2_mt /
      lat_to_area(lat_bin, pixel_size_analysis)
  )

spatial_emissions_by_fleet_total_plot_pseudolog <- spatial_co2_emissions_ais_dark_by_fleet |>
  dplyr::mutate(
    fleet = factor(
      fleet,
      levels = c("emissions_co2_mt", "emissions_co2_dark_mt"),
      labels = c("AIS", "S1")
    )
  ) |>
  dplyr::mutate(
    fishing = ifelse(fishing, "Fishing", "Non-fishing") |>
      forcats::fct_rev()
  ) |>
  dplyr::mutate(
    emissions_mt_per_km2 = dplyr::na_if(emissions_mt_per_km2, 0)
  ) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey92") +
  ggspatial::geom_spatial_tile(aes(
    x = lon_bin + 0.5,
    y = lat_bin + 0.5,
    fill = emissions_mt_per_km2,
    color = emissions_mt_per_km2
  )) +
  theme_void() +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scico::scale_color_scico(
    name = expression(
      Cumulative ~ CO[2] ~ emissions ~ (MT / km^2) ~ from ~ 2017 ~ to ~ 2024
    ),
    palette = "vik",
    begin = 0.5,
    end = 1,
    trans = ggallin::pseudolog10_trans,
    breaks = c(10, 100, 1000, 10000),
    na.value = map_background_color,
    labels = scales::comma,
    guide = "none"
  ) +
  scico::scale_fill_scico(
    name = expression(
      Cumulative ~ CO[2] ~ emissions ~ (MT / km^2) ~ from ~ 2017 ~ to ~ 2024
    ),
    palette = "vik",
    begin = 0.5,
    end = 1,
    trans = ggallin::pseudolog10_trans,
    breaks = c(10, 100, 1000, 10000),
    na.value = map_background_color,
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme(
    legend.position = "bottom",
    legend.title.align = 0.5,
    strip.text.x = element_text(
      margin = margin(t = 10, b = 10),
      size = 10
    ),
    strip.text.y = element_text(
      angle = 0,
      hjust = 0,
      margin = margin(l = 10, r = 10),
      size = 10
    ),
    strip.background = element_rect(fill = "transparent", color = NA),
    panel.spacing = unit(1, "lines")
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "white",
      ticks.colour = "white",
      barwidth = unit(10, "cm"),
      barheight = unit(0.5, "cm")
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  facet_grid(fleet ~ fishing) +
  coord_sf(crs = map_projection)
```

```{r}
#| label: fig-spatial-temporal-richness-by-fleet-total-pseudolog
#| fig-cap: (A) Time series of monthly global CO$_2$ emissions, shown for both AIS-broadcasting and S1-detected non-broadcasting vessels, and for both fishing and non-fishing vessels. (B) Spatial distribution of cumulative 2017-2024 CO$_{2}$ emissions from both AIS-broadcasting and S1-detected non-broadcasting vessels, for both fishing and non-fishing vessels, and shown at a 1x1 degree spatial resolution; emissions are area-normalized for each pixel (metric tonnes per square kilometer, shown on a pseudolog-10 scale).
#| fig-width: 8
#| fig-height: 7

# Figure 2 combined
plot_constrained <- cowplot::plot_grid(
  NULL,
  time_series_by_fleet_fig_co2 +
    theme(plot.margin = margin(b = -15, l = 0)),
  NULL,
  ncol = 3,
  rel_widths = c(0.02, 0.88, 0.10)
)

cowplot::plot_grid(
  plot_constrained,
  spatial_emissions_by_fleet_total_plot_pseudolog,
  ncol = 1,
  rel_heights = c(0.35, 0.65),
  labels = c("A", "B")
)
```

## Figure 3

```{r}
#| label: fig-ais-data-richness
#| fig-cap: "Richness of AIS-broadcasting vessel emissions data, showing: (A) CO$_{2}$ emissions by vessel type for AIS-broadcasting vessels during different activity types (international trips, port stays, and domestic trips), aggregated across 2017-2024. (B) CO$_{2}$ emissions by country for AIS-broadcasting vessels during  during different activity types (international trips, port stays, and domestic trips), aggregated across 2017-2024. For both panels, emissions for each international trip are partitioned equally to its departure and arrival country). Countries with no values are shown in black."
#| fig-width: 9
#| fig-height: 8

emissions_by_vessel_class_barplot <- emissions_by_vessel_class_df |>
  dplyr::mutate(
    type = stringr::str_replace_all(type, " ", "\n") |>
      forcats::fct_relevel("International\ntrips", "Port\nstays")
  ) |>
  ggplot(aes(
    x = emissions_co2_mt,
    y = reorder(vessel_class, emissions_co2_mt),
    fill = type
  )) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
  ggforce::facet_col(~fishing, space = "free", scales = "free") +
  theme_plot() +
  labs(
    x = expression(2017 ~ through ~ 2024 ~ CO[2] ~ emissions ~ (MT)),
    y = "",
    fill = ""
  ) +
  scale_x_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  ) +
  paletteer::scale_fill_paletteer_d(palette = "ltc::trio4") +
  theme(legend.position = "bottom", legend.direction = "horizontal")

all_port_and_trip_emissions_without_vessel_class <- all_port_and_trip_emissions |>
  dplyr::group_by(country_iso3, type) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt)) |>
  dplyr::ungroup()


all_countries_for_plot <- land |>
  sf::st_drop_geometry() |>
  dplyr::select(country_iso3) |>
  tidyr::crossing(
    all_port_and_trip_emissions_without_vessel_class |>
      dplyr::distinct(type)
  )

port_trip_emissions_fig <- ggplot() +
  geom_sf(
    data = world_bbox_sf,
    fill = map_background_color
  ) +
  geom_sf(
    data = all_countries_for_plot |>
      dplyr::left_join(land, by = "country_iso3") |>
      dplyr::left_join(
        all_port_and_trip_emissions_without_vessel_class,
        by = c("country_iso3", "type")
      ) |>
      sf::st_as_sf(),
    aes(fill = emissions_co2_mt),
    color = "white",
    linewidth = 0.01
  ) +
  facet_wrap(type ~ ., ncol = 1) +
  coord_sf(crs = map_projection) +
  theme_void() +
  scale_fill_gradientn(
    # Get the "Blues" palette with 9 colors, but keep only the darker 5 (indices 4-9)
    colors = rev(RColorBrewer::brewer.pal(9, "Blues")[5:9]),
    labels = scales::unit_format(unit = "M", scale = 1e-6)
  ) +
  labs(
    x = "",
    fill = expression(2017 ~ through ~ 2024 ~ CO[2] ~ emissions ~ (MT)),
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "white",
      ticks.colour = "white"
    )
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(1.75, "cm")
  )

cowplot::plot_grid(
  emissions_by_vessel_class_barplot,
  port_trip_emissions_fig,
  ncol = 2,
  labels = c("A", "B"),
  rel_widths = c(0.52, 0.48)
)
```

## Figure 4

```{r}
#| label: fig-emissions-marine-ocean-other
#| fig-cap: (A) Trend in CO$_{2}$ emissions by ocean (normalized to 2017 baseline), use our fused AIS+S1 marine vessel data. (B) Comparison in trend of CO$_{2}$ emissions data (normalized to 2017 baseline) between marine transportation emissions from GFW, other transportation emissions from EDGAR, and all sector emissions from EDGAR. On the right-hand side of the figure, 2024 absolute values (in millions of metric tonnes) are shown as text and color-coded for each line.
#| fig-width: 7.5
#| fig-height: 7
# Calculate normalized annual changes in CO2 emissions by major ocean basin

total_spatial_co2_emissions_by_ocean <- total_spatial_co2_emissions_by_ocean |>
  dplyr::mutate(
    emissions_co2_mt_total = emissions_co2_dark_mt + emissions_co2_mt
  )

# Get analysis_start_year values for time-series normalization
annual_emissions_by_ocean_starting <- total_spatial_co2_emissions_by_ocean |>
  dplyr::filter(year == analysis_start_year) |>
  dplyr::mutate(emissions_co2_mt_total_starting = emissions_co2_mt_total) |>
  dplyr::select(ocean, emissions_co2_mt_total_starting)

annual_emissions_by_ocean_normalized <- total_spatial_co2_emissions_by_ocean |>
  dplyr::inner_join(annual_emissions_by_ocean_starting, by = "ocean") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_co2_mt_total /
      emissions_co2_mt_total_starting -
      1
  ) |>
  dplyr::mutate(
    ocean = paste0(stringr::str_to_sentence(ocean), " Ocean")
  )
ocean_data_ending <- annual_emissions_by_ocean_normalized |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    emissions_co2_mt_ending = paste0(
      prettyNum(signif(emissions_co2_mt_total / 1e6, 3), big.mark = ","),
      " MMT"
    )
  ) |>
  dplyr::select(
    ocean,
    emissions_co2_mt_ending,
    emissions_co2_mt_normalized
  )

annual_ocean_emissions_normalized_fig <- annual_emissions_by_ocean_normalized |>
  ggplot() +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(
    aes(x = year, y = emissions_co2_mt_normalized, color = ocean),
    linewidth = 1.1
  ) +
  scale_x_continuous(
    breaks = seq(2017, 2024),
    limits = c(2017, 2025)
  ) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of,
      CO[2] ~ emissions ~ from ~ 2017 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(0.21, 0.6),
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", color = "gray")
  ) +
  scale_color_brewer(
    palette = "Dark2",
    breaks = c(
      "Southern Ocean",
      "Arctic Ocean",
      "Pacific Ocean",
      "Atlantic Ocean",
      "Indian Ocean"
    )
  ) +
  ggrepel::geom_text_repel(
    data = ocean_data_ending,
    aes(
      x = 2025,
      y = emissions_co2_mt_normalized,
      label = emissions_co2_mt_ending,
      color = ocean
    ),
    direction = "y",
    show.legend = FALSE,
    nudge_x = 0,
    xlim = c(2024.1, 2024.6),
    min.segment.length = Inf,
    fontface = "bold"
  )

# Summarize total emissions by data source for 2023, for putting on figure

all_edgar_gfw_data_ending <- all_edgar_gfw_data_normalized |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    emissions_co2_mt_ending = paste0(
      prettyNum(signif(emissions_co2_mt / 1e9, 3), big.mark = ","),
      " BMT"
    )
  ) |>
  dplyr::select(
    data_source,
    emissions_co2_mt_ending,
    emissions_co2_mt_normalized
  )

emissions_comparison_figure_normalized <- all_edgar_gfw_data_normalized |>
  dplyr::filter(
    data_source %in%
      c(
        "Other transportation sectors (EDGAR)",
        "Marine vessels (AIS + S1)",
        "All sectors (EDGAR)"
      )
  ) |>
  ggplot() +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(
    aes(x = year, y = emissions_co2_mt_normalized, color = data_source),
    linewidth = 1.1
  ) +
  scale_x_continuous(
    breaks = seq(2017, 2024),
    limits = c(2017, 2025)
  ) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of,
      CO[2] ~ emissions ~ from ~ 2017 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(0.3, 0.9),
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", color = "gray")
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c(
      "Marine vessels (AIS + S1)",
      "All sectors (EDGAR)",
      "Other transportation sectors (EDGAR)"
    )
  ) +
  ggrepel::geom_text_repel(
    data = all_edgar_gfw_data_ending,
    aes(
      x = 2025,
      y = emissions_co2_mt_normalized,
      label = emissions_co2_mt_ending,
      color = data_source
    ),
    show.legend = FALSE,
    nudge_x = 0,
    xlim = c(2024.1, 2024.6),
    min.segment.length = Inf,
    fontface = "bold"
  )
cowplot::plot_grid(
  annual_ocean_emissions_normalized_fig,
  emissions_comparison_figure_normalized,
  ncol = 1,
  labels = c("A", "B")
)
```

# Methods

## Figure 6

```{r}
#| label: fig-registered-data-performance
#| fig-cap: Comparison of daily vessel-level CO$_{2}$ emissions estimated using our model with registered emissions from a proprietary vessel database (based on main engine fuel consumption over 24 hours). Estimates are shown using main engine power and design speed from IMO (left), our machine learning random forest models (center), and the same registered dataset (right). The solid red line shows a linear regression fit, while the dashed line represents the 1:1 line. Performance values for traditional R$^{2}$ are shown for each set of vessel characteristics. Over 15,000 vessels are represented in the figure, and blue shading indicates point density. All data are rule-of-three compliant, meaning each cell (20x20 mt) displayed contains at least three vessels.
#| fig-width: 8
#| fig-height: 4
model_levels <- c(
  "IMO report lookup table",
  "ML-derived engine power\nand design speed",
  "Registered data"
)

vessel_long <- registered_validation_data |>
  dplyr::select(
    co2_emissions_tonnes_registered,
    "IMO report lookup table" = co2_emissions_tonnes_estimate_imo,
    "ML-derived engine power\nand design speed" = co2_emissions_tonnes_estimate_rf,
    "Registered data" = co2_emissions_tonnes_estimate_registered
  ) |>
  tidyr::pivot_longer(
    cols = -co2_emissions_tonnes_registered,
    names_to = "Model",
    values_to = "Estimate"
  ) |>
  dplyr::mutate(Model = factor(Model, levels = model_levels)) # Order levels


r2_labels <- vessel_long |>
  dplyr::group_by(Model) |>
  yardstick::rsq_trad(
    truth = co2_emissions_tonnes_registered,
    estimate = Estimate
  ) |>
  dplyr::mutate(label = paste0("Traditional R² = ", round(.estimate, 2)))

ggplot(
  vessel_long,
  aes(
    x = co2_emissions_tonnes_registered,
    y = Estimate
  )
) +
  geom_bin2d(
    bins = 30,
    aes(
      fill = after_stat(count),
      alpha = after_stat(ifelse(count < 3, 0, 1)) # Full transparency for counts < 3
    )
  ) +
  scale_fill_gradientn(
    colours = colorRampPalette(RColorBrewer::brewer.pal(9, "Blues"))(256),
    trans = "log10",
    name = "Count",
    guide = guide_colorbar(
      ticks.linewidth = 0.3,
      frame.colour = "black"
    )
  ) +
  geom_smooth(
    method = "lm",
    linewidth = 0.5,
    color = "red",
    linetype = "solid",
    se = FALSE
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_text(
    data = r2_labels |>
      dplyr::filter(!Model %in% c("RF design speed", "RF engine power")),
    aes(label = label),
    x = 450,
    y = 50,
    inherit.aes = FALSE,
    size = 3,
    fontface = "bold"
  ) +
  coord_fixed(ratio = 1, clip = "on") +
  facet_wrap(Model ~ ., nrow = 1, scales = "fixed") +
  labs(
    x = "Registered CO2 emissions (MT)",
    y = "Estimated CO2 emissions using our model (MT)"
  ) +
  scale_x_continuous(limits = c(0, 600)) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    axis.text.y = element_text(size = 7),
    axis.title = element_text(size = 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.ticks.length = unit(0.25, "cm"),
    panel.spacing = unit(1.5, "lines"),
    strip.placement = "outside",
    legend.position = "none", # Remove legend
  )
```

## Table 1

```{r}
# Generate MRV performance metrics results

## Emissions validation ----
validation_performance_metrics <- yardstick::metric_set(
  yardstick::rsq,
  yardstick::rsq_trad
)

# EU dataset
eu_validation_data_updated <- mrv_data_validation |>
  dplyr::filter(total_time_spent_at_sea_hours != 0) |>
  dplyr::mutate(
    co2_trip_emissions_eu = co2_emissions_from_all_voyages_between_ports_under_a_ms_jurisdiction_m_tonnes +
      co2_emissions_from_all_voyages_which_departed_from_ports_under_a_ms_jurisdiction_m_tonnes +
      co2_emissions_from_all_voyages_to_ports_under_a_ms_jurisdiction_m_tonnes
  ) |>
  dplyr::rename(
    total_time_spent_at_sea_hours_eu = total_time_spent_at_sea_hours
  )
# Emission results by trip
eu_validation_trip_updated <- trip_emissions_for_mrv_validation |>
  dplyr::filter(total_time_spent_at_sea_hours != 0) |>
  dplyr::rename(
    co2_trip_emissions_gfw = total_emissions_co2_mt,
    total_time_spent_at_sea_hours_gfw = total_time_spent_at_sea_hours
  )

find_duplicates <- function(data, key_columns) {
  duplicate_indices <- duplicated(data[key_columns]) |
    duplicated(data[key_columns], fromLast = TRUE)
  duplicates <- data[duplicate_indices, ]
  return(duplicates)
}

aggregate_duplicates <- function(data, group_by_cols, sum_cols, concat_cols) {
  aggregated_data <- data |>
    dplyr::group_by(dplyr::across(dplyr::all_of(group_by_cols))) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(sum_cols), sum, .names = "{.col}"),
      dplyr::across(
        dplyr::all_of(concat_cols),
        ~ paste(unique(.x), collapse = "-"),
        .names = "{.col}"
      ),
      .groups = 'drop'
    )

  return(aggregated_data)
}

group_by_columns <- c("imo_number", "year")
sum_columns <- c(
  "total_time_spent_at_sea_hours_gfw",
  "total_distance_nm",
  "co2_trip_emissions_gfw"
)

concat_columns <- c("ssvid")

eu_validation_trip_updated <- aggregate_duplicates(
  eu_validation_trip_updated,
  group_by_columns,
  sum_columns,
  concat_columns
)

# Data selection and filtering
merged_df <- merge(
  eu_validation_data_updated,
  eu_validation_trip_updated,
  by.x = c("imo_number", "reporting_period"),
  by.y = c("imo_number", "year")
)

merged_df$diff <- abs(
  merged_df$total_time_spent_at_sea_hours_eu -
    merged_df$total_time_spent_at_sea_hours_gfw
)
merged_df$max <- pmax(
  merged_df$total_time_spent_at_sea_hours_eu,
  merged_df$total_time_spent_at_sea_hours_gfw
)

# Function to filter data and calculate performance metrics based on a threshold
calculate_metrics <- function(threshold, merged_df) {
  filtered_hours_df <- subset(merged_df, diff / max < threshold)
  filtered_df <- filtered_hours_df |>
    dplyr::select(co2_trip_emissions_eu, co2_trip_emissions_gfw)

  # Calculate performance metrics
  comparison_results <- filtered_df |>
    tidyr::pivot_longer(
      -co2_trip_emissions_eu,
      names_to = "model",
      values_to = "estimate"
    ) |>
    dplyr::group_by(model) |>
    validation_performance_metrics(
      truth = co2_trip_emissions_eu,
      estimate = estimate
    ) |>
    dplyr::select(-.estimator) |>
    tidyr::pivot_wider(names_from = .metric, values_from = .estimate)

  # Return results with added threshold information
  comparison_results <- comparison_results |>
    dplyr::mutate(threshold = threshold, n_observations = nrow(filtered_df))

  return(comparison_results)
}

# Range of thresholds to test
thresholds <- seq(0.05, 0.4, 0.05)
# Apply calculate_metrics function over range of thresholds
performance_results <- purrr::map_df(
  thresholds,
  ~ calculate_metrics(.x, merged_df)
)

best_performance <- performance_results |>
  dplyr::slice_max(order_by = rsq_trad, n = 1)

# Select data with threshold of 0.15 the one showing best performance
filtered_hours_df <- subset(merged_df, diff / max < best_performance$threshold)
filtered_df <- filtered_hours_df |>
  dplyr::select(co2_trip_emissions_eu, co2_trip_emissions_gfw)

```

```{r}
#| results = 'asis'
mrv_performance_results_tbl <- performance_results |>
  dplyr::mutate(
    rsq = signif(rsq, 2),
    rsq_trad = signif(rsq_trad, 2),
    n_observations = prettyNum(n_observations, big.mark = ",")
  ) |>
  dplyr::select(
    Tolerance = threshold,
    `Number of vessels` = n_observations,
    "R$^{2}$" = rsq,
    "Traditional R$^{2}$" = rsq_trad
  ) |>
  knitr::kable(
    format = "latex",
    label = "tab-performance-tolerance-table",
    caption = "Performance summary across different thresholds for the hours at sea tolerance between datasets. Results are shown using the R$^{2}$ (measuring linear association between observed and predicted values) and traditional R$^{2}$ (proportion of variance explained compared to the mean-only baseline).",
    escape = FALSE
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

mrv_performance_results_tbl

mrv_performance_results_tbl |>
  writeLines(
    here::here("tables/mrv_performance_results.tex")
  )
```

## Figure 7

```{r}
#| label: fig-mrv-performance
#| fig-cap: Comparison of annual vessel-level CO$_{2}$ emissions estimated using our model with published emissions from the EU MRV database. The sample includes 2,276 observations from yearly aggregated vessel trips where hours at sea differ by no more than 15\% between our data and the MRV dataset. The solid red line shows the linear regression fit, and the dashed line represents the 1:1 line.
#| fig-width: 8
#| fig-height: 5

ggplot(
  filtered_df,
  aes(
    x = co2_trip_emissions_eu,
    y = co2_trip_emissions_gfw
  )
) +
  geom_point(
    size = 1.5,
    alpha = 0.5,
    color = "#104E8B"
  ) +
  geom_smooth(
    method = "lm",
    linewidth = 0.5,
    color = "red",
    linetype = "solid",
    se = FALSE
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  annotate(
    geom = "text",
    label = paste0(
      "Traditional R² = ",
      format(best_performance$rsq_trad, digits = 2)
    ),
    x = 130000,
    y = 10000,
    size = 3,
    fontface = "bold"
  ) +
  coord_fixed(ratio = 1, clip = "on") +
  scale_x_continuous(limits = c(0, max(filtered_df)), labels = scales::comma) +
  scale_y_continuous(limits = c(0, max(filtered_df)), labels = scales::comma) +
  labs(
    x = "Published emissions in the EU MRV database (MT)",
    y = "Estimated CO2 emissions using our model (MT)"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    axis.text.y = element_text(size = 7),
    axis.title = element_text(size = 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.ticks.length = unit(0.25, "cm"),
    panel.spacing = unit(1.5, "lines"),
    strip.placement = "outside"
  )
```

## Figure 8

```{r}
#| label: fig-map-fraction-months-imaged
#| fig-cap: Map showing the percentage of months from 2017 through 2024 that were imaged by S1 for each 1x1 degree pixel.
#| fig-width: 7.5
#| fig-height: 5
number_s1_imaged_months_by_pixel |>
  dplyr::mutate(
    fraction_months_imaged = number_months_imaged / number_months
  ) |>
  ggplot() +
  geom_sf(
    data = world_bbox_sf,
    fill = map_background_color
  ) +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      color = fraction_months_imaged,
      fill = fraction_months_imaged
    )
  ) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scale_fill_viridis_c(
    name = "Percentage of months imaged by S1 between 2017 and 2024",
    labels = scales::percent,
    option = "magma"
  ) +
  scale_color_viridis_c(
    name = "Percentage of months imaged by S1 between 2017 and 2024",
    labels = scales::percent,
    option = "magma"
  ) +
  labs(x = "", y = "") +
  theme_void() +
  theme_plot() +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "white",
      barwidth = unit(15, "cm"),
      barheight = unit(1, "cm")
    ),
    color = "none"
  )
```

## Figure 9

```{r}
#| label: fig-length-bin-distributions
#| fig-cap: Distributions showing the number of unique AIS vessels and number of S1 detections for each length size bin, disaggregated by fishing and non-fishing vessels.
#| fig-width: 7.5
#| fig-height: 6
length_size_bin_distributions |>
  dplyr::mutate(
    fishing = ifelse(fishing, "Fishing", "Non-fishing") |> forcats::fct_rev()
  ) |>
  dplyr::mutate(
    type = ifelse(
      type == "ais_vessels",
      "Number AIS vessels",
      "Number S1 detections"
    )
  ) |>
  dplyr::mutate(
    length_size_bin = glue::glue(
      "{length_size_bin}\n({length_bin_min}-{length_bin_max}m)"
    ) |>
      stringr::str_replace("-Infm", "+ m")
  ) |>
  ggplot(aes(x = length_size_bin, y = n)) +
  facet_wrap(type ~ fishing, scales = "free") +
  geom_bar(stat = "identity") +
  theme_plot() +
  labs(x = "Length size bin", y = "Count") +
  scale_y_continuous(labels = scales::comma)
```

## Figure 10

```{r}
#| label: fig-ais-length-power-relationship
#| fig-cap: Relationship between main engine power (killowatts) and length (meters) for all AIS-broadcasting vessels, disaggregated by non-fishing and fishing vessels.
#| fig-height: 7
#| fig-width: 6
vessel_size_info |>
  dplyr::mutate(
    fishing = ifelse(
      fishing,
      "Fishing",
      "Non-fishing"
    ) |>
      forcats::fct_rev()
  ) |>
  ggplot(aes(x = length_m, y = main_engine_power_kw)) +
  geom_point(size = 0.1, alpha = 0.1) +
  theme_plot() +
  facet_wrap(fishing ~ ., ncol = 1, scales = "free") +
  labs(x = "Length (m)", y = "Main engine power (kW)") +
  scale_y_continuous(labels = scales::comma)
```

## Table 2

```{r}
#| results = 'asis'
data_sources_table <- data_sources |>
  dplyr::select(-`Model feature shorthand`) |>
  dplyr::mutate(
    dplyr::across(
      dplyr::everything(),
      ~ stringr::str_replace_all(., "²", "$^2$")
    )
  ) |>
  knitr::kable(
    format = "latex",
    escape = FALSE,
    booktabs = TRUE,
    caption = "Data sources for all included model features, including which random forest model uses the features (M1: detections classification; M2: detections regression; and/or M3: emissions regression), the units, feature type, whether it varies spatially, whether it varies temporally, and the source name and reference citation.",
    label = "data_sources"
  ) |>
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 4
  )

data_sources_table

writeLines(data_sources_table, here::here("tables/data_sources.tex"))
```


## Figure 11

```{r}
#| label: fig-s1-coverage-time-series
#| fig-cap: Monthly statistics of global S1 coverage including total number of S1 scenes, total imaged area across all S1 scenes, total number of vessel detections, total number of vessel detections that could not be matched to an AIS-broadcasting vessels, and fraction of the total number of vessel detections that could not be matched to an AIS-broadcasting vessels.
#| fig-width: 7.5
#| fig-height: 8
s1_time_series |>
  dplyr::mutate(
    month = lubridate::ymd(month),
    fraction_detections_unmatched = n_s1_detections_unmatched /
      (n_s1_detections)
  ) |>
  tidyr::pivot_longer(-month) |>
  dplyr::mutate(
    name = dplyr::case_when(
      name ==
        "fraction_detections_unmatched" ~ "Fraction vessel detections unmatched",
      name == "n_s1_detections" ~ "Number vessel detections",
      name ==
        "n_s1_detections_unmatched" ~ "Number vessel detections unmatched",
      name == "s1_scene_area_km2" ~ "Total imaged area of S1 scenes (km^2)",
      name == "s1_scenes" ~ "Number of S1 scenes"
    ) |>
      forcats::fct_relevel(c(
        "Number of S1 scenes",
        "Total imaged area of S1 scenes (km^2)",
        "Number vessel detections",
        "Number vessel detections unmatched"
      ))
  ) |>
  ggplot(aes(x = month, y = value)) +
  facet_wrap(name ~ ., ncol = 1, scales = "free") +
  geom_line() +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma) +
  theme_plot() +
  scale_x_continuous(
    breaks = seq(
      lubridate::ymd(glue::glue("{analysis_start_year}-01-01")),
      lubridate::ymd(glue::glue("{analysis_end_year+1}-01-01")),
      by = "1 year"
    ),
    labels = seq(analysis_start_year, analysis_end_year + 1)
  ) +
  theme_plot() +
  labs(x = "", y = "")
```

## Figure 12

```{r}
#| label: fig-offshore-outside-footprint-training-testing-map
#| fig-cap: Maps of pixels that were used in the training and testing datasets for the simulated outside S1 footprint performance tests for each model. Pixels $<=$ 100km from the shore were used for training; pixels $>$ 100km from shore were used for testing.
#| fig-width: 7.5
#| fig-height: 7.5
pixels_for_offshore_training_testing_split |>
  dplyr::mutate(type = ifelse(offshore_pixel, "Testing", "Training")) |>
  dplyr::mutate(
    model = stringr::str_replace_all(model, "_", "; ") |>
      stringr::str_to_sentence()
  ) |>
  ggplot() +
  geom_sf(
    data = world_bbox_sf,
    fill = map_background_color
  ) +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      color = type,
      fill = type
    )
  ) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  paletteer::scale_color_paletteer_d(
    name = "Dataset",
    palette = "nbapalettes::pacers_classic",
    direction = -1
  ) +
  paletteer::scale_fill_paletteer_d(
    name = "Dataset",
    palette = "nbapalettes::pacers_classic",
    direction = -1
  ) +
  labs(x = "", y = "") +
  theme_void() +
  theme_plot() +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  facet_wrap(model ~ ., ncol = 1) +
  guides(fill = guide_legend(reverse = TRUE)) +
  guides(color = guide_legend(reverse = TRUE))
```

## Figure 13

```{r}
#| label: fig-pr-curves
#| fig-cap: Precision-recall curves of the detection presence/absence classification model, shown for both the inside S1 footprint test and simulated outside S1 footprint test.
#| fig-height: 3

performance_detections_cls_roc_pr_curves |>
  tidyr::pivot_wider(names_from = "metric", values_from = "estimate") |>
  dplyr::mutate(
    test = stringr::str_replace_all(test, "_", " ") |> stringr::str_to_title()
  ) |>
  ggplot(aes(x = recall, y = precision)) +
  geom_line() +
  coord_equal() +
  facet_wrap(test ~ ., ) +
  labs(
    x = "Recall",
    y = "Precision",
  ) +
  theme_plot()
```

## Figure 14

```{r}
#| label: fig-roc-curves
#| fig-cap: Receiver Operating Characteristic (ROC) curves of the detection classification model, shown for both the inside S1 footprint test and simulated outside S1 footprint test.
#| fig-height: 3

performance_detections_cls_roc_pr_curves |>
  tidyr::pivot_wider(names_from = "metric", values_from = "estimate") |>
  dplyr::mutate(
    test = stringr::str_replace_all(test, "_", " ") |> stringr::str_to_title()
  ) |>
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_line() +
  coord_equal() +
  facet_wrap(test ~ ., ) +
  labs(
    x = "False positive rate (1 - specificity)",
    y = "True positive rate (sensitivity)",
  ) +
  theme_plot() +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1))
```

## Figure 15

```{r}
#| label: fig-conf-mat
#| fig-cap: Confusion matrices of the detection classification model, shown for both the inside S1 footprint test and simulated outside S1 footprint test.
#| fig-height: 3

performance_detections_cls_conf_mat |>
  dplyr::mutate(
    test = stringr::str_replace_all(test, "_", " ") |> stringr::str_to_title()
  ) |>
  dplyr::mutate(prediction = forcats::fct_rev(as.factor(prediction))) |>
  ggplot(aes(x = truth, y = prediction)) +
  geom_tile(aes(fill = count), color = "black") +
  geom_label(aes(label = prettyNum(count, big.mark = ",")), vjust = 1) +
  scale_fill_gradient(low = "gray95", high = "gray10", guide = "none") +
  facet_wrap(test ~ .) +
  coord_equal() +
  labs(x = "Truth", y = "Prediction") +
  theme_minimal()
```

## Table 3

```{r}
#| results = 'asis'
all_performance_metrics_table <- all_performance_metrics |>
  dplyr::mutate(
    model = stringr::str_replace_all(model, "_", " ") |>
      stringr::str_to_sentence(),
    test = stringr::str_replace_all(test, "_", " ") |>
      stringr::str_to_sentence()
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant),
    pollutant = tidyr::replace_na(pollutant, "-")
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::arrange(pollutant) |>
  dplyr::select(
    Model = model,
    Test = test,
    Pollutant = pollutant,
    Metric = metric,
    Estimate = estimate
  ) |>
  dplyr::mutate(
    Metric = dplyr::case_when(
      Metric == "accuracy" ~ "Accuracy",
      Metric == "f_meas" ~ "F1 score",
      Metric == "pr_auc" ~ "Precision-recall AUC",
      Metric == "roc_auc" ~ "ROC AUC",
      Metric == "precision" ~ "Precision",
      Metric == "sensitivity" ~ "Sensitivity",
      Metric == "specificity" ~ "Specificity",
      Metric ==
        "rsq_trad_backtransformed_residuals" ~ "Traditional R² (backtransformed)",
      Metric == "rsq_trad" ~ "Traditional R²",
      Metric == "mae" ~ "MAE",
      Metric == "aggregate_prediction_percent_error" ~ "Total percent error",
      TRUE ~ Metric
    ) |>
      forcats::fct_relevel("Traditional R²")
  ) |>
  dplyr::mutate(Model = forcats::fct_relevel(Model, "Emissions regression")) |>
  dplyr::arrange(Model, Metric) |>
  tidyr::pivot_wider(values_from = Estimate, names_from = Test) |>
  knitr::kable(
    format = "latex",
    label = "all_performance_metrics_table",
    caption = "Summary of dark fleet model performance metrics for the: 1) emissions regression model; 2) detections classification mode; and 3) detections regression model. Performance estimates for each metric are shown for both the inside S1 footprint and outside S1 footprint tests. All metrics are calculated at the unit of observation except for 'Total percent error', which is the percentage error in the total sum of all predictions relative to the total sum of all truth values.",
    digits = 2
  ) |>
  kableExtra::kable_styling(font_size = 7, latex_options = "hold_position")

all_performance_metrics_table

writeLines(
  all_performance_metrics_table,
  here::here("tables/all_performance_metrics_table.tex")
)
```

## Table 4

```{r}
#| results: 'asis'

lm_other_gases_tidy_fit_stats_tbl <- lm_other_gases_tidy_fit_stats |>
  dplyr::mutate(term = stringr::str_remove_all(term, "_")) |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::select(-c(model, run_version_dark_fleet_rf)) |>
  dplyr::arrange(pollutant) |>
  dplyr::mutate(std_error = scales::scientific(std_error)) |>
  dplyr::rename(
    Pollutant = pollutant,
    Term = term,
    `Coefficient estimate` = estimate,
    `Standard error` = std_error,
    `T-statistic` = statistic,
    `p.value` = p_value
  ) |>
  knitr::kable(
    format = "latex",
    caption = "Linear regression fit statistics for the final models trained to estimate emissions of non-CO$_2$ gases from CO$_2$ emissions. The model for each gas has 2 fit terms.",
    label = "lm_other_gases_tidy_fit_stats",
    digits = 3
  ) |>
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 7
  )

lm_other_gases_tidy_fit_stats_tbl

writeLines(
  lm_other_gases_tidy_fit_stats_tbl,
  here::here("tables/lm_other_gases_tidy_fit_stats.tex")
)
```


## Figure 16

```{r}
#| label: fig-feature-importance
#| fig-cap: Feature importance plots for the 1) detections (presence/absence) classification model; 2) detections regression model; and 3) CO$_2$ emissions regression model.
#| fig-height: 9
#| fig-width: 7.5

all_varimp_data |>
  dplyr::left_join(
    data_sources,
    by = c("model_feature" = "Model feature shorthand")
  ) |>
  dplyr::rename(model_feature_long = `Model feature`) |>
  dplyr::mutate(
    model_feature_long = dplyr::case_when(
      stringr::str_detect(
        model_feature,
        "length_size_bin_"
      ) ~ glue::glue(
        "S1 detections per km2 (length bin {stringr::str_remove(model_feature, 'length_size_bin_')})"
      ),
      TRUE ~ model_feature_long
    )
  ) |>
  dplyr::filter(!is.na(model_feature_long)) |>
  dplyr::mutate(
    model = stringr::str_replace_all(model, "_", " ") |>
      stringr::str_to_sentence()
  ) |>
  dplyr::mutate(
    model = ifelse(
      model == "Emissions regression",
      "CO2 emissions regression",
      model
    ) |>
      forcats::fct_relevel(
        "Detections classification",
        "Detections regression",
        "CO2 emissions regression"
      )
  ) |>
  ggplot(aes(
    x = importance,
    y = tidytext::reorder_within(model_feature_long, importance, model)
  )) +
  geom_bar(stat = "identity") +
  labs(
    x = "Feature importance score (Gini impurtity corrected)",
    y = "",
  ) +
  theme_plot() +
  tidytext::scale_y_reordered() +
  facet_wrap(model ~ ., ncol = 1, scales = "free")
```

## Figure 17

```{r}
#| label: fig-spatial-coverage-footprint
#| fig-cap: Spatial coverage of the non-broadcasting emissions model. Areas are categorized according to whether they fall inside or outside the S1 footprint and whether they have been imaged by S1 at any point in time. A given pixel may be imaged in some months but not in others. Pixels outside the footprint have never been imaged. The percentage of total CO$_{2}$ emissions from non-broadcasting vessels attributed to each category is shown.
#| fig-width: 5
#| fig-height: 7

total_spatial_co2_emissions_dark_by_footprint |>
  dplyr::left_join(
    dark_emissions_by_area |>
      dplyr::select(inside_footprint, s1_imaged, percent_total),
    by = c("inside_footprint", "s1_imaged")
  ) |>
  dplyr::mutate(
    footprint = dplyr::case_when(
      inside_footprint == FALSE ~ "Outside S1 footprint",
      inside_footprint == TRUE &
        s1_imaged == TRUE ~ "Inside S1 footprint imaged",
      inside_footprint == TRUE &
        s1_imaged == FALSE ~ "Inside S1 footprint not imaged",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::group_by(footprint) |>
  dplyr::mutate(
    facet_label = glue::glue(
      "{footprint} ({percent_total} of non-broadcasting emissions)"
    )
  ) |>
  dplyr::ungroup() |>
  ggplot() +
  geom_sf(
    data = world_bbox_sf,
    fill = map_background_color,
    color = NA
  ) +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      fill = footprint
    ),
    color = NA
  ) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scale_fill_manual(
    name = "",
    values = c(
      "Outside S1 footprint" = "#d90101ff",
      "Inside S1 footprint imaged" = "#d90101ff",
      "Inside S1 footprint not imaged" = "#d90101ff"
    )
  ) +
  theme_void() +
  facet_wrap(~facet_label, ncol = 1) +
  theme(legend.position = "none")
```

# Supplementary materials: Supplementary figures

## Figure S1

```{r}
#| label: fig-pollutant-maps-qlog10
#| fig-cap: Spatial maps 2024 emissions for each GHG and pollutant, aggregated across AIS-broadcasting vessels and non-broadcasting vessels, and shown at a 1x1 degree spatial resolution. Emissions are area-normalized for each pixel (metric tonnes per square kilometer), and shown using a log-10 scale.
#| fig-width: 8
#| fig-height: 8

pollutant_map_data_all <- total_spatial_emissions_by_pollutant_formatted |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

pollutant_map_list <- purrr::map(
  unique(pollutant_map_data_all$pollutant),
  function(pollutant_tmp) {
    pollutant_tmp <- as.character(pollutant_tmp)

    pollutant_map_data <- pollutant_map_data_all |>
      dplyr::filter(pollutant == pollutant_tmp)

    scale_limits <- pollutant_map_data |>
      dplyr::summarize(
        min = quantile(emissions_mt_per_km2_2024, 0.01, na.rm = TRUE),
        max = max(emissions_mt_per_km2_2024, na.rm = TRUE)
      ) |>
      as.numeric()

    pollutant_map_data |>
      ggplot() +
      geom_sf(
        data = world_bbox_sf,
        fill = map_background_color
      ) +
      ggspatial::geom_spatial_tile(
        aes(
          x = lon_bin + 0.5,
          y = lat_bin + 0.5,
          color = emissions_mt_per_km2_2024,
          fill = emissions_mt_per_km2_2024
        )
      ) +
      geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
      scico::scale_fill_scico(
        name = expression(Emissions ~ (MT / km^2)),
        palette = "vik",
        begin = 0.5,
        end = 1,
        limits = scale_limits,
        trans = "log10",
        na.value = map_background_color,
        labels = scales::scientific
      ) +
      scico::scale_color_scico(
        name = expression(Emissions ~ (MT / km^2)),
        palette = "vik",
        begin = 0.5,
        end = 1,
        limits = scale_limits,
        trans = "log10",
        na.value = map_background_color,
        labels = scales::scientific
      ) +
      labs(x = "", y = "") +
      theme_void() +
      guides(
        fill = guide_colorbar(
          title.position = "top",
          title.hjust = 0.5,
          frame.colour = "white",
          ticks.colour = "white"
        )
      ) +
      scale_x_continuous(expand = c(0, 1)) +
      scale_y_continuous(expand = c(0, 1)) +
      theme(
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.key.width = unit(1, "cm")
      ) +
      facet_wrap(pollutant ~ ., labeller = label_parsed)
  }
)

cowplot::plot_grid(plotlist = pollutant_map_list, ncol = 3)
```

## Figure S2

```{r}
#| label: fig-pollutant-time-series
#| fig-cap: Time series of monthly global emissions for each GHG and pollutant, aggregated across AIS-broadcasting vessels and non-broadcasting vessels for which we only have Sentinel-1 data.
#| fig-width: 8
#| fig-height: 8

pollutant_time_series_data_all <- total_monthly_emissions_by_pollutant |>
  dplyr::mutate(month = lubridate::ymd(month)) |>
  tidyr::pivot_longer(
    -month,
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_total_mt") |>
      stringr::str_to_upper() |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::filter(
    lubridate::year(month) <= analysis_end_year,
    lubridate::year(month) >= analysis_start_year
  ) |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

make_pollutant_time_series_figure <- function(
  pollutant_time_series_data,
  y_axis_label
) {
  {
    ggplot(
      data = pollutant_time_series_data,
      aes(x = month, y = emissions_mt)
    ) +
      geom_line(linewidth = 1.1) +
      scale_y_continuous(labels = scales::comma) +
      theme_plot() +
      labs(x = "", y = y_axis_label) +
      theme(
        panel.grid.minor.x = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0)
      ) +
      scale_x_continuous(
        breaks = seq(
          lubridate::ymd(glue::glue("{analysis_start_year}-01-01")),
          lubridate::ymd(glue::glue("{analysis_end_year+1}-01-01")),
          by = "1 year"
        ),
        labels = seq(analysis_start_year, analysis_end_year + 1)
      ) +
      scale_y_continuous(limits = c(0, NA), labels = scales::comma)
  }
}

make_pollutant_time_series_figure(
  pollutant_time_series_data_all,
  y_axis_label = "Monthly emissions (MT)"
) +
  facet_wrap(
    pollutant ~ .,
    ncol = 3,
    scales = "free_y",
    labeller = label_parsed
  )
```

## Figure S3

```{r}
#| label: fig-annual-emissions-by-ocean-and-data-source
#| fig-cap: Annual total CO2 emissions by ocean, disaggregated by data source (AIS-broadcasting vessels, and non-broadcasting vessels detected by S1).
#| fig-width: 7.5
#| fig-height: 7.5
ocean_time_series_fig_data <- total_spatial_co2_emissions_by_ocean |>
  dplyr::select(ocean, year, emissions_co2_mt, emissions_co2_dark_mt) |>
  dplyr::rename(AIS = emissions_co2_mt, S1 = emissions_co2_dark_mt) |>
  tidyr::pivot_longer(
    -c(ocean, year),
    values_to = "emissions_mt",
    names_to = "fleet"
  ) |>
  dplyr::mutate(ocean = stringr::str_to_title(ocean))

ggplot(
  ocean_time_series_fig_data,
  aes(x = year, y = emissions_mt, fill = fleet)
) +
  geom_area(position = 'stack', alpha = 0.9) +
  geom_line(
    data = subset(ocean_time_series_fig_data, fleet == "S1"),
    aes(x = year, y = emissions_mt),
    color = "white"
  ) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  ) +
  scale_fill_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS", "S1")
  ) +
  scale_x_continuous(
    breaks = seq(
      analysis_start_year,
      analysis_end_year
    )
  ) +
  facet_wrap(~ocean, ncol = 1, scales = "free_y") +
  theme_plot() +
  labs(
    x = "",
    y = expression(Annual ~ CO[2] ~ emissions ~ (MT)),
    fill = "Data source"
  )
```

# Supplementary materials: Supplementary tables

## Table S1

```{r}
#| results = 'asis'
total_percent_change_by_fleet_tbl <- total_percent_change_by_fleet |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "AIS (broadcasting)" ~ "AIS",
      fleet == "S1 (non-broadcasting)" ~ "S1",
      TRUE ~ "AIS + S1"
    )
  ) |>
  dplyr::mutate(across(
    c(`2017`, `2024`),
    ~ signif(., 3) |>
      prettyNum(big.mark = ",")
  )) |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::mutate(
    dplyr::across(
      c(percent_total_2024, percent_change),
      ~ scales::percent(., accuracy = 1)
    )
  ) |>
  dplyr::arrange(pollutant) |>
  dplyr::rename(
    `Source` = fleet,
    Pollutant = pollutant,
    `2017 emissions, MT` = `2017`,
    `2024 emissions, MT` = `2024`,
    `Percent 2024 total` = percent_total_2024,
    `2017 to 2024 change` = percent_change
  ) |>
  knitr::kable(
    format = "latex",
    label = "total_percent_change_by_fleet",
    caption = "Summary of 2017 and 2024 emissions, data source and pollutant."
  ) |>
  kableExtra::kable_styling(font_size = 6, latex_options = "hold_position")

total_percent_change_by_fleet_tbl

writeLines(
  total_percent_change_by_fleet_tbl,
  here::here("tables/total_percent_change_by_fleet.tex")
)
```

## Table S2

```{r}
#| results = 'asis'
pollutant_ais_underestimation_overestimation_summary_tbl <- pollutant_ais_underestimation_overestimation_summary |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::arrange(pollutant) |>
  dplyr::rename(
    Pollutant = pollutant,
    "Underestimation of 2024 emissions" = ais_underestimation_2024,
    "Overestimation of 2017-2024 emissions trend" = trend_overestimation_2017_2024
  ) |>
  knitr::kable(
    format = "latex",
    label = "pollutant_ais_underestimation_overestimation_summary",
    caption = "Underestimation of 2024 emissions using only AIS data, and overestimation of 2017 to 2024 emissions trend using only AIS data, by pollutant."
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

pollutant_ais_underestimation_overestimation_summary_tbl

writeLines(
  pollutant_ais_underestimation_overestimation_summary_tbl,
  here::here(
    "tables/pollutant_ais_underestimation_overestimation_summary.tex"
  )
)
```

## Table S3

```{r}
#| results = 'asis'

# Make and output a table summarizing 2024 CO2 emissions by ocean, and change in emissions from 2017 to 2024 by ocean
# Figure out for each ocean the fraction of observations that were imaged by s1
percent_pixels_imaged_ocean_all_time <- annual_emissions_by_ocean_normalized |>
  dplyr::group_by(ocean) |>
  dplyr::summarise(
    fraction_n_imaged = (sum(number_observations_imaged) /
      sum(number_observations)) |>
      scales::percent(accuracy = 0.01)
  ) |>
  dplyr::ungroup()
emissions_by_ocean_summary_tbl <- annual_emissions_by_ocean_normalized |>
  dplyr::filter(year == 2024) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    percent_annual_total = (emissions_co2_mt_total /
      sum(emissions_co2_mt_total)) |>
      scales::percent(accuracy = 0.01),
    percent_emissions_co2_mt_normalized = emissions_co2_mt_normalized |>
      scales::percent(accuracy = 0.01)
  ) |>
  dplyr::ungroup() |>
  dplyr::left_join(percent_pixels_imaged_ocean_all_time, by = "ocean") |>
  dplyr::mutate(
    `Percent of total CO2 emissions by non-broadcasting` = (emissions_co2_dark_mt /
      emissions_co2_mt_total) |>
      scales::percent(accuracy = 0.01)
  ) |>
  dplyr::mutate(
    ocean = stringr::str_to_sentence(ocean) |>
      stringr::str_remove_all(" ocean")
  ) |>
  dplyr::mutate(
    emissions_co2_mmt = prettyNum(
      signif(emissions_co2_mt / 1e6, 3),
      big.mark = ","
    )
  ) |>
  dplyr::select(
    Ocean = ocean,
    `2024 total CO2 emissions (million MT)` = emissions_co2_mmt,
    `Percent of total CO2 2024 emissions` = percent_annual_total,
    `Percent of total CO2 emissions by non-broadcasting`,
    `Percent change in total CO2 emissions (2017 to 2024)` = percent_emissions_co2_mt_normalized,
    `Percent observations imaged by S1 (across 2017-2024)` = fraction_n_imaged,
  ) |>
  tidyr::pivot_longer(cols = -Ocean) |>
  tidyr::pivot_wider(names_from = Ocean, values_from = value) |>
  dplyr::rename(`Summary statistic` = name) |>
  knitr::kable(
    format = "latex",
    label = "emissions_by_ocean_summary_tbl",
    caption = "For each ocean, summary of: 2024 total (AIS+S1) CO$_2$ emissions (metric tonnes); percentage of global total (AIS+S1) 2024 CO$_2$ emissions from across all oceans; percent of 2024 total (AIS+S1) CO$_2$ emissions by non-broadcasting vessels; percent change in total (AIS+S1) CO$_2$ emissions from 2017 to 2024; and the percentage of all observations in the dataset (pixel-months) from 2017-2024 that were imaged by S1.",
    digits = 3
  ) |>
  kableExtra::kable_styling(latex_options = "hold_position")

emissions_by_ocean_summary_tbl

writeLines(
  emissions_by_ocean_summary_tbl,
  here::here("tables/emissions_by_ocean_summary_tbl.tex")
)
```

## Table S4

```{r}
#| results = 'asis'

# Make and output a table summarizing 2024 CO2 emissions by data source, and change in emissions from 2017 to 2024 by data source
emissions_by_data_source_summary_tbl <- all_edgar_gfw_data_normalized |>
  dplyr::filter(
    data_source %in%
      c(
        "Other transportation sectors (EDGAR)",
        "Marine vessels (AIS + S1)",
        "All sectors (EDGAR)"
      )
  ) |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    emissions_co2_bmt = prettyNum(
      signif(emissions_co2_mt / 1e9, 3),
      big.mark = ","
    ),
    percent_emissions_co2_mt_normalized = emissions_co2_mt_normalized |>
      scales::percent(accuracy = 0.01),
  ) |>
  dplyr::ungroup() |>
  dplyr::select(
    `Data source` = data_source,
    `2024 emissions, billion MT` = emissions_co2_bmt,
    `Change in emissions, 2017 to 2024` = percent_emissions_co2_mt_normalized
  ) |>
  knitr::kable(
    format = "latex",
    label = "emissions_by_data_source_summary_tbl",
    caption = "For each data source, summary of: total 2024 CO$_2$ emissions (metric tonnes); and percent change in CO$_2$ emissions from 2017 to 2024.",
    digits = 3
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

emissions_by_data_source_summary_tbl

writeLines(
  emissions_by_data_source_summary_tbl,
  here::here("tables/emissions_by_data_source_summary_tbl.tex")
)
```

## Table S5

```{r}
#| results: 'asis'

annual_sc_fishing_non_fishing_tbl <- annual_sc_fishing_non_fishing |>
  dplyr::select(
    Year = year,
    `Vessel type` = fishing,
    GHG = pollutant,
    `Emissions (MT)` = emissions_mt,
    `SC (USD/MT)` = sc_usd_per_mt,
    `SC (billion USD)` = sc_billion_usd
  ) |>
  knitr::kable(
    format = "latex",
    label = "social_cost_summary_tbl",
    caption = "For both non-fishing vessels and fishing vessels, we summarize: the total emissions (metric tonnes) of each GHG gas; the social cost associated with each GHG (USD/metric tonne); the social cost of each (USD); the total social cost summed across all three gasses (USD). We use 2021 for non-fishing vessels since it is the year for which we have the total value of all traded goods at sea. We use 2022 for fishing vessels since it is the year for which we have the total value of wild capture fisheries.",
    digits = 3
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

annual_sc_fishing_non_fishing_tbl

writeLines(
  annual_sc_fishing_non_fishing_tbl,
  here::here("tables/annual_sc_fishing_non_fishing_tbl.tex")
)
```

# Supplementary materials: Supplementary figures: Comparing our emission estimates to other inventories

## Figure S4

```{r}
#| label: fig-inventory-comparison
#| fig-cap: Comparison of annual marine vessel CO$_{2}$ emissions inventory estimates.
#| fig-width: 9
#| fig-height: 6

all_inventory_data |>
  ggplot(aes(x = year, y = emissions_co2_mt, color = data_source)) +
  geom_line(linewidth = 1) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "B", scale = 1e-9),
    limits = c(0, 1.5e9),
    breaks = seq(0, 1.5e9, 0.25e9)
  ) +
  theme_plot() +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(analysis_start_year, analysis_end_year)
  ) +
  labs(
    x = "",
    y = expression(Annual ~ CO[2] ~ emissions ~ (MT))
  ) +
  theme(panel.grid.minor = element_blank()) +
  scale_color_manual(
    name = "Emissions inventory",
    values = all_data_source_color_palette,
    breaks = c("GFW (AIS + S1)", "IMO", "GFW (AIS)", "OECD", "EDGAR")
  )
```

## Table S6
```{r}
#| results: 'asis'
all_inventory_data_kable <- all_inventory_data |>
  dplyr::mutate(
    emissions_co2_mt = as.character(signif(emissions_co2_mt, 3) / 1e9)
  ) |>
  tidyr::pivot_wider(
    names_from = "data_source",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::rename(Year = year) |>
  dplyr::relocate(Year, `GFW (AIS + S1)`, `GFW (AIS)`) |>
  dplyr::mutate(across(everything(), ~ tidyr::replace_na(., "-"))) |>
  knitr::kable(
    format = "latex",
    label = "inventory_comparison",
    caption = "Comparison of annual marine vessel CO$_{2}$ emissions inventory estimates (billions of metric tonnes)."
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

all_inventory_data_kable

all_inventory_data_kable |>
  writeLines(
    here::here("tables/inventory_comparison.tex")
  )
```

## Figure S5

```{r}
#| label: fig-annual-emissions-by-ais-receiver-type
#| fig-cap: Annual total CO$_{2}$ emissions (metric tonnes) by AIS-broadcasting vessels, AIS pings (i.e., individual vessel timestamped location messages), and number of unique AIS-broadcasting vessels. Trends are disaggregated by AIS receiver type (satellite, terrestrial, and dynamic), and also shown as a total aggregated across receiver types. A dashed vertical line is shown for 2022, the first year in which dynamic AIS was widely adopted.
# | #| fig-width: 4
#| fig-height: 7
annual_global_emissions_by_receiver_type <- annual_global_emissions_by_receiver_type |>
  # Don't include unknown receiver type
  dplyr::filter(receiver_type != "unknown") |>
  # Fill in missing values with 0s
  tidyr::complete(
    receiver_type,
    tidyr::nesting(year),
    fill = list(emissions_co2_mt = 0, n_distinct_ssvid = 0, number_pings = 0)
  ) |>
  dplyr::mutate(receiver_type = stringr::str_to_sentence(receiver_type)) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025)

annual_global_emissions_by_receiver_type |>
  dplyr::rename(
    `CO[2]~emissions~(MT)` = emissions_co2_mt,
    `Number~unique~vessels` = n_distinct_ssvid,
    `Number~AIS~messages` = number_pings,
  ) |>
  tidyr::pivot_longer(cols = -c(year, receiver_type)) |>
  ggplot(aes(x = year, y = value, color = receiver_type)) +
  facet_wrap(name ~ ., scales = "free_y", ncol = 1, labeller = label_parsed) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  ) +
  scale_x_continuous(
    breaks = unique(annual_global_emissions_by_receiver_type$year)
  ) +
  scale_color_brewer(
    type = "qual",
    palette = "Set2",
    guide = guide_legend(reverse = TRUE)
  ) +
  theme_plot() +
  labs(x = "", y = "", color = "AIS receiver type") +
  theme(panel.grid.minor.x = element_blank()) +
  geom_vline(xintercept = 2022, linetype = 2)
```

## Figure S6

```{r}
#| label: fig-annual-emissions-by-ais-receiver-type-top-flags
#| fig-cap: Annual total CO$_{2}$ emissions (metric tonnes) by AIS-broadcasting vessels, disaggregated by AIS receiver type (satellite, terrestrial, and dynamic) and total aggregated across receiver types, for the top 10 flags with the most emissions in 2024. A dashed vertical line is shown for 2022, the first year in which dynamic AIS was widely adopted.
#| fig-width: 7.5
#| fig-height: 6
annual_global_emissions_by_receiver_type_and_flag <- annual_global_emissions_by_receiver_type_and_flag |>
  # Add column name for country name based on flag

  dplyr::mutate(
    flag = countrycode::countrycode(
      flag,
      origin = "iso3c",
      destination = "country.name"
    )
  ) |>
  # Don't include unknown receiver type
  dplyr::filter(receiver_type != "unknown") |>
  dplyr::mutate(
    receiver_type = stringr::str_to_sentence(receiver_type) |>
      forcats::fct_relevel("Satellite", "Terrestrial", "Dynamic")
  ) |>
  # Fill in missing values with 0s
  tidyr::complete(
    receiver_type,
    tidyr::nesting(year, flag),
    fill = list(emissions_co2_mt = 0, n_distinct_ssvid = 0)
  ) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025)

top_n_flags_2024 <- annual_global_emissions_by_receiver_type_and_flag |>
  dplyr::filter(year == 2024, receiver_type == "Total") |>
  dplyr::slice_max(n = 10, order_by = emissions_co2_mt) |>
  dplyr::select(flag)

annual_global_emissions_by_receiver_type_and_flag |>
  dplyr::inner_join(top_n_flags_2024, by = "flag") |>
  dplyr::mutate(flag = forcats::fct_relevel(flag, top_n_flags_2024$flag)) |>
  ggplot(aes(x = year, y = emissions_co2_mt, color = flag)) +
  facet_grid(receiver_type ~ ., scales = "free_y") +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  scale_x_continuous(
    breaks = unique(annual_global_emissions_by_receiver_type_and_flag$year)
  ) +
  theme_plot() +
  labs(x = "", y = expression(CO[2] ~ emissions ~ (MT)), color = "Flag") +
  theme(panel.grid.minor.x = element_blank()) +
  geom_vline(xintercept = 2022, linetype = 2) +
  paletteer::scale_color_paletteer_d(palette = "rcartocolor::Safe")

```

## Figure S7

```{r}
#| label: fig-co2-emissions-change-map-by-receiver-type
#| fig-cap: Spatial change in CO$_{2}$ emissions by AIS-broadcasting vessels between 2017 and 2024 (metric tonnes) by receiver type (satellite, terrestrial, and dynamic) and total across receiver types, at a 1x1 degree spatial resolution; emissions are area-normalized for each pixel (metric tonnes per square kilometer, shown on a pseudolog-10 scale).
#| fig-width: 7.5
#| fig-height: 8
# Annual spatial CO2 emissions vt receiver type  (satellite; dynamic; or terrestrial)
annual_spatial_emissions_by_receiver_type <- annual_spatial_emissions_by_receiver_type |>
  # Don't include unknown receiver type
  dplyr::filter(receiver_type != "unknown") |>
  dplyr::mutate(
    receiver_type = stringr::str_to_sentence(receiver_type) |>
      forcats::fct_relevel("Satellite", "Terrestrial", "Dynamic")
  ) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025) |>
  dplyr::mutate(
    emissions_mt_per_km2 = emissions_co2_mt /
      lat_to_area(lat_bin, pixel_size_analysis)
  ) |>
  dplyr::select(-emissions_co2_mt)

annual_spatial_emissions_by_receiver_type |>
  dplyr::filter(year %in% c(analysis_start_year, analysis_end_year)) |>
  tidyr::pivot_wider(names_from = year, values_from = emissions_mt_per_km2) |>
  dplyr::mutate(across(c(`2017`, `2024`), ~ tidyr::replace_na(., 0))) |>
  dplyr::mutate(emissions_mt_per_km2 = `2024` - `2017`) |>
  ggplot() +
  geom_sf(
    data = world_bbox_sf,
    fill = map_background_color
  ) +
  ggspatial::geom_spatial_tile(aes(
    x = lon_bin + 0.5,
    y = lat_bin + 0.5,
    fill = emissions_mt_per_km2,
    color = emissions_mt_per_km2
  )) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  facet_wrap(receiver_type ~ ., ncol = 1) +
  scico::scale_fill_scico(
    expression(atop(
      Change ~ of ~ CO[2] ~ emissions,
      from ~ 2017 ~ to ~ 2024 ~ (MT / km^2)
    )),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = map_background_color,
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  scico::scale_color_scico(
    expression(atop(
      Change ~ of ~ CO[2] ~ emissions,
      from ~ 2017 ~ to ~ 2024 ~ (MT / km^2)
    )),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = map_background_color,
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme_void() +
  theme(legend.position = "right", legend.direction = "vertical") +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black",
      barwidth = unit(1, "cm"),
      barheight = unit(5, "cm")
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  coord_sf(crs = map_projection)
```
