---
title: "Quantifying comprehensive marine vessel emissions using data fusion"
author: "emLab, UCSB and GFW"
date: last-modified
date-format: "[Last Updated on] MMMM DD, YYYY"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    fig-pos: H
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
cache: true
---

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE, cache = FALSE}
#| label: document-setup

# Set default chunk options
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  dev = "png",
  fig.path = "../figures/"
)

# Load librariers
library(ggplot2)
library(sf)

analysis_start_year <- 2017
analysis_end_year <- 2024
```

```{r}
#| label: load-data
#| cache: FALSE
targets::tar_load(c(
  "n_unique_vessels",
  "n_ais_messages",
  "n_s1_detections",
  "annual_edgar_emissions",
  "oecd_data",
  "all_performance_metrics",
  "monthly_aggregated_time_series",
  "annual_emissions_all_pollutants",
  "total_spatial_emissions_by_pollutant",
  "annual_ais_to_dark_activity_extrapolation",
  "total_monthly_emissions_by_pollutant",
  "annual_ais_co2_emissions_by_vessel_type",
  "port_visit_co2_emissions_by_country",
  "trip_co2_emissions_by_from_to_countries",
  "annual_global_emissions_by_receiver_type",
  "annual_global_emissions_by_receiver_type_and_flag",
  "annual_spatial_emissions_by_receiver_type",
  "vessel_size_info",
  "registered_validation_data",
  "mrv_data_validation",
  "trip_emissions_for_mrv_validation",
  "registered_validation_data",
  "all_varimp_data",
  "performance_detections_cls_roc_pr_curves",
  "performance_detections_cls_conf_mat",
  "annual_spatial_co2_emissions_ais_dark_by_fleet",
  "total_spatial_co2_emissions_dark_by_footprint",
  "data_sources",
  "pixels_for_offshore_training_testing_split",
  "total_spatial_co2_emissions_by_ocean",
  "number_s1_imaged_months_by_pixel",
  "lm_other_gases_tidy_fit_stats"
))
```

```{r}
#| label: setup-graphics

# For figures and tables, order pollutants first for GHG pollutants,
# then local pollutants. For figures, we use expressions to show subscripts
pollutant_factor_levels <- c(
  "CO[2]",
  "CH[4]",
  "N[2]*O",
  "CO",
  "NO[X]",
  "SO[X]",
  "PM[2.5]",
  "PM[10]",
  "VOCs"
)

# Set pollutant levels for tables
# For tables, we don't use expressions
pollutant_factor_levels_tables <- pollutant_factor_levels |>
  stringr::str_remove_all("\\[") |>
  stringr::str_remove_all("\\]") |>
  stringr::str_remove_all("\\*]*")

# Use Okabe-Ito colorblind-friendly colors
all_data_source_color_palette <- c(
  "Marine vessels (AIS + S1)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[5]],
  "GFW (AIS + S1)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[5]],
  "AIS + S1" = paletteer::paletteer_d("colorblindr::OkabeIto")[[5]],
  "Other transportation sectors (EDGAR)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[7]],
  "GFW (AIS)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[1]],
  "AIS" = paletteer::paletteer_d("colorblindr::OkabeIto")[[1]],
  "S1" = paletteer::paletteer_d("colorblindr::OkabeIto")[[
    6
  ]],
  "All sectors (EDGAR)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[8]],
  "EDGAR" = paletteer::paletteer_d("colorblindr::OkabeIto")[[8]],
  "OECD" = paletteer::paletteer_d("colorblindr::OkabeIto")[[6]],
  "IMO" = paletteer::paletteer_d("colorblindr::OkabeIto")[[3]],
  "Fishing" = paletteer::paletteer_d("colorblindr::OkabeIto")[[5]],
  "Non-fishing" = paletteer::paletteer_d("colorblindr::OkabeIto")[[2]]
)

# Set consistent theme for all plots
theme_plot <- function() {
  theme_minimal() %+replace%
    theme(
      axis.title.x = element_text(face = "bold", size = 12),
      axis.title.y = element_text(
        angle = 90,
        face = "bold",
        vjust = 3,
        size = 12
      ),
      strip.text.x = element_text(angle = 0, face = "bold", size = 14),
      strip.text.y = element_text(angle = 0, face = "bold", size = 14)
    )
}

# Set map projection for all map figures
map_projection <- "+proj=eqearth"

# Create bounding box of globe, for making black map backgrounds
lats <- c(90:-90, -90:90, 90)
longs <- c(rep(c(180, -180), each = 181), 180)
world_bbox_sf <-
  list(cbind(longs, lats)) %>%
  sf::st_polygon() %>%
  sf::st_sfc(
    # create sf geometry list column
    crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
  ) %>%
  sf::st_sf() %>%
  sf::st_transform(map_projection)

# ESRI World Countries Generalized Shapefile
# Downloaded from https://www.arcgis.com/home/item.html?id=fa510018bdd044b08fc64d2a16bc680a
land <- sf::st_read(
  here::here(
    "data/World_Countries_Generalized_Shapefile/World_Countries_Generalized.shp"
  )
) |>
  dplyr::mutate(
    country_iso3 = countrycode::countrycode(ISO, "iso2c", "iso3c")
  )


# Convert lat/lon pixel area to square km
# https://stackoverflow.com/a/1253545
# Need function to convert degrees to radians
# https://stackoverflow.com/q/32370485
deg2rad <- function(deg) {
  (deg * pi) / (180)
}

# Given a latitude, what is a pixel's area?
# https://stackoverflow.com/a/1253545
lat_to_area <- function(lat) {
  110.574 * (111.320 * cos(deg2rad(lat)))
}

```

```{r}
#| label: wrangle-edgar-data
annual_edgar_tranportation_emissions <- annual_edgar_emissions |>
  dplyr::select(
    -c(
      IPCC_annex,
      C_group_IM24_sh,
      Country_code_A3,
      Name,
      ipcc_code_2006_for_standard_report,
      Substance,
      fossil_bio
    )
  ) |>
  dplyr::rename(sector = ipcc_code_2006_for_standard_report_name) |>
  tidyr::pivot_longer(
    cols = -c(sector),
    names_to = "year",
    values_to = "emissions_co2_kt"
  ) |>
  dplyr::mutate(
    year = stringr::str_remove_all(year, "Y_") |>
      as.numeric(),
    emissions_co2_mt = emissions_co2_kt * 1e3
  ) |>
  dplyr::select(-emissions_co2_kt) |>
  dplyr::mutate(
    land_based_tranpsportation = ifelse(
      sector %in%
        c(
          "Civil Aviation",
          "Road Transportation no resuspension",
          "Other Transportation",
          "Railways"
        ),
      TRUE,
      FALSE
    )
  ) |>
  # Only need years that correspond to our dataset
  dplyr::filter(year >= analysis_start_year, year <= analysis_end_year)

# Fraction of global 2024 CO2 emissions that came from transportation sectors
fraction_co2_from_transportation <- annual_edgar_tranportation_emissions |>
  dplyr::mutate(
    transportation = ifelse(
      land_based_tranpsportation | sector == "Water-borne Navigation",
      TRUE,
      FALSE
    )
  ) |>
  dplyr::filter(year == 2024) |>
  dplyr::group_by(transportation) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(
    names_from = "transportation",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::mutate(fraction_from_transportation = `TRUE` / (`TRUE` + `FALSE`)) |>
  dplyr::pull(fraction_from_transportation) |>
  scales::percent()
```

```{r}
#| label: wrangle-gfw-time-series-emissions-data

annual_emissions_all_pollutants <- annual_emissions_all_pollutants |>
  tidyr::pivot_longer(
    -c(year, fishing),
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    fleet = ifelse(
      stringr::str_detect(pollutant, "dark"),
      "Non-broadcasting",
      "AIS-broadcasting"
    )
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_mt") |>
      stringr::str_remove_all("_dark")
  )

annual_emissions_all_pollutants_with_total <- annual_emissions_all_pollutants |>
  # Add total, that aggregates across broadcasting and non-broadcasting
  dplyr::bind_rows(
    annual_emissions_all_pollutants |>
      dplyr::group_by(year, fishing, pollutant) |>
      dplyr::summarize(emissions_mt = sum(emissions_mt)) |>
      dplyr::ungroup() |>
      dplyr::mutate(fleet = "Total")
  )

# Aggregate emissions across fishing/non-fishing
annual_emissions <- annual_emissions_all_pollutants_with_total |>
  dplyr::group_by(year, fleet, pollutant) |>
  dplyr::summarise(emissions_mt = sum(emissions_mt)) |>
  dplyr::ungroup()

# Calculate the total percent change in emissions by fleet
total_percent_change_by_fleet <- annual_emissions |>
  dplyr::filter(
    year %in% c(analysis_start_year, analysis_end_year),
    fleet != "Total"
  ) |>
  tidyr::pivot_wider(names_from = year, values_from = emissions_mt) |>
  dplyr::arrange(pollutant) |>
  dplyr::group_by(pollutant) |>
  dplyr::mutate(
    percent_total_2024 = ifelse(
      fleet == "AIS + S1",
      "-",
      (`2024` / sum(`2024`))
    )
  ) |>
  dplyr::ungroup() %>%
  split(.[, c("pollutant")]) %>%
  purrr::map_df(janitor::adorn_totals, fill = NA) |>
  tidyr::fill(pollutant, .direction = "down") |>
  dplyr::mutate(
    percent_change = ((`2024` - `2017`) /
      `2017`)
  ) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS (broadcasting)",
      TRUE ~ "S1 (non-broadcasting)"
    )
  )
```

```{r}
# For each pollutant, estimate how much emissions are underestimated by just using AIS
# As compared to using AIS + S1
ais_scale_underestimation_2024 <- total_percent_change_by_fleet |>
  dplyr::filter(fleet != "S1 (non-broadcasting)") |>
  dplyr::select(fleet, pollutant, `2024`) |>
  tidyr::pivot_wider(names_from = "fleet", values_from = `2024`) |>
  dplyr::mutate(
    ais_underestimation_2024 = (1 - `AIS (broadcasting)` / `AIS + S1`) |>
      scales::percent(accuracy = 0.01)
  ) |>
  dplyr::rename(
    emissions_mt_2024_ais = `AIS (broadcasting)`,
    emissions_mt_2024_ais_s1 = `AIS + S1`
  )

# For each pollutant, estimate how much the 2017->2024 trend in emissions is underestimated by just using AIS
# As compared to using AIS + S1
ais_trend_overestimation_2017_2024 <- total_percent_change_by_fleet |>
  dplyr::filter(fleet != "S1 (non-broadcasting)") |>
  dplyr::select(fleet, pollutant, percent_change) |>
  tidyr::pivot_wider(names_from = "fleet", values_from = percent_change) |>
  dplyr::mutate(
    trend_overestimation_2017_2024 = (`AIS (broadcasting)` /
      `AIS + S1` -
      1) |>
      scales::percent(accuracy = 0.01)
  ) |>
  dplyr::rename(
    trend_2017_2024_ais = `AIS (broadcasting)`,
    trend_2017_2024_ais_s1 = `AIS + S1`
  )

# Put all this information together
pollutant_ais_underestimation_overestimation_summary <- ais_scale_underestimation_2024 |>
  dplyr::left_join(ais_trend_overestimation_2017_2024, by = "pollutant") |>
  dplyr::select(
    pollutant,
    ais_underestimation_2024,
    trend_overestimation_2017_2024
  )
```

```{r}
#| label: wrangel-edgar-data
# Calculate annual other transportation emissions from EDGAR
land_based_transportation_emissions_edgar <- annual_edgar_tranportation_emissions |>
  dplyr::filter(land_based_tranpsportation) |>
  dplyr::group_by(year) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::mutate(data_source = "Other transportation sectors (EDGAR)")


global_emissions_edgar <- annual_edgar_tranportation_emissions |>
  dplyr::group_by(year) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::mutate(data_source = "All sectors (EDGAR)")

all_edgar_gfw_data <- land_based_transportation_emissions_edgar |>
  dplyr::bind_rows(global_emissions_edgar) |>
  dplyr::bind_rows(
    annual_emissions |>
      dplyr::filter(pollutant == "co2", fleet == "Total") |>
      dplyr::rename(emissions_co2_mt = emissions_mt) |>
      dplyr::mutate(data_source = "Marine vessels (AIS + S1)") |>
      dplyr::select(-c(fleet, pollutant))
  )

# Get analysis_start_year values for time-series normalization
all_edgar_gfw_data_starting <- all_edgar_gfw_data |>
  dplyr::filter(year == analysis_start_year) |>
  dplyr::mutate(emissions_co2_mt_starting = emissions_co2_mt) |>
  dplyr::select(-c(year, emissions_co2_mt))

all_edgar_gfw_data_normalized <- all_edgar_gfw_data |>
  dplyr::inner_join(all_edgar_gfw_data_starting, by = "data_source") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_co2_mt /
      emissions_co2_mt_starting -
      1
  )


percent_transportation_emmissions_from_marine_2024 <- all_edgar_gfw_data |>
  dplyr::filter(year == 2024) |>
  dplyr::select(-year) |>
  tidyr::pivot_wider(
    names_from = "data_source",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::mutate(
    percent = scales::percent(
      `Marine vessels (AIS + S1)` /
        (`Marine vessels (AIS + S1)` + `Other transportation sectors (EDGAR)`)
    )
  ) |>
  dplyr::pull(percent)

percent_increase_from_2017_to_2024 <- all_edgar_gfw_data_normalized |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    emissions_co2_mt_normalized_percent = scales::percent(
      emissions_co2_mt_normalized
    )
  )
```


```{r}
#| label: fig-inventory-comparison
#| #| fig-cap: Comparison of marine vessel CO$_{2}$ emissions inventory estimates, by year.
#| fig-width: 9
#| fig-height: 6
marine_based_transportation_emissions_edgar <- annual_edgar_tranportation_emissions |>
  dplyr::filter(sector == "Water-borne Navigation") |>
  dplyr::group_by(year) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::mutate(data_source = "EDGAR")

imo_data <- tibble::tibble(
  year = c(2016, 2017, 2018),
  emissions_co2_mt = c(1.026e9, 1.064e9, 1.056e9)
) |>
  dplyr::mutate(data_source = "IMO")

all_inventory_data <- all_edgar_gfw_data |>
  dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>
  dplyr::mutate(data_source = "GFW (AIS + S1)") |>
  dplyr::bind_rows(marine_based_transportation_emissions_edgar) |>
  dplyr::bind_rows(oecd_data) |>
  dplyr::bind_rows(imo_data) |>
  dplyr::bind_rows(
    annual_emissions |>
      dplyr::filter(pollutant == "co2", fleet == "AIS-broadcasting") |>
      dplyr::select(year, emissions_co2_mt = emissions_mt) |>
      dplyr::mutate(data_source = "GFW (AIS)")
  ) |>
  dplyr::filter(year >= analysis_start_year, year <= analysis_end_year)

all_inventory_data |>
  ggplot(aes(x = year, y = emissions_co2_mt, color = data_source)) +
  geom_line(linewidth = 1) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "B", scale = 1e-9),
    limits = c(0, 1.5e9),
    breaks = seq(0, 1.5e9, 0.25e9)
  ) +
  theme_plot() +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(analysis_start_year, analysis_end_year)
  ) +
  labs(
    x = "",
    y = expression(Annual ~ CO[2] ~ emissions ~ (MT))
  ) +
  theme(panel.grid.minor = element_blank()) +
  scale_color_manual(
    name = "Emissions inventory",
    values = all_data_source_color_palette,
    breaks = c("GFW (AIS + S1)", "IMO", "GFW (AIS)", "OECD", "EDGAR")
  )

all_inventory_data |>
  dplyr::mutate(
    emissions_co2_mt = as.character(signif(emissions_co2_mt, 3) / 1e9)
  ) |>
  tidyr::pivot_wider(
    names_from = "data_source",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::rename(Year = year) |>
  dplyr::relocate(Year, `GFW (AIS + S1)`, `GFW (AIS)`) |>
  dplyr::mutate(across(everything(), ~ tidyr::replace_na(., "-"))) |>
  knitr::kable(
    format = "latex",
    label = "inventory_comparison",
    caption = "Comparison of annual marine vessel CO$_{2}$ emissions inventory estimates (billions of metric tonnes)."
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position") |>
  writeLines(
    here::here("tables/inventory_comparison.tex")
  )
```

```{r}
#| label: wrangle-gfw-spatials-emissions-data
# Total spatial emissions by pollutant for analysis_start_year and 2024
total_spatial_emissions_by_pollutant <- total_spatial_emissions_by_pollutant |>
  tidyr::pivot_longer(
    -c(lat_bin, lon_bin, year),
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_total_mt") |>
      stringr::str_to_upper() |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(emissions_mt_per_km2 = emissions_mt / lat_to_area(lat_bin)) |>
  dplyr::select(-emissions_mt) |>
  tidyr::pivot_wider(
    names_from = "year",
    values_from = "emissions_mt_per_km2",
    names_prefix = "emissions_mt_per_km2_"
  )
```

```{r}
# Summarize fraction of all estimate dark emissions that occurred inside the S1 footprint and in months that were imaged; inside the S1 footprint and in months that weren't imaged; and outside the S1 footprint entirely
dark_emissions_by_area <- monthly_aggregated_time_series |>
  dplyr::group_by(inside_footprint, s1_imaged) |>
  dplyr::summarise(emissions_co2_dark_mt = sum(emissions_co2_dark_mt)) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    fraction = emissions_co2_dark_mt / sum(emissions_co2_dark_mt),
    percent_total = ifelse(
      fraction < 0.01,
      scales::percent(
        fraction,
        accuracy = 0.001
      ),
      scales::percent(fraction, accuracy = 0.1)
    )
  )

# Summarize fraction of 2024 CO2 emissions that came from non-broadcasting vessels,
# disaggregated by fishing and non-fishing
dark_emissions_by_fishing <- monthly_aggregated_time_series |>
  dplyr::filter(lubridate::year(time) == 2024) |>
  dplyr::group_by(fishing) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    percent_dark = scales::percent(
      emissions_co2_dark_mt / (emissions_co2_dark_mt + emissions_co2_mt),
      accuracy = 0.1
    )
  )

monthly_fraction_fig <- monthly_aggregated_time_series |>
  dplyr::group_by(fishing, year = lubridate::year(time)) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    fraction = emissions_co2_dark_mt /
      (emissions_co2_dark_mt + emissions_co2_mt)
  ) |>
  ggplot(aes(x = year, y = fraction, color = fishing)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA))
```


```{r}
#| label: social-cost-calculation

# Social cost of carbon
# From Moore et al. 2024: https://www.pnas.org/doi/abs/10.1073/pnas.2410733121
social_cost_carbon_usd_per_mt_co2 <- 283

# Social cost of methane
# From Wang and Teng 2023: https://www.nature.com/articles/s41558-023-01803-4
social_cost_methane_usd_per_mt_ch4 <- 2900

# Social cost of n2o
# From Wang and Teng 2023: https://www.nature.com/articles/s41558-023-01803-4
social_cost_nitrous_oxide_usd_per_mt_n2o <- 49600

social_cost_tibble <- tibble::tibble(
  pollutant = c("co2", "ch4", "n2o"),
  sc_usd_per_mt = c(
    social_cost_carbon_usd_per_mt_co2,
    social_cost_methane_usd_per_mt_ch4,
    social_cost_nitrous_oxide_usd_per_mt_n2o
  )
)

annual_sc_fishing_non_fishing <- annual_emissions_all_pollutants_with_total |>
  dplyr::filter(fleet %in% c("AIS-broadcasting", "Non-broadcasting")) |>
  dplyr::group_by(year, fishing, pollutant) |>
  dplyr::summarise(emissions_mt = sum(emissions_mt)) |>
  dplyr::ungroup() |>
  dplyr::inner_join(social_cost_tibble, by = "pollutant") |>
  dplyr::filter((year == 2021 & !fishing) | (year == 2022 & fishing)) |>
  dplyr::mutate(sc_billion_usd = emissions_mt * sc_usd_per_mt / 1e9) |>
  dplyr::mutate(emissions_mt = round(emissions_mt)) |>
  dplyr::mutate(dplyr::across(
    -c(year, sc_billion_usd),
    ~ prettyNum(., big.mark = ",")
  )) |>
  dplyr::group_split(fishing) |>
  purrr::map_df(~ janitor::adorn_totals(., where = "row")) |>
  dplyr::mutate(fishing = ifelse(fishing == "-", NA_character_, fishing)) |>
  tidyr::fill(fishing, .direction = "down") |>
  dplyr::mutate(
    fishing = ifelse(fishing == "TRUE", "Fishing", "Non-fishing")
  ) |>
  dplyr::mutate(pollutant = stringr::str_to_upper(pollutant))

annual_sc_fishing_non_fishing_tbl <- annual_sc_fishing_non_fishing |>
  dplyr::select(
    Year = year,
    `Vessel type` = fishing,
    GHG = pollutant,
    `Emissions (MT)` = emissions_mt,
    `SC (USD/MT)` = sc_usd_per_mt,
    `SC (billion USD)` = sc_billion_usd
  ) |>
  knitr::kable(
    format = "latex",
    label = "social_cost_summary_tbl",
    caption = "For both non-fishing vessels and fishing vessels, we summarize: the total emissions (metric tonnes) of each GHG gas; the social cost associated with each GHG (USD/metric tonne); the social cost of each (USD); the total social cost summed across all three gasses (USD). We use 2021 for non-fishing vessels since it is the year for which we have the total value of all traded goods at sea. We use 2022 for fishing vessels since it is the year for which we have the total value of wild capture fisheries.",
    digits = 3
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

writeLines(
  annual_sc_fishing_non_fishing_tbl,
  here::here("tables/annual_sc_fishing_non_fishing_tbl.tex")
)
```

\# In-line manuscript statistics

-   The transportation sector is a major source of global greenhouse gas
    (GHG) and local air pollution emissions, responsible for
    `r fraction_co2_from_transportation` of all CO$_{2}$ emissions
    globally in 2024

-   For over
    `r scales::comma(signif(n_unique_vessels$n_unique_vessels, 2))`
    AIS-broadcasting vessels, we quantify point-estimate emissions for
    each vessel using precise latitude/longitude/timestamp information
    for over
    `r scales::comma(signif(n_ais_messages$n_billion_ais_messages, 3))`
    billion individual AIS messages

    - We obtained a dataset containing  `r signif(n_s1_detections$n_million_s1_detections,3)` million vessel detections from 2017 to 2024, of which `r signif(n_s1_detections$n_million_s1_detections_unmatched,3)` million detections were not matched to a broadcasting vessel. These detections come from across more than  `r floor(n_s1_detections$n_thousand_s1_scenes)` individual S1 scenes.

-   Because many vessels do not broadcast AIS, current approaches solely
    based on AIS \textit{underestimate} current global emissions by
    `r pollutant_ais_underestimation_overestimation_summary |>   dplyr::filter(pollutant == "co2") |>   dplyr::pull(ais_underestimation_2024)`
    (in 2024), and \textit{overestimate} the increasing trend by
    `r pollutant_ais_underestimation_overestimation_summary |>   dplyr::filter(pollutant == "co2") |>   dplyr::pull(trend_overestimation_2017_2024)`
    (from `r analysis_start_year` to `r analysis_end_year`).

-   Using the latest data available from the EU's Emissions Database for
    Global Atmospheric Research (EDGAR), we estimate that in 2024 marine
    vessels accounted for
    `r percent_transportation_emmissions_from_marine_2024` of all
    transportation-related CO$_{2}$ emissions. Marine vessel CO$_{2}$
    emissions increased by
    `r percent_increase_from_2017_to_2024 |>   dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>   dplyr::pull(emissions_co2_mt_normalized_percent)`
    between `r analysis_start_year` and 2024,
    `r round(   (percent_increase_from_2017_to_2024 |>     dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>     dplyr::pull(emissions_co2_mt_normalized)) /     (percent_increase_from_2017_to_2024 |>       dplyr::filter(data_source == "Other transportation sectors (EDGAR)") |>       dplyr::pull(emissions_co2_mt_normalized)) )`
    times faster than the growth rate of other transportation
    emissions, and `r round(   (percent_increase_from_2017_to_2024 |>     dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>     dplyr::pull(emissions_co2_mt_normalized)) /     (percent_increase_from_2017_to_2024 |>       dplyr::filter(data_source == "All sectors (EDGAR)") |>       dplyr::pull(emissions_co2_mt_normalized)) )` faster than the growth rate of total global emissions across all sectors.

-   Across the entire `r analysis_start_year` to `r analysis_end_year`
    time series,
    `r dark_emissions_by_area |>   dplyr::filter(inside_footprint & s1_imaged) |>   dplyr::pull(percent_total)`
    of non-broadcasting emissions occurred inside the S1 footprint and
    in pixel-months that were imaged;
    `r dark_emissions_by_area |>   dplyr::filter(inside_footprint & !s1_imaged) |>   dplyr::pull(percent_total)`
    occurred inside the S1 footprint but in pixel-months that weren't
    imaged; and
    `r dark_emissions_by_area |>   dplyr::filter(!inside_footprint) |>   dplyr::pull(percent_total)`
    occurred outside the S1 footprint entirely.

-   In `r analysis_end_year`,
    `r dark_emissions_by_fishing |> dplyr::filter(fishing) |> dplyr::pull(percent_dark)`
    of CO$_{2}$ emissions from fishing vessels were from
    non-broadcasting vessels, and
    `r dark_emissions_by_fishing |> dplyr::filter(!fishing) |> dplyr::pull(percent_dark)`
    of CO$_{2}$ emissions from non-fishing vessels were from
    non-broadcasting vessels. For both types of vessels, this fraction
    of emissions from non-broadcasting vessels has been declining
    steadily since 2020.

    - For non-fishing vessels, the total social cost of CO$_{2}$, CH$_{4}$, and N$_{2}$O emissions in 2021 was `r annual_sc_fishing_non_fishing |> dplyr::filter(fishing == "Non-fishing",year=="Total") |> dplyr::pull(sc_billion_usd) |> round()` billion USD. This represents about `r scales::percent((annual_sc_fishing_non_fishing |> dplyr::filter(fishing == "Non-fishing",year=="Total") |> dplyr::pull(sc_billion_usd))/1e3/6.63)` of the total value of all goods transported by maritime trade (using 6.63 trillion USD as the 2021 value of maritime-traded goods).
    
    - For fishing vessels, the total social cost of CO$_{2}$, CH$_{4}$, and N$_{2}$O emissions in 2022 was `r annual_sc_fishing_non_fishing |> dplyr::filter(fishing == "Fishing",year=="Total") |> dplyr::pull(sc_billion_usd) |> round()` billion USD. This represents about `r scales::percent((annual_sc_fishing_non_fishing |> dplyr::filter(fishing == "Fishing",year=="Total") |> dplyr::pull(sc_billion_usd))/159)` of the total value of all wild capture fisheries (using 159 billion USD as the 2022 value of wild capture fisheries).

    # Figures

    ## Figure 1

    ### Figure 1, panel A

```{r}
total_annual_emissions_by_fleet_fig <- annual_emissions |>
  dplyr::filter(pollutant == "co2") |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS",
      TRUE ~ "S1"
    )
  ) |>
  ggplot(aes(x = year, y = emissions_mt, color = fleet)) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "B", scale = 1e-9),
    limits = c(0, 1.5e9),
    breaks = seq(0, 1.5e9, 0.25e9)
  ) +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(analysis_start_year, analysis_end_year)
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS + S1", "AIS", "S1")
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(Annual ~ CO[2] ~ emissions ~ (MT))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.75, 0.4),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  )
```

### Figure 2, panel B

```{r}
# Get analysis_start_year values for time-series normalization
annual_co2_emissions_starting <- annual_emissions |>
  dplyr::filter(pollutant == "co2", year == analysis_start_year) |>
  dplyr::mutate(emissions_co2_mt_starting = emissions_mt) |>
  dplyr::select(-c(year, emissions_mt))

normalized_annual_emissions_by_fleet_fig <- annual_emissions |>
  dplyr::filter(pollutant == "co2") |>
  dplyr::left_join(annual_co2_emissions_starting, by = "fleet") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_mt / emissions_co2_mt_starting - 1
  ) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS",
      TRUE ~ "S1"
    )
  ) |>
  ggplot(aes(x = year, y = emissions_co2_mt_normalized, color = fleet)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(analysis_start_year, analysis_end_year)
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS", "AIS + S1", "S1")
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of ~ CO[2],
      emissions ~ from ~ 2017 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.25, 0.85),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  )
```

### Figure 1, panel A

```{r}
pollutant_map_data_co2 <- total_spatial_emissions_by_pollutant |>
  dplyr::filter(pollutant %in% c("CO2")) |>
  dplyr::mutate(
    emissions_mt_per_km2_delta_abs = emissions_mt_per_km2_2024 -
      emissions_mt_per_km2_2017
  ) |>
  dplyr::select(
    lon_bin,
    lat_bin,
    emissions_mt_per_km2_delta_abs,
    emissions_mt_per_km2_2024
  ) |>
  tidyr::pivot_longer(
    cols = c(emissions_mt_per_km2_delta_abs, emissions_mt_per_km2_2024)
  )

co2_map <- pollutant_map_data_co2 |>
  dplyr::mutate(
    name = ifelse(
      name == "emissions_mt_per_km2_delta_abs",
      "Change in emissions from 2017 to 2024",
      "2024 emissions"
    )
  ) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      color = value,
      fill = value
    )
  ) +
  facet_wrap(name ~ ., ncol = 1) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scico::scale_fill_scico(
    name = expression(CO[2] ~ emissions ~ (MT / km^2)),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  scico::scale_color_scico(
    name = expression(CO[2] ~ emissions ~ (MT / km^2)),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme_void() +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black"
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(2, "cm")
  )
```

### Figure 1, combined

```{r}
#| label: fig-emissions-by-data-source-and-maps
#| #| fig-cap: (A) Annual marine vessel CO2 emissions, with each line colored based on the data source. (B) Relative change in annual marine vessel CO2 emissions from a 2017 baseline, with each line colored based on the data source. (C) Spatial map of CO2 emissions in 2024 (top) and change in CO2 emissions between 2017 and 2024 (bottom), aggregated across AIS-broadcasting vessels and non-broadcasting vessels, and shown at a 1x1 degree spatial resolution; emissions are area-normalized for each pixel.
#| fig-width: 9
#| fig-height: 6
cowplot::plot_grid(
  cowplot::plot_grid(
    total_annual_emissions_by_fleet_fig,
    normalized_annual_emissions_by_fleet_fig,
    ncol = 1,
    labels = c("A", "B")
  ),
  co2_map,
  ncol = 2,
  labels = c("", "C"),
  rel_widths = c(0.4, 0.6),
  align = "vh"
)
```

### Figure 2

```{r}
#| label: fig-emissions-marine-ocean-other
#| fig-cap: (A) Trend in CO2 emissions by ocean (normalized to 2017 baesline) . (B) Comparison in trend of CO2 emissions data (normalized to 2017 baesline) between marine transportation emissions from GFW, other transportation emissions from EDGAR, and all sector emissions from EDGAR.
#| #| fig-width: 7.5
#| fig-height: 7
# Calculate normalized annual changes in CO2 emissions by major ocean basin

total_spatial_co2_emissions_by_ocean <- total_spatial_co2_emissions_by_ocean |>
  dplyr::mutate(
    emissions_co2_mt_total = emissions_co2_dark_mt + emissions_co2_mt
  )

# Get analysis_start_year values for time-series normalization
annual_emissions_by_ocean_starting <- total_spatial_co2_emissions_by_ocean |>
  dplyr::filter(year == analysis_start_year) |>
  dplyr::mutate(emissions_co2_mt_total_starting = emissions_co2_mt_total) |>
  dplyr::select(ocean, emissions_co2_mt_total_starting)

annual_emissions_by_ocean_normalized <- total_spatial_co2_emissions_by_ocean |>
  dplyr::inner_join(annual_emissions_by_ocean_starting, by = "ocean") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_co2_mt_total /
      emissions_co2_mt_total_starting -
      1
  )
ocean_data_ending <- annual_emissions_by_ocean_normalized |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    emissions_co2_mt_ending = paste0(
      prettyNum(signif(emissions_co2_mt_total / 1e6, 3), big.mark = ","),
      " MMT"
    ),
    ocean = stringr::str_to_sentence(ocean)
  ) |>
  dplyr::select(
    ocean,
    emissions_co2_mt_ending,
    emissions_co2_mt_normalized
  ) |>
  # Manually move these down, so labels don't interact with each other
  dplyr::mutate(
    emissions_co2_mt_normalized = dplyr::case_when(
      ocean == "indian" ~ emissions_co2_mt_normalized - 0.25,
      ocean == "atlantic" ~ emissions_co2_mt_normalized - .1,
      TRUE ~ emissions_co2_mt_normalized
    )
  )

annual_ocean_emissions_normalized_fig <- annual_emissions_by_ocean_normalized |>
  dplyr::mutate(ocean = paste0(stringr::str_to_sentence(ocean), " Ocean")) |>
  ggplot() +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(
    aes(x = year, y = emissions_co2_mt_normalized, color = ocean),
    linewidth = 1.1
  ) +
  scale_x_continuous(
    breaks = seq(2017, 2024),
    limits = c(2017, 2025.5)
  ) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of,
      CO[2] ~ emissions ~ from ~ 2017 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.175, 0.6),
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", color = "gray")
  ) +
  scale_color_brewer(
    palette = "Dark2",
    breaks = c(
      "Southern Ocean",
      "Arctic Ocean",
      "Pacific Ocean",
      "Indian Ocean",
      "Atlantic Ocean"
    )
  ) +
  geom_label(
    data = ocean_data_ending,
    aes(
      x = 2025,
      y = emissions_co2_mt_normalized,
      label = emissions_co2_mt_ending
    )
  )


# Summarize total emissions by data source for 2023, for putting on figure

all_edgar_gfw_data_ending <- all_edgar_gfw_data_normalized |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    emissions_co2_mt_ending = paste0(
      prettyNum(signif(emissions_co2_mt / 1e9, 3), big.mark = ","),
      " BMT"
    )
  ) |>
  dplyr::select(
    data_source,
    emissions_co2_mt_ending,
    emissions_co2_mt_normalized
  )

emissions_comparison_figure_normalized <- all_edgar_gfw_data_normalized |>
  dplyr::filter(
    data_source %in%
      c(
        "Other transportation sectors (EDGAR)",
        "Marine vessels (AIS + S1)",
        "All sectors (EDGAR)"
      )
  ) |>
  ggplot() +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(
    aes(x = year, y = emissions_co2_mt_normalized, color = data_source),
    linewidth = 1.1
  ) +
  scale_x_continuous(
    breaks = seq(2017, 2024),
    limits = c(2017, 2025.5)
  ) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of,
      CO[2] ~ emissions ~ from ~ 2017 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.35, 0.9),
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", color = "gray")
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c(
      "Marine vessels (AIS + S1)",
      "All sectors (EDGAR)",
      "Other transportation sectors (EDGAR)"
    )
  ) +
  geom_label(
    data = all_edgar_gfw_data_ending,
    aes(
      x = 2025,
      y = emissions_co2_mt_normalized,
      label = emissions_co2_mt_ending
    )
  )
cowplot::plot_grid(
  annual_ocean_emissions_normalized_fig,
  emissions_comparison_figure_normalized,
  ncol = 1,
  labels = c("A", "B")
)
```

```{r}
#| results = 'asis'

# Make and output a table summarizing 2024 CO2 emissions by ocean, and change in emissions from 2017 to 2024 by ocean
# Figure out for each ocean the fraction of observations that were imaged by s1
percent_pixels_imaged_ocean_all_time <- annual_emissions_by_ocean_normalized |>
  dplyr::group_by(ocean) |>
  dplyr::summarise(
    fraction_n_imaged = (sum(number_observations_imaged) /
      sum(number_observations)) |>
      scales::percent(accuracy = 0.01)
  ) |>
  dplyr::ungroup()
emissions_by_ocean_summary_tbl <- annual_emissions_by_ocean_normalized |>
  dplyr::filter(year == 2024) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    percent_annual_total = (emissions_co2_mt_total /
      sum(emissions_co2_mt_total)) |>
      scales::percent(accuracy = 0.01),
    percent_emissions_co2_mt_normalized = emissions_co2_mt_normalized |>
      scales::percent(accuracy = 0.01)
  ) |>
  dplyr::ungroup() |>
  dplyr::left_join(percent_pixels_imaged_ocean_all_time, by = "ocean") |>
  dplyr::mutate(
    `Percent of total (AIS+S1) CO2 emissions by non-broadcasting vessels` = (emissions_co2_dark_mt /
      emissions_co2_mt_total) |>
      scales::percent(accuracy = 0.01)
  ) |>
  dplyr::mutate(ocean = stringr::str_to_sentence(ocean)) |>
  dplyr::mutate(
    emissions_co2_mmt = prettyNum(
      signif(emissions_co2_mt / 1e6, 3),
      big.mark = ","
    )
  ) |>
  dplyr::select(
    Ocean = ocean,
    `2024 total (AIS+S1) CO2 emissions (million MT)` = emissions_co2_mmt,
    `Percent of global total (AIS+S1) CO2 2024 emissions` = percent_annual_total,
    `Percent of total (AIS+S1) CO2 emissions by non-broadcasting vessels`,
    `Percent change in total (AIS+S1) CO2 emissions from 2017 to 2024` = percent_emissions_co2_mt_normalized,
    `Percent observations imaged by S1 (across 2017-2024)` = fraction_n_imaged,
  ) |>
  tidyr::pivot_longer(cols = -Ocean) |>
  tidyr::pivot_wider(names_from = Ocean, values_from = value) |>
  dplyr::rename(`Summary statistic` = name) |>
  knitr::kable(
    format = "latex",
    label = "emissions_by_ocean_summary_tbl",
    caption = "For each ocean, summary of: 2024 total (AIS+S1) CO$_2$ emissions (metric tonnes); percentage of global total (AIS+S1) 2024 CO$_2$ emissions from across all oceans; percent of 2024 total (AIS+S1) CO$_2$ emissions by non-broadcasting vessels; percent change in total (AIS+S1) CO$_2$ emissions from 2017 to 2024; and the percentage of all observations in the dataset (pixel-months) from 2017-2024 that were imaged by S1.",
    digits = 3
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

emissions_by_ocean_summary_tbl

writeLines(
  emissions_by_ocean_summary_tbl,
  here::here("tables/emissions_by_ocean_summary_tbl.tex")
)
```

```{r}
#| results = 'asis'

# Make and output a table summarizing 2024 CO2 emissions by data source, and change in emissions from 2017 to 2024 by data source
emissions_by_data_source_summary_tbl <- all_edgar_gfw_data_normalized |>
  dplyr::filter(
    data_source %in%
      c(
        "Other transportation sectors (EDGAR)",
        "Marine vessels (AIS + S1)",
        "All sectors (EDGAR)"
      )
  ) |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    emissions_co2_bmt = prettyNum(
      signif(emissions_co2_mt / 1e9, 3),
      big.mark = ","
    ),
    percent_emissions_co2_mt_normalized = emissions_co2_mt_normalized |>
      scales::percent(accuracy = 0.01),
  ) |>
  dplyr::ungroup() |>
  dplyr::select(
    `Data source` = data_source,
    `CO2 emissions in 2024 (billion MT)` = emissions_co2_bmt,
    `Percent change in CO2 emissions from 2017 to 2024` = percent_emissions_co2_mt_normalized
  ) |>
  knitr::kable(
    format = "latex",
    label = "emissions_by_data_source_summary_tbl",
    caption = "For each data source, summary of: total 2024 CO$_2$ emissions (metric tonnes); and percent change in CO$_2$ emissions from 2017 to 2024.",
    digits = 3
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

emissions_by_data_source_summary_tbl

writeLines(
  emissions_by_data_source_summary_tbl,
  here::here("tables/emissions_by_data_source_summary_tbl.tex")
)
```


# Supplementary figures

### Relationship between length and engine power

```{r}
#| label: fig-ais-length-power-relationship
#| fig-cap: The relationship between main engine power (kilowatts) and length (meters), where each point is a specific AIS-broadcasting vessel. The relationship is disaggregated for non-fishing vessels and fishing vessels.
#| fig-height: 7
#| fig-width: 6
vessel_size_info |>
  dplyr::mutate(
    fishing = ifelse(
      fishing,
      "Fishing",
      "Non-fishing"
    ) |>
      forcats::fct_rev()
  ) |>
  ggplot(aes(x = length_m, y = main_engine_power_kw)) +
  geom_point(size = 0.1, alpha = 0.1) +
  theme_plot() +
  facet_wrap(fishing ~ ., ncol = 1, scales = "free") +
  labs(x = "Length (m)", y = "Main engine power (kW)") +
  scale_y_continuous(labels = scales::comma)
```

### Dark fleet emissions, by vessel size and type

```{r}
#| label: fig-global-annual-co2-dark-emissions-by-type
#| fig-cap: Summary of global annual CO2 dark emissions over time, disaggregated by vessel type.
#| fig-height: 8
#| fig-width: 8
#| include: false
#| eval: false

summarized_s1_dark_fleet_data |>
  ggplot(aes(
    x = year,
    y = emissions_co2_dark_mt,
    fill = factor(length_size_class_percentile)
  )) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_continuous(breaks = unique(summarized_s1_dark_fleet_data$year)) +
  labs(x = "", y = "CO2 dark emissions (MT)", fill = "Length bin") +
  theme_plot() +
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  facet_grid(fishing ~ ., scales = "free_y") +
  paletteer::scale_fill_paletteer_d(palette = "jcolors::pal10")
```

```{r}
#| label: fig-pollutant-maps-pseudolog
#| fig-cap: Spatial map of 2024 emissions for each pollutant, aggregated across the AIS-broadcasting fleet and non-broadcasting, and shown at a 1x1 degree spatial resolution
#| fig-width: 8
#| fig-height: 8

pollutant_map_data_all <- total_spatial_emissions_by_pollutant |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

pollutant_map_list <- purrr::map(
  unique(pollutant_map_data_all$pollutant),
  function(pollutant_tmp) {
    pollutant_tmp <- as.character(pollutant_tmp)

    pollutant_map_data_all |>
      dplyr::filter(pollutant == pollutant_tmp) |>
      ggplot() +
      geom_sf(data = world_bbox_sf, fill = "grey90") +
      ggspatial::geom_spatial_tile(
        aes(
          x = lon_bin + 0.5,
          y = lat_bin + 0.5,
          color = emissions_mt_per_km2_2024,
          fill = emissions_mt_per_km2_2024
        )
      ) +
      geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
      scico::scale_fill_scico(
        name = expression(Emissions ~ (MT / km^2)),
        palette = "vik",
        midpoint = 0,
        trans = ggallin::pseudolog10_trans,
        na.value = "grey90",
        labels = scales::comma
      ) +
      scico::scale_color_scico(
        name = expression(Emissions ~ (MT / km^2)),
        palette = "vik",
        midpoint = 0,
        trans = ggallin::pseudolog10_trans,
        na.value = "grey90",
        labels = scales::comma
      ) +
      labs(x = "", y = "") +
      theme_void() +
      guides(
        fill = guide_colorbar(
          title.position = "top",
          title.hjust = 0.5,
          frame.colour = "black",
          ticks.colour = "black"
        )
      ) +
      scale_x_continuous(expand = c(0, 1)) +
      scale_y_continuous(expand = c(0, 1)) +
      theme(
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.key.width = unit(1, "cm")
      ) +
      facet_wrap(pollutant ~ ., labeller = label_parsed)
  }
)

#cowplot::plot_grid(plotlist = pollutant_map_list, ncol = 3)
```


```{r}
#| label: fig-pollutant-maps-qlog10
#| fig-cap: Spatial map of 2024 emissions for each pollutant, aggregated across the AIS-broadcasting fleet and non-broadcasting, and shown at a 1x1 degree spatial resolution. A log-10 scale is applied for visualization to manage the high dynamic range of emissions. To ensure visual clarity and suppress background noise, the bottom 1% of the emission distribution has been excluded from the color mapping.
#| fig-width: 8
#| fig-height: 8

pollutant_map_data_all <- total_spatial_emissions_by_pollutant |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

pollutant_map_list <- purrr::map(
  unique(pollutant_map_data_all$pollutant),
  function(pollutant_tmp) {
    pollutant_tmp <- as.character(pollutant_tmp)

    pollutant_map_data <- pollutant_map_data_all |>
      dplyr::filter(pollutant == pollutant_tmp)

    scale_limits <- pollutant_map_data |>
      dplyr::summarize(
        min = quantile(emissions_mt_per_km2_2024, 0.01, na.rm = TRUE),
        max = max(emissions_mt_per_km2_2024, na.rm = TRUE)
      ) |>
      as.numeric()

    pollutant_map_data |>
      ggplot() +
      geom_sf(data = world_bbox_sf, fill = "grey90") +
      ggspatial::geom_spatial_tile(
        aes(
          x = lon_bin + 0.5,
          y = lat_bin + 0.5,
          color = emissions_mt_per_km2_2024,
          fill = emissions_mt_per_km2_2024
        )
      ) +
      geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
      scico::scale_fill_scico(
        name = expression(Emissions ~ (MT / km^2)),
        palette = "vik",
        begin = 0.5,
        end = 1,
        limits = scale_limits,
        trans = "log10",
        na.value = "grey90",
        labels = scales::scientific
      ) +
      scico::scale_color_scico(
        name = expression(Emissions ~ (MT / km^2)),
        palette = "vik",
        begin = 0.5,
        end = 1,
        limits = scale_limits,
        trans = "log10",
        na.value = "grey90",
        labels = scales::scientific
      ) +
      labs(x = "", y = "") +
      theme_void() +
      guides(
        fill = guide_colorbar(
          title.position = "top",
          title.hjust = 0.5,
          frame.colour = "white",
          ticks.colour = "white"
        )
      ) +
      scale_x_continuous(expand = c(0, 1)) +
      scale_y_continuous(expand = c(0, 1)) +
      theme(
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.key.width = unit(1, "cm")
      ) +
      facet_wrap(pollutant ~ ., labeller = label_parsed)
  }
)

cowplot::plot_grid(plotlist = pollutant_map_list, ncol = 3)
```


```{r}
#| label: make-fig-co2-emissions-time-series

# Total monthly non-spatial emissions by pollutant
total_monthly_emissions_by_pollutant <- total_monthly_emissions_by_pollutant |>
  dplyr::mutate(month = lubridate::ymd(month)) |>
  tidyr::pivot_longer(
    -month,
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_total_mt") |>
      stringr::str_to_upper() |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::filter(
    lubridate::year(month) <= analysis_end_year,
    lubridate::year(month) >= analysis_start_year
  )

pollutant_time_series_data_co2 <- total_monthly_emissions_by_pollutant |>
  dplyr::filter(pollutant == "CO2")

```


```{r}
#| label: fig-ais-data-richness
#| fig-cap: (A) 2024 CO$_{2}$ emissions by vessel type for AIS-broadcasting vessels. (B) 2024 CO$_{2}$ emissions by country for AIS-broadcasting vessels during domestic trips, international trips (where emissions of each trip are partitioned equally to its departure and arrival country), and during port stays. Countries with no values are shown in black.
#| fig-width: 9
#| fig-height: 8

time_series_data_monthly_emissions_by_fishing <- monthly_aggregated_time_series |>
  dplyr::mutate(
    time = lubridate::ymd(time)
  ) |>
  dplyr::group_by(time, fishing) |>
  dplyr::summarise(
    total_emissions_co2_mt = sum(emissions_co2_mt + emissions_co2_dark_mt)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fishing == TRUE ~ "Fishing",
      fishing == FALSE ~ "Non-fishing"
    ),
    fleet = factor(fleet, levels = c("Non-fishing", "Fishing"))
  )

time_series_by_fishing_fig_co2 <- ggplot(
  time_series_data_monthly_emissions_by_fishing,
  aes(x = time, y = total_emissions_co2_mt, fill = fleet)
) +
  geom_area(position = 'stack', alpha = 0.9) +
  geom_line(
    data = subset(
      time_series_data_monthly_emissions_by_fishing,
      fleet == "Fishing"
    ),
    aes(x = time, y = total_emissions_co2_mt),
    color = "white"
  ) +
  scale_y_continuous(
    labels = scales::comma,
    limits = c(0, NA)
  ) +
  scale_fill_manual(
    values = all_data_source_color_palette,
    breaks = c("Non-fishing", "Fishing")
  ) +
  scale_x_continuous(
    breaks = seq(
      lubridate::ymd(glue::glue("{analysis_start_year}-01-01")),
      lubridate::ymd(glue::glue("{analysis_end_year+1}-01-01")),
      by = "1 year"
    ),
    labels = seq(analysis_start_year, analysis_end_year + 1)
  ) +
  theme_plot() +
  labs(x = "", y = expression(Monthly ~ CO[2] ~ emissions ~ (MT))) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.15, 0.9),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  ) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  )

emissions_by_vessel_class_barplot <- annual_ais_co2_emissions_by_vessel_type |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    fishing = ifelse(
      vessel_class %in%
        c(
          "trawlers",
          "squid_jigger",
          "tuna_purse_seines",
          "set_longlines",
          "pole_and_line",
          "set_gillnets",
          "pots_and_traps",
          "dredge_fishing",
          "other_seines",
          "other_purse_seines",
          "trollers",
          "drifting_longlines",
          "driftnets",
          "fish_factory",
          "other_fishing"
        ),
      "Fishing",
      "Non-fishing"
    ) |>
      forcats::fct_rev()
  ) |>
  dplyr::mutate(
    vessel_class = stringr::str_replace_all(
      vessel_class,
      "\\.",
      ": "
    ) |>
      stringr::str_replace_all("_", " ") |>
      stringr::str_to_sentence() |>
      stringr::str_replace_all(
        "chemical oil",
        "chemical or oil"
      )
  ) |>
  ggplot(aes(
    x = emissions_co2_mt,
    y = reorder(vessel_class, emissions_co2_mt)
  )) +
  geom_col(position = position_dodge2(preserve = "single")) +
  ggforce::facet_col(~fishing, space = "free", scales = "free") +
  theme_plot() +
  labs(
    x = expression(2024 ~ CO[2] ~ emissions ~ (MT)),
    y = ""
  ) +
  scale_x_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  )


# Divde international trip emissions equally to from and to countries
international_trip_emissions_by_country <- trip_co2_emissions_by_from_to_countries |>
  dplyr::filter(from_country_iso3 != to_country_iso3) |>
  dplyr::mutate(emissions_co2_mt = emissions_co2_mt / 2) |>
  dplyr::select(country_iso3 = from_country_iso3, emissions_co2_mt) |>
  dplyr::bind_rows(
    trip_co2_emissions_by_from_to_countries |>
      dplyr::mutate(emissions_co2_mt = emissions_co2_mt / 2) |>
      dplyr::select(country_iso3 = to_country_iso3, emissions_co2_mt)
  ) |>
  dplyr::group_by(country_iso3) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt)) |>
  dplyr::ungroup()

domestic_trip_emissions_by_country <- trip_co2_emissions_by_from_to_countries |>
  dplyr::filter(from_country_iso3 == to_country_iso3) |>
  dplyr::select(country_iso3 = from_country_iso3, emissions_co2_mt)

all_port_and_trip_emissions <- international_trip_emissions_by_country |>
  dplyr::mutate(type = "International trips") |>
  dplyr::bind_rows(
    domestic_trip_emissions_by_country |>
      dplyr::mutate(type = "Domestic trips")
  ) |>
  dplyr::bind_rows(
    port_visit_co2_emissions_by_country |>
      dplyr::rename(country_iso3 = port_country_iso3) |>
      dplyr::mutate(type = "Port stays")
  )


all_countries_for_plot <- land |>
  sf::st_drop_geometry() |>
  dplyr::select(country_iso3) |>
  tidyr::crossing(
    all_port_and_trip_emissions |>
      dplyr::distinct(type)
  )

port_trip_emissions_fig <- ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  geom_sf(
    data = all_countries_for_plot |>
      dplyr::left_join(land, by = ) |>
      dplyr::left_join(
        all_port_and_trip_emissions,
        by = c("country_iso3", "type")
      ) |>
      sf::st_as_sf(),
    aes(fill = emissions_co2_mt),
    color = "white",
    linewidth = 0.01
  ) +
  facet_wrap(type ~ ., ncol = 1) +
  coord_sf(crs = map_projection) +
  theme_void() +
  scale_fill_gradientn(
    # Get the "Blues" palette with 9 colors, but keep only the darker 5 (indices 4-9)
    colors = rev(RColorBrewer::brewer.pal(9, "Blues")[5:9]),
    labels = scales::unit_format(unit = "M", scale = 1e-6)
  ) +
  labs(
    x = "",
    fill = expression(2024 ~ CO[2] ~ emissions ~ (MT)),
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "white",
      ticks.colour = "white"
    )
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(1.75, "cm")
  )

# cowplot::plot_grid(
#   cowplot::plot_grid(
#     time_series_by_fishing_fig_co2,
#     emissions_by_vessel_class_barplot,
#     ncol = 1,
#     rel_heights = c(0.4, 0.6),
#     labels = c("A", "B")
#   ),
#   port_trip_emissions_fig,
#   labels = c("", "C"),
#   rel_widths = c(0.52, 0.48)
# )

cowplot::plot_grid(
  emissions_by_vessel_class_barplot,
  port_trip_emissions_fig,
  ncol = 2,
  labels = c("A", "B"),
  rel_widths = c(0.52, 0.48)
)
```

# Figure 3.1 -----



#### Spatial emissions by fleet (annual)

```{r}
#| label: fig-spatial-emissions-by-fleet-2024
#| fig-cap: Spatial distribution of cumulative CO$_{2}$ emissions (Mt/km2) for AIS/S1, and fishing/non-fishing fleets, from 2017 to 2024, in MT/Km$^{2}$
#| fig-width: 8
#| fig-height: 4

target_lons <- unique(as.integer(longs))
target_lats <- unique(as.integer(lats))

spatial_co2_emissions_ais_dark_by_fleet <- annual_spatial_co2_emissions_ais_dark_by_fleet |>
  dplyr::group_by(year, lon_bin, lat_bin, fishing) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt, na.rm = TRUE)
  ) |>
  dplyr::ungroup() |>
  dplyr::filter(year == analysis_end_year) |>
  tidyr::pivot_longer(
    cols = c(emissions_co2_mt, emissions_co2_dark_mt),
    names_to = "fleet",
    values_to = "emissions_co2_mt"
  ) |>
  dplyr::mutate(
    emissions_mt_per_km2 = emissions_co2_mt / lat_to_area(lat_bin)
  ) |>
  tidyr::complete(
    lon_bin = target_lons,
    lat_bin = target_lats,
    fishing,
    fleet,
    fill = list(emissions_mt_per_km2 = NA)
  )

scale_limits <- c(
  0.001,
  max(
    spatial_co2_emissions_ais_dark_by_fleet$emissions_mt_per_km2,
    na.rm = TRUE
  )
)

spatial_emissions_by_fleet_2024_plot <- spatial_co2_emissions_ais_dark_by_fleet |>
  dplyr::mutate(
    fleet = factor(
      fleet,
      levels = c("emissions_co2_mt", "emissions_co2_dark_mt"),
      labels = c("AIS", "S1")
    )
  ) |>
  dplyr::mutate(
    fishing = ifelse(fishing, "Fishing", "Non-fishing") |>
      forcats::fct_rev()
  ) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "black") +
  ggspatial::geom_spatial_tile(aes(
    x = lon_bin + 0.5,
    y = lat_bin + 0.5,
    fill = emissions_mt_per_km2,
    color = emissions_mt_per_km2
  )) +
  theme_void() +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scico::scale_color_scico(
    name = expression(
      2024 ~ CO[2] ~ emissions ~ (MT / km^2)
    ),
    palette = "vik",
    begin = 0.5,
    end = 1,
    limits = scale_limits,
    trans = "log10",
    na.value = "grey90",
    # labels = scales::comma,
    labels = c("0", "0.1", "10", "1000"),
    guide = "none"
  ) +
  scico::scale_fill_scico(
    name = expression(
      2024 ~ CO[2] ~ emissions ~ (MT / km^2)
    ),
    palette = "vik",
    begin = 0.5,
    end = 1,
    limits = scale_limits,
    trans = "log10",
    na.value = "grey90",
    # labels = scales::comma
    labels = c("0", "0.1", "10", "1000")
  ) +
  labs(x = "", y = "") +
  theme(
    legend.position = "bottom",
    legend.title.align = 0.5,
    strip.text.x = element_text(
      margin = margin(t = 10, b = 10),
      size = 10
    ),
    strip.text.y = element_text(
      angle = 0,
      hjust = 0,
      margin = margin(l = 10, r = 10),
      size = 10
    ),
    strip.background = element_rect(fill = "transparent", color = NA),
    panel.spacing = unit(1, "lines")
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "white",
      ticks.colour = "white",
      barwidth = unit(10, "cm"),
      barheight = unit(0.5, "cm")
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  facet_grid(fleet ~ fishing) +
  coord_sf(crs = map_projection)

spatial_emissions_by_fleet_2024_plot
```

#### Spatial emissions by fleet (total)

```{r}
#| label: fig-spatial-emissions-by-fleet-total-qlog10
#| fig-cap: Spatial distribution of cumulative CO$_{2}$ emissions (Mt/km2) for AIS/S1, and fishing/non-fishing fleets, from 2017 to 2024, in MT/Km$^{2}. A log-10 scale is applied for visualization to manage the high dynamic range of emissions. To ensure visual clarity and suppress background noise, the bottom 1% of the emission distribution has been excluded from the color mapping.$
#| fig-width: 8
#| fig-height: 4

target_lons <- unique(as.integer(longs))
target_lats <- unique(as.integer(lats))

spatial_co2_emissions_ais_dark_by_fleet <- annual_spatial_co2_emissions_ais_dark_by_fleet |>
  dplyr::group_by(lon_bin, lat_bin, fishing) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt, na.rm = TRUE)
  ) |>
  dplyr::ungroup() |>
  tidyr::pivot_longer(
    cols = c(emissions_co2_mt, emissions_co2_dark_mt),
    names_to = "fleet",
    values_to = "emissions_co2_mt"
  ) |>
  dplyr::mutate(
    emissions_mt_per_km2 = emissions_co2_mt / lat_to_area(lat_bin),
    # emissions_mt_per_km2_year = emissions_mt_per_km2 / length(analysis_end_year:analysis_start_year) # If use, update SQL to pull data by year
  ) |>
  tidyr::complete(
    lon_bin = target_lons,
    lat_bin = target_lats,
    fishing,
    fleet,
    fill = list(emissions_mt_per_km2 = NA)
  )


scale_limits <- spatial_co2_emissions_ais_dark_by_fleet |>
  dplyr::filter(fleet == "emissions_co2_mt") |>
  dplyr::summarize(
    min = quantile(emissions_mt_per_km2, 0.01, na.rm = TRUE),
    max = max(emissions_mt_per_km2, na.rm = TRUE)
  ) |>
  as.numeric()

spatial_emissions_by_fleet_total_plot_qlog10 <- spatial_co2_emissions_ais_dark_by_fleet |>
  dplyr::mutate(
    fleet = factor(
      fleet,
      levels = c("emissions_co2_mt", "emissions_co2_dark_mt"),
      labels = c("AIS", "S1")
    )
  ) |>
  dplyr::mutate(
    fishing = ifelse(fishing, "Fishing", "Non-fishing") |>
      forcats::fct_rev()
  ) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "black") +
  ggspatial::geom_spatial_tile(aes(
    x = lon_bin + 0.5,
    y = lat_bin + 0.5,
    fill = emissions_mt_per_km2,
    color = emissions_mt_per_km2
  )) +
  theme_void() +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scico::scale_color_scico(
    name = expression(
      Cumulative ~ CO[2] ~ emissions ~ (MT / km^2) ~ from ~ 2017 ~ to ~ 2024
    ),
    palette = "vik",
    begin = 0.5,
    end = 1,
    limits = scale_limits,
    trans = "log10",
    na.value = "grey90",
    labels = scales::comma,
    guide = "none"
  ) +
  scico::scale_fill_scico(
    name = expression(
      Cumulative ~ CO[2] ~ emissions ~ (MT / km^2) ~ from ~ 2017 ~ to ~ 2024
    ),
    palette = "vik",
    begin = 0.5,
    end = 1,
    limits = scale_limits,
    trans = "log10",
    na.value = "grey90",
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme(
    legend.position = "bottom",
    legend.title.align = 0.5,
    strip.text.x = element_text(
      margin = margin(t = 10, b = 10),
      size = 10
    ),
    strip.text.y = element_text(
      angle = 0,
      hjust = 0,
      margin = margin(l = 10, r = 10),
      size = 10
    ),
    strip.background = element_rect(fill = "transparent", color = NA),
    panel.spacing = unit(1, "lines")
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "white",
      ticks.colour = "white",
      barwidth = unit(10, "cm"),
      barheight = unit(0.5, "cm")
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  facet_grid(fleet ~ fishing) +
  coord_sf(crs = map_projection)
```

```{r}
#| label: fig-spatial-emissions-by-fleet-total-pseudolog
#| fig-cap: Spatial distribution of cumulative CO$_{2}$ emissions (Mt/km2) for AIS/S1, and fishing/non-fishing fleets, from 2017 to 2024, in MT/Km$^{2}. A pseudolog-10 scale is applied for visualization to manage the high dynamic range of emissions, suppressing near-zero values while preserving emission hotspots.$
#| fig-width: 8
#| fig-height: 4

spatial_emissions_by_fleet_total_plot_pseudolog <- spatial_co2_emissions_ais_dark_by_fleet |>
  dplyr::mutate(
    fleet = factor(
      fleet,
      levels = c("emissions_co2_mt", "emissions_co2_dark_mt"),
      labels = c("AIS", "S1")
    )
  ) |>
  dplyr::mutate(
    fishing = ifelse(fishing, "Fishing", "Non-fishing") |>
      forcats::fct_rev()
  ) |>
  dplyr::mutate(
    emissions_mt_per_km2 = dplyr::na_if(emissions_mt_per_km2, 0)
  ) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey92") +
  ggspatial::geom_spatial_tile(aes(
    x = lon_bin + 0.5,
    y = lat_bin + 0.5,
    fill = emissions_mt_per_km2,
    color = emissions_mt_per_km2
  )) +
  theme_void() +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scico::scale_color_scico(
    name = expression(
      Cumulative ~ CO[2] ~ emissions ~ (MT / km^2) ~ from ~ 2017 ~ to ~ 2024
    ),
    palette = "vik",
    begin = 0.5,
    end = 1,
    trans = ggallin::pseudolog10_trans,
    breaks = c(10, 100, 1000, 10000),
    na.value = "grey95",
    labels = scales::comma,
    guide = "none"
  ) +
  scico::scale_fill_scico(
    name = expression(
      Cumulative ~ CO[2] ~ emissions ~ (MT / km^2) ~ from ~ 2017 ~ to ~ 2024
    ),
    palette = "vik",
    begin = 0.5,
    end = 1,
    trans = ggallin::pseudolog10_trans,
    breaks = c(10, 100, 1000, 10000),
    na.value = "grey95",
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme(
    legend.position = "bottom",
    legend.title.align = 0.5,
    strip.text.x = element_text(
      margin = margin(t = 10, b = 10),
      size = 10
    ),
    strip.text.y = element_text(
      angle = 0,
      hjust = 0,
      margin = margin(l = 10, r = 10),
      size = 10
    ),
    strip.background = element_rect(fill = "transparent", color = NA),
    panel.spacing = unit(1, "lines")
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "white",
      ticks.colour = "white",
      barwidth = unit(10, "cm"),
      barheight = unit(0.5, "cm")
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  facet_grid(fleet ~ fishing) +
  coord_sf(crs = map_projection)
```

```{r}
#| label: fig-spatial-temporal-richness-by-fleet-total-qlog10
#| fig-cap: (A) Time series of monthly global CO$_2$ emissions, aggregated across AIS-broadcasting and non-broadcasting fleets, for both fishing and non-fishing boats. (B) Spatial distribution of cumulative CO$_{2}$ emissions (Mt/km2) for AIS/S1, and fishing/non-fishing fleets, from 2017 to 2024, in MT/Km$^{2}. A log-10 scale is applied for visualization to manage the high dynamic range of emissions. To ensure visual clarity and suppress background noise, the bottom 1% of the emission distribution has been excluded from the color mapping.$
#| fig-width: 8
#| fig-height: 7

time_series_data_monthly_emissions_by_fleet <- monthly_aggregated_time_series |>
  dplyr::group_by(time, fishing) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(time = lubridate::ymd(time)) |>
  tidyr::pivot_longer(
    cols = c(emissions_co2_mt, emissions_co2_dark_mt),
    names_to = "fleet",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "emissions_co2_mt" ~ "AIS",
      fleet == "emissions_co2_dark_mt" ~ "S1"
    ),
    fleet = factor(fleet, levels = c("AIS", "S1")),
    fishing = dplyr::case_when(
      fishing == TRUE ~ "Fishing",
      fishing == FALSE ~ "Non-fishing"
    ),
    fishing = factor(fishing, levels = c("Non-fishing", "Fishing"))
  )

make_pollutant_by_fleet_time_series_figure <- function(
  pollutant_time_series_data,
  y_axis_label
) {
  ggplot(
    pollutant_time_series_data,
    aes(x = time, y = emissions_mt, fill = fleet)
  ) +
    geom_area(position = 'stack', alpha = 0.9) +
    geom_line(
      data = subset(pollutant_time_series_data, fleet == "S1"),
      aes(x = time, y = emissions_mt),
      color = "white"
    ) +
    scale_y_continuous(
      labels = scales::comma,
      limits = c(0, NA)
    ) +
    scale_fill_manual(
      values = all_data_source_color_palette,
      breaks = c("AIS", "S1")
    ) +
    scale_x_continuous(
      breaks = seq(
        lubridate::ymd(glue::glue("{analysis_start_year}-01-01")),
        lubridate::ymd(glue::glue("{analysis_end_year+1}-01-01")),
        by = "1 year"
      ),
      labels = seq(analysis_start_year, analysis_end_year + 1)
    ) +
    facet_wrap(~fishing, nrow = 1, scales = "free_y") +
    theme_plot() +
    labs(x = "", y = y_axis_label) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = c(0.60, 0.90),
      legend.background = element_rect(fill = "white", color = "gray"),
      legend.title = element_blank(),
      axis.title.y = element_text(size = 9),
      axis.text.x = element_text(angle = 25, hjust = 0.7),
      legend.text = element_text(size = 8),
      legend.key.size = unit(0.5, "cm"),
      strip.text.x = element_text(size = 12)
    )
}

time_series_by_fleet_fig_co2 <- make_pollutant_by_fleet_time_series_figure(
  time_series_data_monthly_emissions_by_fleet,
  y_axis_label = expression(Monthly ~ CO[2] ~ emissions ~ (MT))
) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  )

plot_constrained <- cowplot::plot_grid(
  NULL,
  time_series_by_fleet_fig_co2 +
    theme(plot.margin = margin(b = -15, l = 0)),
  NULL,
  ncol = 3,
  rel_widths = c(0.02, 0.88, 0.10)
)
```


```{r}
#| label: fig-spatial-temporal-richness-by-fleet-total-pseudolog
#| fig-cap: (A) Time series of monthly global CO$_2$ emissions, aggregated across AIS-broadcasting and non-broadcasting fleets, for both fishing and non-fishing boats. (B) Spatial distribution of cumulative CO$_{2}$ emissions (Mt/km2) for AIS/S1, and fishing/non-fishing fleets, from 2017 to 2024, in MT/Km$^{2}. A pseudolog-10 scale is applied for visualization to manage the high dynamic range of emissions, suppressing near-zero values while preserving emission hotspots.$
#| fig-width: 8
#| fig-height: 7

cowplot::plot_grid(
  plot_constrained,
  spatial_emissions_by_fleet_total_plot_pseudolog,
  ncol = 1,
  rel_heights = c(0.35, 0.65),
  labels = c("A", "B")
)

```

### Other plots

```{r}
#| label: tbl-top-port-stay-emissions-by-country
#| tbl-cap: Top 10 countries in terms of 2024 port stay emissions.
all_port_and_trip_emissions |>
  dplyr::mutate(
    country = countrycode::countrycode(
      country_iso3,
      "iso3c",
      "country.name"
    )
  ) |>
  dplyr::group_by(type) |>
  dplyr::slice_max(n = 3, order_by = emissions_co2_mt) |>
  knitr::kable()
```

```{r}
#| label: fig-pollutant-time-series
#| fig-cap: Time series of monthly global emissions for each pollutant, aggregated across AIS-broadcasting vessels and non-broadcasting vessels.
#| fig-width: 8
#| fig-height: 8

pollutant_time_series_data_all <- total_monthly_emissions_by_pollutant |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

make_pollutant_time_series_figure <- function(
  pollutant_time_series_data,
  y_axis_label
) {
  {
    ggplot(
      data = pollutant_time_series_data,
      aes(x = month, y = emissions_mt)
    ) +
      geom_line(linewidth = 1.1) +
      scale_y_continuous(labels = scales::comma) +
      theme_plot() +
      labs(x = "", y = y_axis_label) +
      theme(
        panel.grid.minor.x = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0)
      ) +
      scale_x_continuous(
        breaks = seq(
          lubridate::ymd(glue::glue("{analysis_start_year}-01-01")),
          lubridate::ymd(glue::glue("{analysis_end_year+1}-01-01")),
          by = "1 year"
        ),
        labels = seq(analysis_start_year, analysis_end_year + 1)
      ) +
      scale_y_continuous(limits = c(0, NA), labels = scales::comma)
  }
}

make_pollutant_time_series_figure(
  pollutant_time_series_data_all,
  y_axis_label = "Monthly emissions (MT)"
) +
  facet_wrap(
    pollutant ~ .,
    ncol = 3,
    scales = "free_y",
    labeller = label_parsed
  )
```


### Spatial emissions by footprint group

```{r}
#| label: fig-spatial-coverage-footprint
#| fig-cap: Spatial coverage of the non-broadcasting emissions model. Areas are classified according to whether they fall inside or outside the S1 footprint and whether they have been imaged by S1 at any point in time. A given pixel may be imaged in some months but not in others. Pixels outside the footprint have never been imaged. The percentage of total CO$_{2}$ emissions from non-broadcasting vessels attributed to each category is shown.
#| fig-width: 5
#| fig-height: 7

total_spatial_co2_emissions_dark_by_footprint |>
  dplyr::left_join(
    dark_emissions_by_area |>
      dplyr::select(inside_footprint, s1_imaged, percent_total),
    by = c("inside_footprint", "s1_imaged")
  ) |>
  dplyr::mutate(
    footprint = dplyr::case_when(
      inside_footprint == FALSE ~ "Outside S1 footprint",
      inside_footprint == TRUE &
        s1_imaged == TRUE ~ "Inside S1 footprint imaged",
      inside_footprint == TRUE &
        s1_imaged == FALSE ~ "Inside S1 footprint not imaged",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::group_by(footprint) |>
  dplyr::mutate(
    facet_label = glue::glue(
      "{footprint} ({percent_total} of non-broadcasting emissions)"
    )
  ) |>
  dplyr::ungroup() |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90", color = NA) +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      fill = footprint
    ),
    color = NA
  ) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scale_fill_manual(
    name = "",
    values = c(
      "Outside S1 footprint" = "#d90101ff",
      "Inside S1 footprint imaged" = "#d90101ff",
      "Inside S1 footprint not imaged" = "#d90101ff"
    )
  ) +
  theme_void() +
  facet_wrap(~facet_label, ncol = 1) +
  theme(legend.position = "none")
```

### Registered data validation

```{r}
#| label: fig-registered-data-performance
#| fig-cap: Comparison of registered and simulated CO$_{2}$ emissions from main engine fuel consumption over 24 hours. Estimates are shown using main engine power and design speed from IMO (left), our machine learning random forest models (center), and the same registered dataset (right). The solid red line shows a linear regression fit, while the dashed line represents the 1:1 line. Performance values for traditional R$^{2}$ are shown for each set of vessel characteristics. Blue shading indicates point density. All data are rule-of-three compliant, meaning each cell (20×20 mt) displayed contains at least three vessels.
#| fig-width: 8
#| fig-height: 4
model_levels <- c(
  "IMO report lookup table",
  "ML-derived engine power\nand design speed",
  "Registered data"
)

vessel_long <- registered_validation_data |>
  dplyr::select(
    co2_emissions_tonnes_registered,
    "IMO report lookup table" = co2_emissions_tonnes_estimate_imo,
    "ML-derived engine power\nand design speed" = co2_emissions_tonnes_estimate_rf,
    "Registered data" = co2_emissions_tonnes_estimate_registered
  ) |>
  tidyr::pivot_longer(
    cols = -co2_emissions_tonnes_registered,
    names_to = "Model",
    values_to = "Estimate"
  ) |>
  dplyr::mutate(Model = factor(Model, levels = model_levels)) # Order levels


r2_labels <- vessel_long |>
  dplyr::group_by(Model) |>
  yardstick::rsq_trad(
    truth = co2_emissions_tonnes_registered,
    estimate = Estimate
  ) |>
  dplyr::mutate(label = paste0("Traditional R² = ", round(.estimate, 2)))

ggplot(
  vessel_long,
  aes(
    x = co2_emissions_tonnes_registered,
    y = Estimate
  )
) +
  geom_bin2d(
    bins = 30,
    aes(
      fill = after_stat(count),
      alpha = after_stat(ifelse(count < 3, 0, 1)) # Full transparency for counts < 3
    )
  ) +
  scale_fill_gradientn(
    colours = colorRampPalette(RColorBrewer::brewer.pal(9, "Blues"))(256),
    trans = "log10",
    name = "Count",
    guide = guide_colorbar(
      ticks.linewidth = 0.3,
      frame.colour = "black"
    )
  ) +
  geom_smooth(
    method = "lm",
    linewidth = 0.5,
    color = "red",
    linetype = "solid",
    se = FALSE
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_text(
    data = r2_labels |>
      dplyr::filter(!Model %in% c("RF design speed", "RF engine power")),
    aes(label = label),
    x = 450,
    y = 50,
    inherit.aes = FALSE,
    size = 3,
    fontface = "bold"
  ) +
  coord_fixed(ratio = 1, clip = "on") +
  facet_wrap(Model ~ ., nrow = 1, scales = "fixed") +
  labs(
    x = "Registered CO2 emissions (MT)",
    y = "Estimated CO2 emissions using our model (MT)"
  ) +
  scale_x_continuous(limits = c(0, 600)) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    axis.text.y = element_text(size = 7),
    axis.title = element_text(size = 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.ticks.length = unit(0.25, "cm"),
    panel.spacing = unit(1.5, "lines"),
    strip.placement = "outside",
    legend.position = "none", # Remove legend
  )
```

### MRV validation

```{r}
#| label: fig-mrv-performance
#| fig-cap: Comparison of published MRV emissions and simulated CO$_{2}$ emissions. The sample includes 2,276 observations from yearly aggregated vessel trips where hours at sea differ by no more than 15\% between our data and the MRV dataset. The solid red line shows the linear regression fit, and the dashed line represents the 1:1 line.}.
#| fig-width: 8
#| fig-height: 5

## Emissions validation ----
validation_performance_metrics <- yardstick::metric_set(
  yardstick::rsq,
  yardstick::rsq_trad
)

# EU dataset
eu_validation_data_updated <- mrv_data_validation |>
  dplyr::filter(total_time_spent_at_sea_hours != 0) |>
  dplyr::mutate(
    co2_trip_emissions_eu = co2_emissions_from_all_voyages_between_ports_under_a_ms_jurisdiction_m_tonnes +
      co2_emissions_from_all_voyages_which_departed_from_ports_under_a_ms_jurisdiction_m_tonnes +
      co2_emissions_from_all_voyages_to_ports_under_a_ms_jurisdiction_m_tonnes
  ) |>
  dplyr::rename(
    total_time_spent_at_sea_hours_eu = total_time_spent_at_sea_hours
  )
# Emission results by trip
eu_validation_trip_updated <- trip_emissions_for_mrv_validation |>
  dplyr::filter(total_time_spent_at_sea_hours != 0) |>
  dplyr::rename(
    co2_trip_emissions_gfw = total_emissions_co2_mt,
    total_time_spent_at_sea_hours_gfw = total_time_spent_at_sea_hours
  )

find_duplicates <- function(data, key_columns) {
  duplicate_indices <- duplicated(data[key_columns]) |
    duplicated(data[key_columns], fromLast = TRUE)
  duplicates <- data[duplicate_indices, ]
  return(duplicates)
}

aggregate_duplicates <- function(data, group_by_cols, sum_cols, concat_cols) {
  aggregated_data <- data |>
    dplyr::group_by(dplyr::across(dplyr::all_of(group_by_cols))) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(sum_cols), sum, .names = "{.col}"),
      dplyr::across(
        dplyr::all_of(concat_cols),
        ~ paste(unique(.x), collapse = "-"),
        .names = "{.col}"
      ),
      .groups = 'drop'
    )

  return(aggregated_data)
}

group_by_columns <- c("imo_number", "year")
sum_columns <- c(
  "total_time_spent_at_sea_hours_gfw",
  "total_distance_nm",
  "co2_trip_emissions_gfw"
)

concat_columns <- c("ssvid")

eu_validation_trip_updated <- aggregate_duplicates(
  eu_validation_trip_updated,
  group_by_columns,
  sum_columns,
  concat_columns
)

# Data selection and filtering
merged_df <- merge(
  eu_validation_data_updated,
  eu_validation_trip_updated,
  by.x = c("imo_number", "reporting_period"),
  by.y = c("imo_number", "year")
)

merged_df$diff <- abs(
  merged_df$total_time_spent_at_sea_hours_eu -
    merged_df$total_time_spent_at_sea_hours_gfw
)
merged_df$max <- pmax(
  merged_df$total_time_spent_at_sea_hours_eu,
  merged_df$total_time_spent_at_sea_hours_gfw
)

# Function to filter data and calculate performance metrics based on a threshold
calculate_metrics <- function(threshold, merged_df) {
  filtered_hours_df <- subset(merged_df, diff / max < threshold)
  filtered_df <- filtered_hours_df |>
    dplyr::select(co2_trip_emissions_eu, co2_trip_emissions_gfw)

  # Calculate performance metrics
  comparison_results <- filtered_df |>
    tidyr::pivot_longer(
      -co2_trip_emissions_eu,
      names_to = "model",
      values_to = "estimate"
    ) |>
    dplyr::group_by(model) |>
    validation_performance_metrics(
      truth = co2_trip_emissions_eu,
      estimate = estimate
    ) |>
    dplyr::select(-.estimator) |>
    tidyr::pivot_wider(names_from = .metric, values_from = .estimate)

  # Return results with added threshold information
  comparison_results <- comparison_results |>
    dplyr::mutate(threshold = threshold, n_observations = nrow(filtered_df))

  return(comparison_results)
}

# Range of thresholds to test
thresholds <- seq(0.05, 0.4, 0.05)
# Apply calculate_metrics function over range of thresholds
performance_results <- purrr::map_df(
  thresholds,
  ~ calculate_metrics(.x, merged_df)
)

best_performance <- performance_results |>
  dplyr::filter(rsq_trad == max(rsq_trad))

# Select data with threshold of 0.15 the one showing best performance
filtered_hours_df <- subset(merged_df, diff / max < best_performance$threshold)
filtered_df <- filtered_hours_df |>
  dplyr::select(co2_trip_emissions_eu, co2_trip_emissions_gfw)

ggplot(
  filtered_df,
  aes(
    x = co2_trip_emissions_eu,
    y = co2_trip_emissions_gfw
  )
) +
  geom_point(
    size = 1.5,
    alpha = 0.5,
    color = "#104E8B"
  ) +
  geom_smooth(
    method = "lm",
    linewidth = 0.5,
    color = "red",
    linetype = "solid",
    se = FALSE
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  annotate(
    geom = "text",
    label = paste0(
      "Traditional R² = ",
      format(best_performance$rsq_trad, digits = 2)
    ),
    x = 130000,
    y = 10000,
    size = 3,
    fontface = "bold"
  ) +
  coord_fixed(ratio = 1, clip = "on") +
  scale_x_continuous(limits = c(0, max(filtered_df)), labels = scales::comma) +
  scale_y_continuous(limits = c(0, max(filtered_df)), labels = scales::comma) +
  labs(
    x = "Published emissions in the EU MRV database (MT)",
    y = "Estimated CO2 emissions using our model (MT)"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    axis.text.y = element_text(size = 7),
    axis.title = element_text(size = 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.ticks.length = unit(0.25, "cm"),
    panel.spacing = unit(1.5, "lines"),
    strip.placement = "outside"
  )
```

# Supplementary tables

## Dark fleet model performance metrics 

```{r}
#| results = 'asis'
all_performance_metrics_table <- all_performance_metrics |>
  dplyr::mutate(
    model = stringr::str_replace_all(model, "_", "; ") |>
      stringr::str_to_sentence(),
    test = stringr::str_replace_all(test, "_", " ") |>
      stringr::str_to_sentence()
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant),
    pollutant = tidyr::replace_na(pollutant, "-")
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::arrange(pollutant) |>
  dplyr::select(
    Model = model,
    Test = test,
    Pollutant = pollutant,
    Metric = metric,
    Estimate = estimate
  ) |>
  dplyr::mutate(
    Metric = dplyr::case_when(
      Metric == "accuracy" ~ "Accuracy",
      Metric == "f_meas" ~ "F1 score",
      Metric == "pr_auc" ~ "Precision-recall AUC",
      Metric == "roc_auc" ~ "ROC AUC",
      Metric == "precision" ~ "Precision",
      Metric == "sensitivity" ~ "Sensitivity",
      Metric == "specificity" ~ "Specificity",
      Metric ==
        "rsq_trad_backtransformed_residuals" ~ "Traditional R² (backtransformed)",
      Metric == "rsq_trad" ~ "Traditional R²",
      Metric == "mae" ~ "MAE",
      Metric ==
        "aggregate_prediction_percent_error" ~ "Percent error of total summed predictions",
      TRUE ~ Metric
    ) |>
      forcats::fct_relevel("Traditional R²")
  ) |>
  dplyr::mutate(Model = forcats::fct_relevel(Model, "Emissions; regression")) |>
  dplyr::arrange(Model, Metric) |>
  tidyr::pivot_wider(values_from = Estimate, names_from = Test) |>
  knitr::kable(
    format = "latex",
    label = "all_performance_metrics_table",
    caption = "Summary of dark fleet model performance metrics for the: 1) emissions regression model; 2) detections classification mode; and 3) detections regression model. Performance estimates for each metric are shown for both the inside S1 footprint and outside S1 footprint tests. All metrics are calculated at the unit of observation except for 'Percent error of total summed predictions', which is the percentage error in the total sum of all predictions relative to the total sum of all truth values.",
    digits = 2
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

all_performance_metrics_table

writeLines(
  all_performance_metrics_table,
  here::here("tables/all_performance_metrics_table.tex")
)
```

## Summary of 2017 and 2024 emissions, by pollutant and data source
  
```{r}
total_percent_change_by_fleet_tbl <- total_percent_change_by_fleet |>
  dplyr::mutate(across(
    c(`2017`, `2024`),
    ~ signif(., 3) |>
      prettyNum(big.mark = ",")
  )) |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::mutate(
    dplyr::across(
      c(percent_total_2024, percent_change),
      ~ scales::percent(., accuracy = 1)
    )
  ) |>
  dplyr::arrange(pollutant) |>
  dplyr::rename(
    `Data source` = fleet,
    Pollutant = pollutant,
    `2017 emissions (MT)` = `2017`,
    `2024 emissions (MT)` = `2024`,
    `Percent 2024 total` = percent_total_2024,
    `Percent change from 2017 to 2024` = percent_change
  ) |>
  knitr::kable(
    format = "latex",
    label = "total_percent_change_by_fleet",
    caption = "Summary of 2017 and 2024 emissions, by pollutant and data source."
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

writeLines(
  total_percent_change_by_fleet_tbl,
  here::here("tables/total_percent_change_by_fleet.tex")
)
```

## Underestimation of 2024 emissions using only AIS data, and overestimation of 2017 to 2024 emissions trend using only AIS data, by pollutant.

```{r}
pollutant_ais_underestimation_overestimation_summary_tbl <- pollutant_ais_underestimation_overestimation_summary |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::arrange(pollutant) |>
  dplyr::rename(
    Pollutant = pollutant,
    "AIS-only underestimation of 2024 emissions" = ais_underestimation_2024,
    "AIS-only overestimation of 2017-2024 emissions trend" = trend_overestimation_2017_2024
  ) |>
  knitr::kable(
    format = "latex",
    label = "pollutant_ais_underestimation_overestimation_summary",
    caption = "Underestimation of 2024 emissions using only AIS data, and overestimation of 2017 to 2024 emissions trend using only AIS data, by pollutant."
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

writeLines(
  pollutant_ais_underestimation_overestimation_summary_tbl,
  here::here(
    "tables/pollutant_ais_underestimation_overestimation_summary.tex"
  )
)
```

### Feature importance

```{r}
#| label: fig-feature-importance
#| fig-cap: Feature importance plots for the three final model fit random forests
#| fig-height: 9
#| fig-width: 7

all_varimp_data |>
  dplyr::left_join(
    data_sources,
    by = c("model_feature" = "Model feature shorthand")
  ) |>
  dplyr::rename(model_feature_long = `Model feature`) |>
  dplyr::mutate(
    model_feature_long = dplyr::case_when(
      stringr::str_detect(
        model_feature,
        "length_size_bin_"
      ) ~ glue::glue(
        "S1 detections per km2 (length bin {stringr::str_remove(model_feature, 'length_size_bin_')})"
      ),
      TRUE ~ model_feature_long
    )
  ) |>
  dplyr::filter(!is.na(model_feature_long)) |>
  dplyr::mutate(
    model = stringr::str_replace_all(model, "_", " ") |>
      stringr::str_to_sentence() |>
      forcats::fct_relevel(
        "Detections classification",
        "Detections regression",
        "Emissions regression"
      ),
  ) |>
  ggplot(aes(
    x = importance,
    y = tidytext::reorder_within(model_feature_long, importance, model)
  )) +
  geom_bar(stat = "identity") +
  labs(
    x = "Feature importance score (Gini impurtity corrected)",
    y = "",
  ) +
  theme_plot() +
  tidytext::scale_y_reordered() +
  facet_wrap(model ~ ., ncol = 1, scales = "free")
```

### Classification model performance

```{r}
#| label: fig-pr-curves
#| fig-cap: Precision recall curves for the detection classification model, shown for the inside S1 footprint and outside S1 footprint performance tests
#| fig-height: 3

performance_detections_cls_roc_pr_curves |>
  tidyr::pivot_wider(names_from = "metric", values_from = "estimate") |>
  dplyr::mutate(
    test = stringr::str_replace_all(test, "_", " ") |> stringr::str_to_title()
  ) |>
  ggplot(aes(x = recall, y = precision)) +
  geom_line() +
  coord_equal() +
  facet_wrap(test ~ ., ) +
  labs(
    x = "Recall",
    y = "Precision",
  ) +
  theme_plot()
```

```{r}
#| label: fig-roc-curves
#| fig-cap: ROC curves for the detection classification model, shown for the inside S1 footprint and outside S1 footprint performance tests
#| fig-height: 3

performance_detections_cls_roc_pr_curves |>
  tidyr::pivot_wider(names_from = "metric", values_from = "estimate") |>
  dplyr::mutate(
    test = stringr::str_replace_all(test, "_", " ") |> stringr::str_to_title()
  ) |>
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_line() +
  coord_equal() +
  facet_wrap(test ~ ., ) +
  labs(
    x = "False positive rate (1 - specificity)",
    y = "True positive rate (sensitivity)",
  ) +
  theme_plot() +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1))
```


```{r}
#| label: fig-conf-mat
#| fig-cap: Confusion matricies for the detection classification model, shown for the inside S1 footprint and outside S1 footprint performance tests
#| fig-height: 3

performance_detections_cls_conf_mat |>
  dplyr::mutate(
    test = stringr::str_replace_all(test, "_", " ") |> stringr::str_to_title()
  ) |>
  dplyr::mutate(prediction = forcats::fct_rev(as.factor(prediction))) |>
  ggplot(aes(x = truth, y = prediction)) +
  geom_tile(aes(fill = count), color = "black") +
  geom_label(aes(label = prettyNum(count, big.mark = ",")), vjust = 1) +
  scale_fill_gradient(low = "gray95", high = "gray10", guide = "none") +
  facet_wrap(test ~ .) +
  coord_equal() +
  labs(x = "Truth", y = "Prediction") +
  theme_minimal()
```


### Dynamic AIS - annual emissions and vessels by AIS receiver type

```{r}
#| label: fig-annual-emissions-by-ais-receiver-type
#| fig-cap: Annual total CO2 emissions (metric tonnes) by the AIS-broadcasting fleet, disaggregated by AIS receiver type (satellite, terrestrial, and dynamic) and total aggregated across receiver types. A dashed vertical line is shown for 2022, the first year in which dynamic AIS was widely adopted.
# | #| fig-width: 4
#| fig-height: 7
annual_global_emissions_by_receiver_type <- annual_global_emissions_by_receiver_type |>
  # Don't include unknown receiver type
  dplyr::filter(receiver_type != "unknown") |>
  # Fill in missing values with 0s
  tidyr::complete(
    receiver_type,
    tidyr::nesting(year),
    fill = list(emissions_co2_mt = 0, n_distinct_ssvid = 0, number_pings = 0)
  ) |>
  dplyr::mutate(receiver_type = stringr::str_to_sentence(receiver_type)) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025)

annual_global_emissions_by_receiver_type |>
  dplyr::rename(
    `CO[2]~emissions~(MT)` = emissions_co2_mt,
    `Number~unique~vessels` = n_distinct_ssvid,
    `Number~AIS~messages` = number_pings,
  ) |>
  tidyr::pivot_longer(cols = -c(year, receiver_type)) |>
  ggplot(aes(x = year, y = value, color = receiver_type)) +
  facet_wrap(name ~ ., scales = "free_y", ncol = 1, labeller = label_parsed) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  ) +
  scale_x_continuous(
    breaks = unique(annual_global_emissions_by_receiver_type$year)
  ) +
  scale_color_brewer(
    type = "qual",
    palette = "Set2",
    guide = guide_legend(reverse = TRUE)
  ) +
  theme_plot() +
  labs(x = "", y = "", color = "AIS receiver type") +
  theme(panel.grid.minor.x = element_blank()) +
  geom_vline(xintercept = 2022, linetype = 2)
```

### Dynamic AIS - annual emissions by AIS receiver type for top flags

```{r}
#| label: fig-annual-emissions-by-ais-receiver-type-top-flags
#| fig-cap: Annual total CO2 emissions (metric tonnes) by the AIS-broadcasting fleet, disaggregated by AIS receiver type (satellite, terrestrial, and dynamic) and total aggregated across receiver types, for the top 10 flags with the most emissions in 2024. A dashed vertical line is shown for 2022, the first year in which dynamic AIS was widely adopted.
#| fig-width: 7.5
#| fig-height: 6
annual_global_emissions_by_receiver_type_and_flag <- annual_global_emissions_by_receiver_type_and_flag |>
  # Add column name for country name based on flag

  dplyr::mutate(
    flag = countrycode::countrycode(
      flag,
      origin = "iso3c",
      destination = "country.name"
    )
  ) |>
  # Don't include unknown receiver type
  dplyr::filter(receiver_type != "unknown") |>
  dplyr::mutate(
    receiver_type = stringr::str_to_sentence(receiver_type) |>
      forcats::fct_relevel("Satellite", "Terrestrial", "Dynamic")
  ) |>
  # Fill in missing values with 0s
  tidyr::complete(
    receiver_type,
    tidyr::nesting(year, flag),
    fill = list(emissions_co2_mt = 0, n_distinct_ssvid = 0)
  ) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025)

top_n_flags_2024 <- annual_global_emissions_by_receiver_type_and_flag |>
  dplyr::filter(year == 2024, receiver_type == "Total") |>
  dplyr::slice_max(n = 10, order_by = emissions_co2_mt) |>
  dplyr::select(flag)

annual_global_emissions_by_receiver_type_and_flag |>
  dplyr::inner_join(top_n_flags_2024, by = "flag") |>
  dplyr::mutate(flag = forcats::fct_relevel(flag, top_n_flags_2024$flag)) |>
  ggplot(aes(x = year, y = emissions_co2_mt, color = flag)) +
  facet_grid(receiver_type ~ ., scales = "free_y") +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  scale_x_continuous(
    breaks = unique(annual_global_emissions_by_receiver_type_and_flag$year)
  ) +
  theme_plot() +
  labs(x = "", y = expression(CO[2] ~ emissions ~ (MT)), color = "Flag") +
  theme(panel.grid.minor.x = element_blank()) +
  geom_vline(xintercept = 2022, linetype = 2) +
  paletteer::scale_color_paletteer_d(palette = "rcartocolor::Safe")

```

### Dynamic AIS - map showing spatial change in emissions by AIS receiver type

```{r}
#| label: fig-co2-emissions-change-map-by-receiver-type
#| fig-cap: Spatial change in CO2 emissions by the AIS-broadcasting fleet between 2017 and 2024 (metric tonnes) by receiver type (satellite, terrestrial, and dynamic) and total across receiver types, at a 1x1 degree spatial resolution.
#| fig-width: 7.5
#| fig-height: 8
# Annual spatial CO2 emissions vt receiver type  (satellite; dynamic; or terrestrial)
annual_spatial_emissions_by_receiver_type <- annual_spatial_emissions_by_receiver_type |>
  dplyr::mutate(
    receiver_type = stringr::str_to_sentence(receiver_type) |>
      forcats::fct_relevel("Satellite", "Terrestrial", "Dynamic")
  ) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025) |>
  dplyr::mutate(
    emissions_mt_per_km2 = emissions_co2_mt / lat_to_area(lat_bin)
  ) |>
  dplyr::select(-emissions_co2_mt)

annual_spatial_emissions_by_receiver_type |>
  dplyr::filter(year %in% c(analysis_start_year, analysis_end_year)) |>
  tidyr::pivot_wider(names_from = year, values_from = emissions_mt_per_km2) |>
  dplyr::mutate(across(c(`2017`, `2024`), ~ tidyr::replace_na(., 0))) |>
  dplyr::mutate(emissions_mt_per_km2 = `2024` - `2017`) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  ggspatial::geom_spatial_tile(aes(
    x = lon_bin + 0.5,
    y = lat_bin + 0.5,
    fill = emissions_mt_per_km2,
    color = emissions_mt_per_km2
  )) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  facet_wrap(receiver_type ~ ., ncol = 1) +
  scico::scale_fill_scico(
    expression(atop(
      Change ~ of ~ CO[2] ~ emissions,
      from ~ 2017 ~ to ~ 2024 ~ (MT / km^2)
    )),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  scico::scale_color_scico(
    expression(atop(
      Change ~ of ~ CO[2] ~ emissions,
      from ~ 2017 ~ to ~ 2024 ~ (MT / km^2)
    )),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme_void() +
  theme(legend.position = "right", legend.direction = "vertical") +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black",
      barwidth = unit(1, "cm"),
      barheight = unit(5, "cm")
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  coord_sf(crs = map_projection)
```

## MRV performance validation table

```{r}
performance_results |>
  dplyr::mutate(
    rsq = signif(rsq, 2),
    rsq_trad = signif(rsq_trad, 2),
    n_observations = prettyNum(n_observations, big.mark = ",")
  ) |>
  dplyr::select(
    Tolerance = threshold,
    `Number of vessels` = n_observations,
    "R$^{2}$" = rsq,
    "Traditional R$^{2}$" = rsq_trad
  ) |>
  knitr::kable(
    format = "latex",
    label = "tab-performance-tolerance-table",
    caption = "Performance summary across different thresholds for the hours at sea tolerance between datasets. Results are shown using the R$^{2}$ (measuring linear association between observed and predicted values) and traditional R$^{2}$ (proportion of variance explained compared to the mean-only baseline).",
    escape = FALSE
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position") |>
  writeLines(
    here::here("tables/mrv_performance_results.tex")
  )
```


# Non-broadcasting model data sources


```{r}
data_sources_table <- data_sources |>
  dplyr::select(-`Model feature shorthand`) |>
  dplyr::mutate(
    dplyr::across(
      dplyr::everything(),
      ~ stringr::str_replace_all(., "²", "$^2$")
    )
  ) |>
  knitr::kable(
    format = "latex",
    escape = FALSE,
    booktabs = TRUE,
    longtable = FALSE,
    caption = "Data sources for all included model features, including which random forest model uses the features (detections classification; detections regression; and/or emissions regression), the units, feature type, whether it varies spatially, whether it varies temporally, and the source name and reference citation.",
    label = "data_sources"
  ) |>
  kableExtra::kable_styling(
    latex_options = c("scale_down", "hold_position"),
    font_size = 7
  ) |>
  kableExtra::landscape()

writeLines(data_sources_table, here::here("tables/data_sources.tex"))
```


```{r}
#| label: fig-offshore-outside-footprint-training-testing-map
#| fig-cap: Maps of pixels that were used in the training and testing datasets for the outside S1 performance tests for each model. Pixels <=100km from the shore were used for training; pixels >100km from shore were used for testing. The exact training/testing split for each of the 3 models is shown.
#| fig-width: 7.5
#| fig-height: 7.5
pixels_for_offshore_training_testing_split |>
  dplyr::mutate(type = ifelse(offshore_pixel, "Testing", "Training")) |>
  dplyr::mutate(
    model = stringr::str_replace_all(model, "_", "; ") |>
      stringr::str_to_sentence()
  ) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      color = type,
      fill = type
    )
  ) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  paletteer::scale_color_paletteer_d(
    name = "Dataset",
    palette = "nbapalettes::pacers_classic",
    direction = -1
  ) +
  paletteer::scale_fill_paletteer_d(
    name = "Dataset",
    palette = "nbapalettes::pacers_classic",
    direction = -1
  ) +
  labs(x = "", y = "") +
  theme_void() +
  theme_plot() +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  facet_wrap(model ~ ., ncol = 1) +
  guides(fill = guide_legend(reverse = TRUE)) +
  guides(color = guide_legend(reverse = TRUE))
```

```{r}
#| label: fig-map-fraction-months-imaged
#| fig-cap: Map showing the fraction of months from 2017 through 2024 that were imaged by S1 for each 1x1 degree pixel.
#| fig-width: 7.5
#| fig-height: 5
number_s1_imaged_months_by_pixel |>
  dplyr::mutate(
    fraction_months_imaged = number_months_imaged / number_months
  ) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      color = fraction_months_imaged,
      fill = fraction_months_imaged
    )
  ) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scale_fill_viridis_c(
    name = "Percentage of months imaged by S1 between 2017 and 2024",
    labels = scales::percent,
    option = "magma"
  ) +
  scale_color_viridis_c(
    name = "Percentage of months imaged by S1 between 2017 and 2024",
    labels = scales::percent,
    option = "magma"
  ) +
  labs(x = "", y = "") +
  theme_void() +
  theme_plot() +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "white",
      barwidth = unit(15, "cm"),
      barheight = unit(1, "cm")
    ),
    color = "none"
  )
```