---
title: "Quantifying comprehensive marine vessel emissions using data fusion"
author: "emLab, UCSB and GFW"
date: last-modified
date-format: "[Last Updated on] MMMM DD, YYYY"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    fig-pos: H
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
cache: true
---

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE, cache = FALSE}
#| label: document-setup

# Set default chunk options
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  dev = "png",
  fig.path = "../figures/"
)

# Load librariers
library(ggplot2)
library(sf)

analysis_start_year <- 2016
analysis_end_year <- 2024
```

```{r}
#| label: load-data
#| cache: FALSE
targets::tar_load(c(
  "n_unique_vessels",
  "n_ais_messages",
  "annual_edgar_emissions",
  "annual_emissions_all_pollutants",
  "total_spatial_emissions_by_pollutant",
  "annual_ais_to_dark_activity_extrapolation",
  "total_monthly_emissions_by_pollutant",
  "annual_ais_co2_emissions_by_vessel_type",
  "annual_non_broadcasting_detections_emissions",
  "s1_mean_spatial_knn_ratios",
  "annual_global_emissions_by_receiver_type",
  "annual_global_emissions_by_receiver_type_and_flag",
  "annual_spatial_emissions_by_receiver_type",
  "vessel_size_info",
  "s1_ais_vessel_size_class_comparison"
))
```

```{r}
#| label: setup-graphics

# For figures and tables, order pollutants first for GHG pollutants,
# then local pollutants. For figures, we use expressions to show subscripts
pollutant_factor_levels <- c(
  "CO[2]",
  "CH[4]",
  "N[2]*O",
  "CO",
  "NO[X]",
  "SO[X]",
  "PM[2.5]",
  "PM[10]",
  "VOCs"
)

# Set pollutant levels for tables
# For tables, we don't use expressions
pollutant_factor_levels_tables <- pollutant_factor_levels |>
  stringr::str_remove_all("\\[") |>
  stringr::str_remove_all("\\]") |>
  stringr::str_remove_all("\\*]*")

# Use Okabe-Ito colorblind-friendly colors
all_data_source_color_palette <- c(
  "Marine vessels (AIS + S1)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[5]],
  "AIS + S1" = paletteer::paletteer_d("colorblindr::OkabeIto")[[5]],
  "Terrestrial transportation (EDGAR)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[6]],
  "AIS (broadcasting)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[3]],
  "S1 (non-broadcasting)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[
    7
  ]],
  "EDGAR" = paletteer::paletteer_d("colorblindr::OkabeIto")[[8]]
)

# Set consistent theme for all plots
theme_plot <- function() {
  theme_minimal() %+replace%
    theme(
      axis.title.x = element_text(face = "bold"),
      axis.title.y = element_text(angle = 90, face = "bold", vjust = 3),
      strip.text.x = element_text(angle = 0, face = "bold"),
      strip.text.y = element_text(angle = 0, face = "bold")
    )
}

# Set map projection for all map figures
map_projection <- "+proj=eqearth"

# Create bounding box of globe, for making black map backgrounds
lats <- c(90:-90, -90:90, 90)
longs <- c(rep(c(180, -180), each = 181), 180)
world_bbox_sf <-
  list(cbind(longs, lats)) %>%
  sf::st_polygon() %>%
  sf::st_sfc(
    # create sf geometry list column
    crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
  ) %>%
  sf::st_sf() %>%
  sf::st_transform(map_projection)

land <- rnaturalearth::ne_countries(returnclass = "sf") |>
  dplyr::select(geometry)


# Convert lat/lon pixel area to square km
# https://stackoverflow.com/a/1253545
# Need function to convert degrees to radians
# https://stackoverflow.com/q/32370485
deg2rad <- function(deg) {
  (deg * pi) / (180)
}

# Given a latitude, what is a pixel's area?
# https://stackoverflow.com/a/1253545
lat_to_area <- function(lat) {
  110.574 * (111.320 * cos(deg2rad(lat)))
}

```

```{r}
#| label: wrangle-edgar-data
annual_edgar_tranportation_emissions <- annual_edgar_emissions |>
  dplyr::select(
    -c(
      IPCC_annex,
      C_group_IM24_sh,
      Country_code_A3,
      Name,
      ipcc_code_2006_for_standard_report,
      Substance,
      fossil_bio
    )
  ) |>
  dplyr::rename(sector = ipcc_code_2006_for_standard_report_name) |>
  tidyr::pivot_longer(
    cols = -c(sector),
    names_to = "year",
    values_to = "emissions_co2_kt"
  ) |>
  dplyr::mutate(
    year = stringr::str_remove_all(year, "Y_") |>
      as.numeric(),
    emissions_co2_mt = emissions_co2_kt * 1e3
  ) |>
  dplyr::select(-emissions_co2_kt) |>
  dplyr::mutate(
    land_based_tranpsportation = ifelse(
      sector %in%
        c(
          "Civil Aviation",
          "Road Transportation no resuspension",
          "Other Transportation",
          "Railways"
        ),
      TRUE,
      FALSE
    )
  ) |>
  # Only need years that correspond to our dataset
  dplyr::filter(year >= analysis_start_year, year <= analysis_end_year)

# Fraction of global 2023 CO2 emissions that came from transportation sectors
fraction_co2_from_transportation <- annual_edgar_tranportation_emissions |>
  dplyr::mutate(
    transportation = ifelse(
      land_based_tranpsportation | sector == "Water-borne Navigation",
      TRUE,
      FALSE
    )
  ) |>
  dplyr::filter(year == 2023) |>
  dplyr::group_by(transportation) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(
    names_from = "transportation",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::mutate(fraction_from_transportation = `TRUE` / (`TRUE` + `FALSE`)) |>
  dplyr::pull(fraction_from_transportation) |>
  scales::percent()
```

```{r}
#| label: wrangle-gfw-time-series-emissions-data

annual_emissions_all_pollutants <- annual_emissions_all_pollutants |>
  tidyr::pivot_longer(
    -c(year, fishing),
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    fleet = ifelse(
      stringr::str_detect(pollutant, "dark"),
      "Non-broadcasting",
      "AIS-broadcasting"
    )
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_mt") |>
      stringr::str_remove_all("_dark")
  )

annual_emissions_all_pollutants_with_total <- annual_emissions_all_pollutants |>
  # Add total, that aggregates across broadcasting and non-broadcasting
  dplyr::bind_rows(
    annual_emissions_all_pollutants |>
      dplyr::group_by(year, fishing, pollutant) |>
      dplyr::summarize(emissions_mt = sum(emissions_mt)) |>
      dplyr::ungroup() |>
      dplyr::mutate(fleet = "Total")
  )

# Aggregate emissions across fishing/non-fishing
annual_emissions <- annual_emissions_all_pollutants_with_total |>
  dplyr::group_by(year, fleet, pollutant) |>
  dplyr::summarise(emissions_mt = sum(emissions_mt)) |>
  dplyr::ungroup()

# Calculate the total percent change in emissions by fleet
total_percent_change_by_fleet <- annual_emissions |>
  dplyr::filter(year %in% c(2016, 2024), fleet != "Total") |>
  tidyr::pivot_wider(names_from = year, values_from = emissions_mt) |>
  dplyr::arrange(pollutant) |>
  dplyr::group_by(pollutant) |>
  dplyr::mutate(percent_total_2024 = `2024` / sum(`2024`)) |>
  dplyr::ungroup() %>%
  split(.[, c("pollutant")]) %>%
  purrr::map_df(janitor::adorn_totals, fill = NA) |>
  tidyr::fill(pollutant, .direction = "down") |>
  # janitor::adorn_totals("row") |>
  dplyr::mutate(percent_total_2024 = scales::percent(percent_total_2024)) |>
  dplyr::mutate(percent_change = 100 * (`2024` - `2016`) / `2016`) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS (broadcasting)",
      TRUE ~ "S1 (non-broadcasting)"
    )
  )
```

```{r}
# For each pollutant, estimate how much emissions are underestimated by just using AIS
# As compared to using AIS + S1
ais_scale_underestimation_2024 <- total_percent_change_by_fleet |>
  dplyr::filter(fleet != "S1 (non-broadcasting)") |>
  dplyr::select(fleet, pollutant, `2024`) |>
  tidyr::pivot_wider(names_from = "fleet", values_from = `2024`) |>
  dplyr::mutate(
    ais_underestimation_2024 = (1 - `AIS (broadcasting)` / `AIS + S1`) |>
      scales::percent()
  ) |>
  dplyr::rename(
    emissions_mt_2024_ais = `AIS (broadcasting)`,
    emissions_mt_2024_ais_s1 = `AIS + S1`
  )

# For each pollutant, estimate how much the 2016->2024 trend in emissions is underestimated by just using AIS
# As compared to using AIS + S1
ais_trend_overestimation_2016_2024 <- total_percent_change_by_fleet |>
  dplyr::filter(fleet != "S1 (non-broadcasting)") |>
  dplyr::select(fleet, pollutant, percent_change) |>
  tidyr::pivot_wider(names_from = "fleet", values_from = percent_change) |>
  dplyr::mutate(
    trend_overestimation_2016_2024 = (`AIS (broadcasting)` /
      `AIS + S1` -
      1) |>
      scales::percent()
  ) |>
  dplyr::rename(
    trend_2016_2024_ais = `AIS (broadcasting)`,
    trend_2016_2024_ais_s1 = `AIS + S1`
  )

# Put all this information together
pollutant_ais_underestimation_overestimation_summary <- ais_scale_underestimation_2024 |>
  dplyr::left_join(ais_trend_overestimation_2016_2024, by = "pollutant") |>
  dplyr::select(
    pollutant,
    ais_underestimation_2024,
    trend_overestimation_2016_2024
  )
```

```{r}
#| label: wrangle-gfw-spatials-emissions-data
# Total spatial emissions by pollutant for 2016 and 2024
total_spatial_emissions_by_pollutant <- total_spatial_emissions_by_pollutant |>
  tidyr::pivot_longer(
    -c(lat_bin, lon_bin, year),
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_total_mt") |>
      stringr::str_to_upper() |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(emissions_mt_per_km2 = emissions_mt / lat_to_area(lat_bin)) |>
  dplyr::select(-emissions_mt) |>
  tidyr::pivot_wider(
    names_from = "year",
    values_from = "emissions_mt_per_km2",
    names_prefix = "emissions_mt_per_km2_"
  )
```


```{r}
#| label: dark-fleet-inferred-stats
percent_dark_emissions_inferred <- annual_non_broadcasting_detections_emissions |>
  dplyr::group_by(ratio_inferred) |>
  dplyr::summarize(emissions_co2_dark_mt = sum(emissions_co2_dark_mt)) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(
    names_from = "ratio_inferred",
    values_from = "emissions_co2_dark_mt"
  ) |>
  dplyr::mutate(percent_inferred = scales::percent(`TRUE` / (`TRUE` + `FALSE`)))

percent_dark_emissions_inferred_by_type <- annual_non_broadcasting_detections_emissions |>
  dplyr::group_by(fishing, ratio_inferred) |>
  dplyr::summarize(emissions_co2_dark_mt = sum(emissions_co2_dark_mt)) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(
    names_from = "ratio_inferred",
    values_from = "emissions_co2_dark_mt"
  ) |>
  dplyr::mutate(percent_inferred = scales::percent(`TRUE` / (`TRUE` + `FALSE`)))
```

# In-line manuscript statistics

- The transportation sector is a major source of global greenhouse gas (GHG) and local air pollution emissions, responsible for `r fraction_co2_from_transportation` of all CO$_{2}$ emissions globally in 2023 

- For over `r scales::comma(signif(n_unique_vessels$n_unique_vessels,2))` AIS-broadcasting vessels, we quantify point-estimate emissions for each vessel using precise latitude/longitude/timestamp information for over `r scales::comma(signif(n_ais_messages$n_billion_ais_messages,3))` billion individual AIS messages


- Because many vessels do not broadcast AIS, current approaches solely based on AIS \textit{underestimate} current global emissions by `r pollutant_ais_underestimation_overestimation_summary |> dplyr::filter(pollutant == "co2") |> dplyr::pull(ais_underestimation_2024)` (in 2024), and \textit{overestimate} the increasing trend by `r pollutant_ais_underestimation_overestimation_summary |> dplyr::filter(pollutant == "co2") |> dplyr::pull(trend_overestimation_2016_2024)` (from 2016 to 2024). 

- Across the entire 2016 to 2024 time series, `r percent_dark_emissions_inferred$percent_inferred` of non-broadcasting emissions were estimated using inferred ratios from the KNN approach (`r percent_dark_emissions_inferred_by_type |> dplyr::filter(!fishing) |> dplyr::pull(percent_inferred)` for non-fishing vessels, and `r percent_dark_emissions_inferred_by_type |> dplyr::filter(fishing) |> dplyr::pull(percent_inferred)` for fishing vessels).

# Figures


## Figure 1

### Figure 1, panel A

```{r}
total_annual_emissions_by_fleet_fig <- annual_emissions |>
  dplyr::filter(pollutant == "co2") |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS (broadcasting)",
      TRUE ~ "S1 (non-broadcasting)"
    )
  ) |>
  ggplot(aes(x = year, y = emissions_mt, color = fleet)) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "B", scale = 1e-9),
    limits = c(0, 1.5e9),
    breaks = seq(0, 1.5e9, 0.25e9)
  ) +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(2016, 2024)
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS + S1", "AIS (broadcasting)", "S1 (non-broadcasting)")
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(Annual ~ CO[2] ~ emissions ~ (MT))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.65, 0.4),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  )
```

### Figure 2, panel B

```{r}
# Get 2016 values for time-series normalization
annual_co2_emissions_starting <- annual_emissions |>
  dplyr::filter(pollutant == "co2", year == 2016) |>
  dplyr::mutate(emissions_co2_mt_starting = emissions_mt) |>
  dplyr::select(-c(year, emissions_mt))

normalized_annual_emissions_by_fleet_fig <- annual_emissions |>
  dplyr::filter(pollutant == "co2") |>
  dplyr::left_join(annual_co2_emissions_starting, by = "fleet") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_mt / emissions_co2_mt_starting - 1
  ) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS (broadcasting)",
      TRUE ~ "S1 (non-broadcasting)"
    )
  ) |>
  ggplot(aes(x = year, y = emissions_co2_mt_normalized, color = fleet)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(-0.2, 0.35),
    breaks = seq(-.2, .35, .05)
  ) +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(2016, 2024)
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS (broadcasting)", "AIS + S1", "S1 (non-broadcasting)")
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of ~ CO[2] ~ emissions,
      from ~ 2016 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.35, 0.85),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  )
```

### Figure 1, panel A

```{r}
pollutant_map_data_co2 <- total_spatial_emissions_by_pollutant |>
  dplyr::filter(pollutant %in% c("CO2")) |>
  dplyr::mutate(
    emissions_mt_per_km2_delta_abs = emissions_mt_per_km2_2024 -
      emissions_mt_per_km2_2016
  ) |>
  dplyr::select(
    lon_bin,
    lat_bin,
    emissions_mt_per_km2_delta_abs,
    emissions_mt_per_km2_2024
  ) |>
  tidyr::pivot_longer(
    cols = c(emissions_mt_per_km2_delta_abs, emissions_mt_per_km2_2024)
  )

co2_map <- pollutant_map_data_co2 |>
  dplyr::mutate(
    name = ifelse(
      name == "emissions_mt_per_km2_delta_abs",
      "Change in emissions from 2016 to 2024",
      "2024 emissions"
    )
  ) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      color = value,
      fill = value
    )
  ) +
  facet_wrap(name ~ ., ncol = 1) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scico::scale_fill_scico(
    name = expression(CO[2] ~ emissions ~ (MT / km^2)),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  scico::scale_color_scico(
    name = expression(CO[2] ~ emissions ~ (MT / km^2)),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme_void() +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black"
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(2, "cm")
  )
```


### Figure 1, combined

```{r}
#| label: fig-emissions-by-data-source-and-maps
#| #| fig-cap: (A) Annual marine vessel CO2 emissions, with each line colored based on the data source. (B) Relative change in annual marine vessel CO2 emissions from a 2016 baseline, with each line colored based on the data source. (C) Spatial map of CO2 emissions in 2024 (top) and change in CO2 emissions between 2016 and 2024 (bottom), aggregated across AIS-broadcasting vessels and non-broadcasting vessels, and shown at a 1x1 degree spatial resolution; emissions are area-normalized for each pixel.
#| fig-width: 11
#| fig-height: 7
cowplot::plot_grid(
  cowplot::plot_grid(
    total_annual_emissions_by_fleet_fig,
    normalized_annual_emissions_by_fleet_fig,
    ncol = 1,
    labels = c("A", "B")
  ),
  co2_map,
  ncol = 2,
  labels = c("", "B"),
  rel_widths = c(0.4, 0.6),
  align = "vh"
)
```

### Figure 2

```{r}
#| label: fig-emissions-marine-terrestrial
#| fig-cap: Comparison of emissions data between marine transportation emissions from GFW and terrestrial transportation emissions from EDGAR.
#| fig-width: 7
#| fig-height: 6
# Calculate annual terrestrial transportation emissions from EDGAR
land_based_transportation_emissions_edgar <- annual_edgar_tranportation_emissions |>
  dplyr::filter(land_based_tranpsportation) |>
  dplyr::group_by(year) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::mutate(data_source = "Terrestrial transportation (EDGAR)")


all_tranportation_data <- land_based_transportation_emissions_edgar |>
  dplyr::bind_rows(
    annual_emissions |>
      dplyr::filter(pollutant == "co2", fleet == "Total") |>
      dplyr::rename(emissions_co2_mt = emissions_mt) |>
      dplyr::mutate(data_source = "Marine vessels (AIS + S1)") |>
      dplyr::select(-c(fleet, pollutant))
  )


# Get 2016 values for time-series normalization
all_tranportation_data_starting <- all_tranportation_data |>
  dplyr::filter(year == 2016) |>
  dplyr::mutate(emissions_co2_mt_starting = emissions_co2_mt) |>
  dplyr::select(-c(year, emissions_co2_mt))

all_tranportation_data_normalized <- all_tranportation_data |>
  dplyr::inner_join(all_tranportation_data_starting, by = "data_source") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_co2_mt /
      emissions_co2_mt_starting -
      1
  )

# Summarize total emissions by data source for 2023, for putting on figure
emissions_comparison_figure <- all_tranportation_data |>
  ggplot(aes(x = year, y = emissions_co2_mt, color = data_source)) +
  geom_line(linewidth = 1.05) +
  scale_x_continuous(
    breaks = unique(all_tranportation_data_normalized$year),
    limits = c(2016, 2023)
  ) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "B", scale = 1e-9),
    limits = c(0, NA)
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(
      atop(Annual ~ transportation, ~ CO[2] ~ emissions ~ (MT))
    ),
    color = "Data source"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.65, 0.5),
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", color = "gray")
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c(
      "Terrestrial transportation (EDGAR)",
      "Marine vessels (AIS + S1)"
    )
  )

emissions_comparison_figure_normalized <- all_tranportation_data_normalized |>
  dplyr::filter(
    data_source %in%
      c("Terrestrial transportation (EDGAR)", "Marine vessels (AIS + S1)")
  ) |>
  ggplot(aes(x = year, y = emissions_co2_mt_normalized, color = data_source)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(linewidth = 1.1) +
  scale_x_continuous(
    breaks = unique(all_tranportation_data_normalized$year),
    limits = c(2016, 2023)
  ) +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(-0.15, 0.25),
    breaks = seq(-.15, .25, .05)
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of,
      CO[2] ~ emissions ~ from ~ 2016 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.25, 0.875),
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", color = "gray")
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c(
      "Marine vessels (AIS + S1)",
      "Terrestrial transportation (EDGAR)"
    )
  )

cowplot::plot_grid(
  emissions_comparison_figure,
  emissions_comparison_figure_normalized,
  ncol = 1,
  labels = c("A", "B")
)
```

## Figure 3

```{r}
#| label: build-decomposition-model

# For aggregate annual extrapolation dataset, calculated total emissions, kw-hours,distance, hours
annual_extrapolation_dataset <- annual_ais_to_dark_activity_extrapolation |>
  dplyr::mutate(
    emissions_co2_mt_total = emissions_co2_mt + emissions_co2_mt_dark,
    kw_hours_total = kw_hours + kw_hours_dark,
    distance_nm_total = distance_nm + distance_nm_dark,
    hours_total = hours + hours_dark
  ) |>
  dplyr::select(
    -c(
      emissions_co2_mt,
      emissions_co2_mt_dark,
      kw_hours,
      kw_hours_dark,
      distance_nm,
      distance_nm_dark,
      hours,
      hours_dark,
      avg_speed_knots
    )
  ) |>
  # Calculate average aggregate speed as total distance traveled divided by total hours
  dplyr::mutate(avg_speed_knots = distance_nm_total / hours_total)

# Using annual_extrapolation_dataset, use tidymodels to train a random forest model
# to predict emissions
set.seed(123)

# Set initial split of data
inital_split <- annual_extrapolation_dataset |>
  rsample::initial_split(prop = 0.8)

training_data <- rsample::training(inital_split)
testing_data <- rsample::testing(inital_split)

# Specify recipe for model
model_recipe <- recipes::recipe(
  emissions_co2_mt_total ~
    fishing + length_size_class_percentile + kw_hours_total + avg_speed_knots,
  data = head(rsample::training(inital_split))
) |>
  recipes::step_num2factor(
    length_size_class_percentile,
    levels = as.character(unique(
      annual_extrapolation_dataset$length_size_class_percentile
    ))
  )

# Specify model type
model_type <- parsnip::rand_forest(
  trees = 1000,
  # Default of ranger is squareroot of number of model features. We have 4 features, so use a value of 2.
  mtry = 2,
  # Default of ranger is 5 for regression
  min_n = 5
) |>
  parsnip::set_engine("ranger") |>
  parsnip::set_mode("regression")

# Initialize model workflow
model_workflow <- workflows::workflow() |>
  workflows::add_recipe(model_recipe) |>
  workflows::add_model(model_type)

# Fit model on training data
workflow_fit <- parsnip::fit(model_workflow, training_data)

# Make predictions on testing data, and augment this to testing data frame
worflow_predictions <- parsnip::augment(workflow_fit, testing_data)


# Calculate performance metrics

# Multiple regression metrics
multi_metric <- yardstick::metric_set(
  yardstick::rsq,
  yardstick::mape,
  yardstick::rsq_trad
)

performance_metrics <- multi_metric(
  data = worflow_predictions,
  truth = emissions_co2_mt_total,
  estimate = .pred
) |>
  dplyr::select(-.estimator)

# Final model on all data
# Fit final model on all data
workflow_fit_final <- parsnip::fit(model_workflow, annual_extrapolation_dataset)
```

```{r}
#| label: tbl-decomposition-model-performance
#| tbl-cap: Performance metrics for random forest model that predicts annual CO2 emissions, by vessel size class and type, using hours and average speed.

performance_metrics |>
  knitr::kable(digits = 3)
```

```{r}
#| label: predict-using-decomposition-model
# Find initial/starting 2016 values of kw-hours and speed, by fishing and length_size_class_percentile, for decomposition
starting_values <- annual_extrapolation_dataset |>
  dplyr::filter(year == 2016) |>
  dplyr::select(
    fishing,
    length_size_class_percentile,
    kw_hours_total_starting = kw_hours_total,
    hours_total_starting = hours_total,
    avg_speed_knots_starting = avg_speed_knots
  )

# Add starting values to predictions dataset
predictions_dataset <- annual_extrapolation_dataset |>
  dplyr::left_join(
    starting_values,
    by = c("fishing", "length_size_class_percentile")
  )

final_model_fit <- workflows::extract_fit_parsnip(workflow_fit_final)

# Now run predictions for 4 different scenarios
final_prediction_dataset <- predictions_dataset |>
  # Add model predictions using observed kw-hours and observed speed
  dplyr::bind_cols(
    observed_kw_hours_observed_speed = predict(
      final_model_fit,
      predictions_dataset
    )$.pred
  ) |>
  # Add model predictions using observed kw-hours and constant speed (from 2016)
  dplyr::bind_cols(
    observed_kw_hours_constant_speed = predict(
      final_model_fit,
      predictions_dataset |>
        dplyr::select(-avg_speed_knots) |>
        dplyr::rename(avg_speed_knots = avg_speed_knots_starting)
    )$.pred
  ) |>
  # Add model predictions using constant kw-hours (from 2016) and observed speed
  dplyr::bind_cols(
    constant_kw_hours_observed_speed = predict(
      final_model_fit,
      predictions_dataset |>
        dplyr::select(-kw_hours_total) |>
        dplyr::rename(kw_hours_total = kw_hours_total_starting)
    )$.pred
  ) |>
  # Add model predictions using constant kw-hours (from 2016) and constant speed (from 2016)
  dplyr::bind_cols(
    constant_kw_hours_constant_speed = predict(
      final_model_fit,
      predictions_dataset |>
        dplyr::select(-avg_speed_knots) |>
        dplyr::rename(avg_speed_knots = avg_speed_knots_starting) |>
        dplyr::select(-kw_hours_total) |>
        dplyr::rename(kw_hours_total = kw_hours_total_starting)
    )$.pred
  )
```

## Figure 3

```{r}
#| label: fig-hours-speed-decomposition
#| fig-cap: Estimated decomposition of change in total CO2 emissions from 2016 baseline (aggregating across both broadcasting and non-broadcasting vessels), using the random forest model outputs of four different scenarios that decompose the relative impact of changes in total kW-hours and changes in average speed/
#| fig-width: 7.5
#| fig-height: 3
# Aggregate prediction values by year (across vessel size classes and type)
final_annual_prediction_dataset <- final_prediction_dataset |>
  dplyr::group_by(year) |>
  dplyr::summarise(across(
    c(
      observed_kw_hours_observed_speed,
      observed_kw_hours_constant_speed,
      constant_kw_hours_observed_speed,
      constant_kw_hours_constant_speed
    ),
    ~ sum(., na.rm = TRUE)
  )) |>
  dplyr::rename(
    `Observed kW-hours, observed speed` = observed_kw_hours_observed_speed,
    `Observed kW-hours, constant speed` = observed_kw_hours_constant_speed,
    `Constant kW-hours, observed speed` = constant_kw_hours_observed_speed,
    `Constant kW-hours, constant speed` = constant_kw_hours_constant_speed
  ) |>
  tidyr::pivot_longer(-year)

# We will normalize all of these predictions to the 2016 starting prediction
starting_value_from_prediction <- final_annual_prediction_dataset |>
  dplyr::filter(year == 2016) |>
  dplyr::rename(starting_value = value) |>
  dplyr::select(-year)

time_series_predicted_decomposition_df <- final_annual_prediction_dataset |>
  dplyr::inner_join(starting_value_from_prediction, by = "name") |>
  dplyr::mutate(normalized_value = value / starting_value - 1)

time_series_predicted_decomposition_df |>
  ggplot(aes(x = year, y = normalized_value, color = name)) +
  geom_line(linewidth = 1.1) +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_color_manual(
    "Prediction scenario",
    values = c(
      "coral3",
      "coral4",
      paletteer::paletteer_d("colorblindr::OkabeIto")[[2]],
      paletteer::paletteer_d("colorblindr::OkabeIto")[[5]]
    )
  ) +
  scale_x_continuous(breaks = unique(final_prediction_dataset$year)) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "",
    y = ""
  ) +
  theme_plot() +
  theme(panel.grid.minor.x = element_blank())
```

# Supplementary figures

### Relationship between length and engine power

```{r}
#| label: fig-ais-length-power-relationship
#| fig-cap: The relationship between main engine power (kilowatts) and length (meters), where each point is a specific AIS-broadcasting vessel. The relationship is disaggregated for non-fishing vessels and fishing vessels.
#| #| fig-height: 10
#| #| fig-width: 6
vessel_size_info |>
  dplyr::mutate(
    fishing = ifelse(
      fishing,
      "Fishing",
      "Non-fishing"
    ) |>
      forcats::fct_rev()
  ) |>
  ggplot(aes(x = length_m, y = main_engine_power_kw)) +
  geom_point(size = 0.1, alpha = 0.5) +
  theme_plot() +
  facet_wrap(fishing ~ ., ncol = 1, scales = "free") +
  labs(x = "Length (m)", y = "Main engine power (kW)") +
  scale_y_continuous(labels = scales::comma)
```

### Size distribution of vessel class for S1 and AIS data

```{r}
#| label: fig-s1-ais-size-class-comparison
#| #| fig-cap: Figure summarizing the length distributions for each vessel type (fishing and non-fishing) and size class, separately for AIS-broadcasting vessels and S1 detections. The point represents the median value, the vertical bars represent the interquartile range (i.e., 25th percentile and 75th percentile), and the lightly colored horizonal lines represents the full range of values.
#| fig-height: 6
#| fig-width: 8
#|
s1_ais_vessel_size_class_comparison |>
  dplyr::mutate(
    fishing = ifelse(fishing, "Fishing", "Non-fishing") |>
      forcats::fct_rev()
  ) |>
  dplyr::mutate(
    SOURCE = dplyr::case_when(
      source == "s1" ~ "S1 unmatched detections",
      source == "ais" ~ "AIS-broadcasting"
    ),
    length_size_class_percentile = glue::glue(
      "Length bin: {length_size_class_percentile}"
    ) %>%
      forcats::fct_relevel("Length bin: 10", after = Inf)
  ) |>
  ggplot(aes(x = median_length_m, y = SOURCE)) +
  geom_point(size = 1) +
  geom_errorbar(
    aes(xmin = min_length_m, xmax = max_length_m),
    width = 0,
    alpha = 0.25
  ) +
  geom_errorbar(aes(xmin = iqr_lower_length_m, xmax = iqr_upper_length_m)) +
  facet_grid(length_size_class_percentile ~ fishing, scales = "free_x") +
  theme_plot() +
  labs(x = "Vessel length (m)", y = "") +
  scale_x_continuous(limits = c(0, NA))
```

###  Fraction matched detections, by vessel size and type

```{r}
#| label: fig-dark-vessel-fraction-detected-time-series
#| fig-cap: Fraction of S1-detected vessels matched AIS-broadcasting vessels over time, disaggregated by vessel type and length size class
#| fig-height: 6
#| fig-width: 8

summarized_s1_dark_fleet_data <- annual_non_broadcasting_detections_emissions |>
  dplyr::group_by(year, fishing, length_size_class_percentile) %>%
  dplyr::summarize(across(
    c(number_dark_detections, number_ais_detections, emissions_co2_dark_mt),
    ~ sum(., na.rm = TRUE)
  )) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    ratio_dark_to_ais_detections = number_dark_detections /
      number_ais_detections,
    fraction_tracked = number_ais_detections /
      (number_ais_detections + number_dark_detections)
  ) |>
  dplyr::mutate(
    fishing = ifelse(fishing, "Fishing", "Non-fishing") |>
      forcats::fct_rev()
  )

summarized_s1_dark_fleet_data |>
  ggplot(aes(
    x = year,
    y = ratio_dark_to_ais_detections,
    color = factor(length_size_class_percentile),
    group = length_size_class_percentile
  )) +
  geom_line(linewidth = 1.1) +
  paletteer::scale_color_paletteer_d(palette = "jcolors::pal10") +
  theme_plot() +
  facet_grid(. ~ fishing) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(
    color = "Length bin",
    x = "",
    y = "Ratio of S1-detected dark vessels\nto matched AIS-broadcasting vessels"
  ) +
  guides(
    linetype = guide_legend(order = 2, reverse = TRUE),
    color = guide_legend(order = 1)
  )
```

### Outside and within S1 footprint ratios

```{r}
#| label: fig-map-dark-to-ais-s1-ratio
#| fig-cap: Mean ratio value within and outside S1 footprint, aggregated across all years, and vessel lengths, for representation purposes. Model inputs are disaggregated by pixel, month, vessel class and length bin.

scale_limits <- c(
  0.001,
  max(s1_mean_spatial_knn_ratios$avg_neighbor_ratio, na.rm = TRUE)
)

s1_mean_spatial_knn_ratios |>
  dplyr::mutate(
    within_footprint = ifelse(
      within_footprint,
      "Within S1 footprint",
      "Outside S1 footprint"
    ) |>
      forcats::fct_rev()
  ) |>
  dplyr::mutate(
    fishing = ifelse(fishing, "Fishing", "Non-fishing") |>
      forcats::fct_rev()
  ) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "black") +
  ggspatial::geom_spatial_tile(aes(
    x = lon_bin + 0.5,
    y = lat_bin + 0.5,
    fill = avg_neighbor_ratio,
    color = avg_neighbor_ratio
  )) +
  theme_void() +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scale_fill_viridis_c(
    "Average ratios (log scale)",
    trans = "log10",
    option = "mako",
    na.value = "black",
    limits = scale_limits
  ) +
  scale_color_viridis_c(
    "Average ratios (log scale)",
    trans = "log10",
    option = "mako",
    na.value = "black",
    limits = scale_limits
  ) +
  labs(x = "", y = "") +
  theme(legend.position = "bottom") +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black",
      barwidth = unit(10, "cm"),
      barheight = unit(0.5, "cm")
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  facet_grid(fishing ~ within_footprint) +
  coord_sf(crs = map_projection)
```

###  Dark fleet emissions, by vessel size and type

```{r}
#| label: fig-global-annual-co2-dark-emissions-by-type
#| fig-cap: Summary of global annual CO2 dark emissions over time, disaggregated by vessel type.
#| fig-height: 8
#| fig-width: 8
#|
summarized_s1_dark_fleet_data |>
  ggplot(aes(
    x = year,
    y = emissions_co2_dark_mt,
    fill = factor(length_size_class_percentile)
  )) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_continuous(breaks = unique(summarized_s1_dark_fleet_data$year)) +
  labs(x = "", y = "CO2 dark emissions (MT)", fill = "Length bin") +
  theme_plot() +
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  facet_grid(fishing ~ ., scales = "free_y") +
  paletteer::scale_fill_paletteer_d(palette = "jcolors::pal10")
```


```{r}
#| label: fig-pollutant-maps
#| fig-cap: Spatial map of 2024 emissions for each pollutant, aggregated across the AIS-broadcasting fleet and non-broadcasting, and shown at a 1x1 degree spatial resolution
#| fig-width: 8
#| fig-height: 6
pollutant_map_data_all <- total_spatial_emissions_by_pollutant |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

pollutant_map_data_all |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      color = emissions_mt_per_km2_2024,
      fill = emissions_mt_per_km2_2024
    )
  ) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scico::scale_fill_scico(
    name = expression(2024 ~ emissions ~ (MT / km^2)),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(0, 0.1, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  scico::scale_color_scico(
    name = expression(2024 ~ emissions ~ (MT / km^2)),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(0, 0.1, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme_void() +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black"
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(2, "cm")
  ) +
  facet_wrap(pollutant ~ ., ncol = 3, labeller = label_parsed)
```

```{r}
#| label: fig-co2-emissions-time-series
#| fig-cap: Time series of monthly global CO2 emissions, aggregated across AIS-broadcasting vessels and non-broadcasting vessels.
#| fig-width: 8
#| fig-height: 3

# Total monthly non-spatial emissions by pollutant
total_monthly_emissions_by_pollutant <- total_monthly_emissions_by_pollutant |>
  dplyr::mutate(month = lubridate::ymd(month)) |>
  tidyr::pivot_longer(
    -month,
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_total_mt") |>
      stringr::str_to_upper() |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::filter(
    lubridate::year(month) <= analysis_end_year,
    lubridate::year(month) >= analysis_start_year
  )

pollutant_time_series_data_co2 <- total_monthly_emissions_by_pollutant |>
  dplyr::filter(pollutant == "CO2")

make_pollutant_time_series_figure <- function(
    pollutant_time_series_data,
    y_axis_label) {{ ggplot(
  data = pollutant_time_series_data,
  aes(x = month, y = emissions_mt)
) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::comma) +
  theme_plot() +
  labs(x = "", y = y_axis_label) +
  theme(panel.grid.minor.x = element_blank()) +
  scale_x_continuous(
    breaks = seq(
      lubridate::ymd(glue::glue("{analysis_start_year}-01-01")),
      lubridate::ymd(glue::glue("{analysis_end_year+1}-01-01")),
      by = "1 year"
    ),
    labels = seq(analysis_start_year, analysis_end_year + 1)
  ) +
  scale_y_continuous(limits = c(0, NA)) }}

make_pollutant_time_series_figure(
  pollutant_time_series_data_co2,
  y_axis_label = expression(Monthly ~ CO[2] ~ emissions ~ (MT))
) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  )
```

```{r}
#| label: fig-pollutant-time-series
#| fig-cap: Time series of monthly global emissions for each pollutant, aggregated across AIS-broadcasting vessels and non-broadcasting vessels.
#| fig-width: 8
#| fig-height: 8

pollutant_time_series_data_all <- total_monthly_emissions_by_pollutant |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

make_pollutant_time_series_figure(
  pollutant_time_series_data_all,
  y_axis_label = "Monthly emissions (MT)"
) +
  facet_wrap(
    pollutant ~ .,
    ncol = 3,
    scales = "free_y",
    labeller = label_parsed
  )
```

```{r}
#| label: fig-co2-emissions-by-vessel-type
#| fig-cap: Total 2024 CO2 emissions by vessel type for AIS-broadcasting vessels (millions metric tonnes).
#| fig-width: 4
#| fig-height: 6
annual_ais_co2_emissions_by_vessel_type |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    fishing = ifelse(
      vessel_class %in%
        c(
          "trawlers",
          "squid_jigger",
          "tuna_purse_seines",
          "set_longlines",
          "pole_and_line",
          "set_gillnets",
          "pots_and_traps",
          "dredge_fishing",
          "other_seines",
          "other_purse_seines",
          "trollers"
        ),
      "Fishing",
      "Non-fishing"
    ) |>
      forcats::fct_rev()
  ) |>
  dplyr::mutate(
    vessel_class = stringr::str_replace_all(
      vessel_class,
      "\\.",
      ": "
    ) |>
      stringr::str_replace_all("_", " ") |>
      stringr::str_to_sentence() |>
      stringr::str_replace_all(
        "chemical oil",
        "chemical or oil"
      )
  ) |>
  ggplot(aes(
    x = emissions_co2_mt,
    y = reorder(vessel_class, emissions_co2_mt)
  )) +
  geom_col(position = position_dodge2(preserve = "single")) +
  ggforce::facet_col(~fishing, space = "free", scales = "free") +
  theme_plot() +
  labs(
    x = "2024 global CO2 emissions (MT)",
    y = ""
  ) +
  scale_x_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  )
```

# Supplementary tables

## Summary of 2016 and 2024 emissions, by pollutant and data source

```{r}
total_percent_change_by_fleet_tbl <- total_percent_change_by_fleet |>
  dplyr::mutate(across(
    c(`2016`, `2024`, percent_change),
    ~ signif(., 3) |>
      prettyNum(big.mark = ",")
  )) |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::arrange(pollutant) |>
  dplyr::rename(
    `Data source` = fleet,
    Pollutant = pollutant,
    `2016 emissions (MT)` = `2016`,
    `2024 emissions (MT)` = `2024`,
    `Percent 2024 total` = percent_total_2024,
    `Percent change from 2016 to 2024` = percent_change
  ) |>
  knitr::kable(
    format = "latex",
    label = "total_percent_change_by_fleet",
    caption = "Summary of 2016 and 2024 emissions, by pollutant and data source."
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

writeLines(
  total_percent_change_by_fleet_tbl,
  here::here("tables/total_percent_change_by_fleet.tex")
)
```

## Underestimation of 2024 emissions using only AIS data, and overestimation of 2016 to 2024 emissions trend using only AIS data, by pollutant.

```{r}
pollutant_ais_underestimation_overestimation_summary_tbl <- pollutant_ais_underestimation_overestimation_summary |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::arrange(pollutant) |>
  dplyr::rename(
    Pollutant = pollutant,
    "AIS-only underestimation of 2024 emissions" = ais_underestimation_2024,
    "AIS-only overestimation of 2016-2024 emissions trend" = trend_overestimation_2016_2024
  ) |>
  knitr::kable(
    format = "latex",
    label = "pollutant_ais_underestimation_overestimation_summary",
    caption = "Underestimation of 2024 emissions using only AIS data, and overestimation of 2016 to 2024 emissions trend using only AIS data, by pollutant."
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

writeLines(
  pollutant_ais_underestimation_overestimation_summary_tbl,
  here::here(
    "tables/pollutant_ais_underestimation_overestimation_summary.tex"
  )
)
```

### Dynamic AIS - annual emissions and vessels by AIS receiver type

```{r}
#| label: wrangle-receiver-type-data

annual_global_emissions_by_receiver_type <- annual_global_emissions_by_receiver_type |>
  # Don't include unknown receiver type
  dplyr::filter(receiver_type != "unknown") |>
  # Fill in missing values with 0s
  tidyr::complete(
    receiver_type,
    tidyr::nesting(year),
    fill = list(emissions_co2_mt = 0, n_distinct_ssvid = 0)
  ) |>
  dplyr::mutate(receiver_type = stringr::str_to_sentence(receiver_type)) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025)

# Annual global emissions and number of unique vessels by receiver type (satellite; dynamic; or terrestrial) and flag
annual_global_emissions_by_receiver_type_and_flag <- annual_global_emissions_by_receiver_type_and_flag |>
  # Add column name for country name based on flag

  dplyr::mutate(
    flag = countrycode::countrycode(
      flag,
      origin = "iso3c",
      destination = "country.name"
    )
  ) |>
  # Don't include unknown receiver type
  dplyr::filter(receiver_type != "unknown") |>
  dplyr::mutate(
    receiver_type = stringr::str_to_sentence(receiver_type) |>
      forcats::fct_relevel("Satellite", "Terrestrial", "Dynamic")
  ) |>
  # Fill in missing values with 0s
  tidyr::complete(
    receiver_type,
    tidyr::nesting(year, flag),
    fill = list(emissions_co2_mt = 0, n_distinct_ssvid = 0)
  ) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025)

# Annual spatial CO2 emissions vt receiver type  (satellite; dynamic; or terrestrial)
annual_spatial_emissions_by_receiver_type <- annual_spatial_emissions_by_receiver_type |>
  dplyr::mutate(
    receiver_type = stringr::str_to_sentence(receiver_type) |>
      forcats::fct_relevel("Satellite", "Terrestrial", "Dynamic")
  ) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025)
```


```{r}
#| label: fig-annual-emissions-by-ais-receiver-type
#| fig-cap: Annual total CO2 emissions (metric tonnes) by the AIS-broadcasting fleet, disaggregated by AIS receiver type (satellite, terrestrial, and dynamic) and total aggregated across receiver types. A dashed vertical line is shown for 2022, the first year in which dynamic AIS was widely adopted.
#| #| fig-width: 7.5
#| fig-height: 6
annual_global_emissions_by_receiver_type |>
  dplyr::rename(
    `CO[2]~emissions~(MT)` = emissions_co2_mt,
    `Number~unique~vessels` = n_distinct_ssvid
  ) |>
  tidyr::pivot_longer(cols = -c(year, receiver_type)) |>
  ggplot(aes(x = year, y = value, color = receiver_type)) +
  facet_wrap(name ~ ., scales = "free_y", ncol = 1, labeller = label_parsed) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  scale_x_continuous(
    breaks = unique(annual_global_emissions_by_receiver_type$year)
  ) +
  scale_color_brewer(
    type = "qual",
    palette = "Set2",
    guide = guide_legend(reverse = TRUE)
  ) +
  theme_plot() +
  labs(x = "", y = "", color = "AIS receiver type") +
  theme(panel.grid.minor.x = element_blank()) +
  geom_vline(xintercept = 2022, linetype = 2)
```

### Dynamic AIS - annual emissions by AIS receiver type for top flags

```{r}
#| label: fig-annual-emissions-by-ais-receiver-type-top-flags
#| fig-cap: Annual total CO2 emissions (metric tonnes) by the AIS-broadcasting fleet, disaggregated by AIS receiver type (satellite, terrestrial, and dynamic) and total aggregated across receiver types, for the top 10 flags with the most emissions in 2024. A dashed vertical line is shown for 2022, the first year in which dynamic AIS was widely adopted.
#| fig-width: 7.5
#| fig-height: 6

top_n_flags_2024 <- annual_global_emissions_by_receiver_type_and_flag |>
  dplyr::filter(year == 2024, receiver_type == "Total") |>
  dplyr::slice_max(n = 10, order_by = emissions_co2_mt) |>
  dplyr::select(flag)

annual_global_emissions_by_receiver_type_and_flag |>
  dplyr::inner_join(top_n_flags_2024, by = "flag") |>
  dplyr::mutate(flag = forcats::fct_relevel(flag, top_n_flags_2024$flag)) |>
  ggplot(aes(x = year, y = emissions_co2_mt, color = flag)) +
  facet_grid(receiver_type ~ ., scales = "free_y") +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  scale_x_continuous(
    breaks = unique(annual_global_emissions_by_receiver_type_and_flag$year)
  ) +
  theme_plot() +
  labs(x = "", y = expression(CO[2] ~ emissions ~ (MT)), color = "Flag") +
  theme(panel.grid.minor.x = element_blank()) +
  geom_vline(xintercept = 2022, linetype = 2) +
  paletteer::scale_color_paletteer_d(palette = "rcartocolor::Safe")

```


### Dynamic AIS - map showing spatial change in emissions by AIS receiver type


```{r}
#| label: fig-co2-emissions-change-map-by-receiver-type
#| fig-cap: Spatial change in CO2 emissions by the AIS-broadcasting fleet between 2016 and 2024 (metric tonnes) by receiver type (satellite, terrestrial, and dynamic) and total across receiver types, at a 1x1 degree spatial resolution.
#| fig-width: 7.5
#| fig-height: 8

annual_spatial_emissions_by_receiver_type |>
  dplyr::filter(year %in% c(2016, 2024)) |>
  tidyr::pivot_wider(names_from = year, values_from = emissions_co2_mt) |>
  dplyr::mutate(across(c(`2016`, `2024`), ~ tidyr::replace_na(., 0))) |>
  dplyr::mutate(change_emissions_co2 = `2024` - `2016`) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  ggspatial::geom_spatial_tile(aes(
    x = lon_bin + 0.5,
    y = lat_bin + 0.5,
    fill = change_emissions_co2,
    color = change_emissions_co2
  )) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  facet_wrap(receiver_type ~ ., ncol = 1) +
  scico::scale_fill_scico(
    expression(atop(
      Change ~ of ~ CO[2] ~ emissions,
      from ~ 2016 ~ to ~ 2024 ~ (MT)
    )),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e6, -1e3, 0, 1e3, 1e6),
    labels = scales::comma
  ) +
  scico::scale_color_scico(
    expression(atop(
      Change ~ of ~ CO[2] ~ emissions,
      from ~ 2016 ~ to ~ 2024 ~ (MT)
    )),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e6, -1e3, 0, 1e3, 1e6),
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme_void() +
  theme(legend.position = "right", legend.direction = "vertical") +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black",
      barwidth = unit(1, "cm"),
      barheight = unit(5, "cm")
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  coord_sf(crs = map_projection)
```