---
title: "Quantifying comprehensive marine vessel emissions using data fusion"
author: "emLab, UCSB and GFW"
date: last-modified
date-format: "[Last Updated on] MMMM DD, YYYY"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    fig-pos: H
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
cache: true
---

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE, cache = FALSE}
#| label: document-setup

# Set default chunk options
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  dev = "png",
  fig.path = "figures/"
)

# Load librariers
library(ggplot2)
library(sf)

analysis_start_year <- 2016
analysis_end_year <- 2024
```

```{r}
#| label: load-data
targets::tar_load(c(
  "n_unique_vessels",
  "n_ais_messages",
  "annual_edgar_emissions"
))
```

```{r}
#| label: wrangle-edgar-data
annual_edgar_tranportation_emissions <- annual_edgar_emissions |>
  dplyr::select(
    -c(
      IPCC_annex,
      C_group_IM24_sh,
      Country_code_A3,
      Name,
      ipcc_code_2006_for_standard_report,
      Substance,
      fossil_bio
    )
  ) |>
  dplyr::rename(sector = ipcc_code_2006_for_standard_report_name) |>
  tidyr::pivot_longer(
    cols = -c(sector),
    names_to = "year",
    values_to = "emissions_co2_kt"
  ) |>
  dplyr::mutate(
    year = stringr::str_remove_all(year, "Y_") |>
      as.numeric(),
    emissions_co2_mt = emissions_co2_kt * 1e3
  ) |>
  dplyr::select(-emissions_co2_kt) |>
  dplyr::mutate(
    land_based_tranpsportation = ifelse(
      sector %in%
        c(
          "Civil Aviation",
          "Road Transportation no resuspension",
          "Other Transportation",
          "Railways"
        ),
      TRUE,
      FALSE
    )
  ) |>
  # Only need years that correspond to our dataset
  dplyr::filter(year >= analysis_start_year, year <= analysis_end_year)

# Fraction of global 2023 CO2 emissions that came from transportation sectors
fraction_co2_from_transportation <- annual_edgar_tranportation_emissions |>
  dplyr::mutate(
    transportation = ifelse(
      land_based_tranpsportation | sector == "Water-borne Navigation",
      TRUE,
      FALSE
    )
  ) |>
  dplyr::filter(year == 2023) |>
  dplyr::group_by(transportation) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(
    names_from = "transportation",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::mutate(fraction_from_transportation = `TRUE` / (`TRUE` + `FALSE`)) |>
  dplyr::pull(fraction_from_transportation) |>
  scales::percent()
```

# In-line manuscript statistics

- The transportation sector is a major source of global greenhouse gas (GHG) and local air pollution emissions, responsible for `r fraction_co2_from_transportation` of all CO$_{2}$ emissions globally in 2023 

- For over `r scales::comma(signif(n_unique_vessels$n_unique_vessels,2))` AIS-broadcasting vessels, we quantify point-estimate emissions for each vessel using precise latitude/longitude/timestamp information for over `r scales::comma(signif(n_ais_messages$n_billion_ais_messages,3))` billion individual AIS messages