---
title: "Quantifying comprehensive marine vessel emissions using data fusion"
author: "emLab, UCSB and GFW"
date: last-modified
date-format: "[Last Updated on] MMMM DD, YYYY"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    fig-pos: H
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
cache: true
---

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE, cache = FALSE}
#| label: document-setup

# Set default chunk options
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  dev = "png",
  fig.path = "../figures/"
)

# Load librariers
library(ggplot2)
library(sf)

analysis_start_year <- 2017
analysis_end_year <- 2024
```

```{r}
#| label: load-data
#| cache: FALSE
targets::tar_load(c(
  "n_unique_vessels",
  "n_ais_messages",
  "n_s1_detections",
  "annual_edgar_emissions",
  "all_performance_metrics",
  "monthly_aggregated_time_series",
  "annual_emissions_all_pollutants",
  "total_spatial_emissions_by_pollutant",
  "annual_ais_to_dark_activity_extrapolation",
  "total_monthly_emissions_by_pollutant",
  "annual_ais_co2_emissions_by_vessel_type",
  "port_visit_co2_emissions_by_country",
  "trip_co2_emissions_by_from_to_countries",
  "annual_non_broadcasting_detections_emissions",
  "annual_global_emissions_by_receiver_type",
  "annual_global_emissions_by_receiver_type_and_flag",
  "annual_spatial_emissions_by_receiver_type",
  "vessel_size_info",
  "s1_ais_vessel_size_class_comparison",
  "registered_validation_data",
  "mrv_data_validation",
  "trip_emissions_for_mrv_validation",
  "registered_validation_data",
  "all_varimp_data",
  "performance_detections_cls_roc_pr_curves",
  "performance_detections_cls_conf_mat"
))
```

```{r}
#| label: setup-graphics

# For figures and tables, order pollutants first for GHG pollutants,
# then local pollutants. For figures, we use expressions to show subscripts
pollutant_factor_levels <- c(
  "CO[2]",
  "CH[4]",
  "N[2]*O",
  "CO",
  "NO[X]",
  "SO[X]",
  "PM[2.5]",
  "PM[10]",
  "VOCs"
)

# Set pollutant levels for tables
# For tables, we don't use expressions
pollutant_factor_levels_tables <- pollutant_factor_levels |>
  stringr::str_remove_all("\\[") |>
  stringr::str_remove_all("\\]") |>
  stringr::str_remove_all("\\*]*")

# Use Okabe-Ito colorblind-friendly colors
all_data_source_color_palette <- c(
  "Marine vessels (AIS + S1)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[5]],
  "GFW (AIS + S1)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[5]],
  "AIS + S1" = paletteer::paletteer_d("colorblindr::OkabeIto")[[5]],
  "Terrestrial transportation (EDGAR)" = paletteer::paletteer_d(
    "colorblindr::OkabeIto"
  )[[6]],
  "GFW (AIS)" = paletteer::paletteer_d("colorblindr::OkabeIto")[[3]],
  "AIS" = paletteer::paletteer_d("colorblindr::OkabeIto")[[3]],
  "S1" = paletteer::paletteer_d("colorblindr::OkabeIto")[[
    7
  ]],
  "EDGAR" = paletteer::paletteer_d("colorblindr::OkabeIto")[[8]],
  "OECD" = paletteer::paletteer_d("colorblindr::OkabeIto")[[6]],
  "IMO" = paletteer::paletteer_d("colorblindr::OkabeIto")[[1]]
)

# Set consistent theme for all plots
theme_plot <- function() {
  theme_minimal() %+replace%
    theme(
      axis.title.x = element_text(face = "bold", size = 12),
      axis.title.y = element_text(
        angle = 90,
        face = "bold",
        vjust = 3,
        size = 12
      ),
      strip.text.x = element_text(angle = 0, face = "bold", size = 14),
      strip.text.y = element_text(angle = 0, face = "bold", size = 14)
    )
}

# Set map projection for all map figures
map_projection <- "+proj=eqearth"

# Create bounding box of globe, for making black map backgrounds
lats <- c(90:-90, -90:90, 90)
longs <- c(rep(c(180, -180), each = 181), 180)
world_bbox_sf <-
  list(cbind(longs, lats)) %>%
  sf::st_polygon() %>%
  sf::st_sfc(
    # create sf geometry list column
    crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
  ) %>%
  sf::st_sf() %>%
  sf::st_transform(map_projection)

# ESRI World Countries Generalized Shapefile
# Downloaded from https://www.arcgis.com/home/item.html?id=fa510018bdd044b08fc64d2a16bc680a
land <- sf::st_read(
  here::here(
    "data/World_Countries_Generalized_Shapefile/World_Countries_Generalized.shp"
  )
) |>
  dplyr::mutate(
    country_iso3 = countrycode::countrycode(ISO, "iso2c", "iso3c")
  )


# Convert lat/lon pixel area to square km
# https://stackoverflow.com/a/1253545
# Need function to convert degrees to radians
# https://stackoverflow.com/q/32370485
deg2rad <- function(deg) {
  (deg * pi) / (180)
}

# Given a latitude, what is a pixel's area?
# https://stackoverflow.com/a/1253545
lat_to_area <- function(lat) {
  110.574 * (111.320 * cos(deg2rad(lat)))
}

```

```{r}
#| label: wrangle-edgar-data
annual_edgar_tranportation_emissions <- annual_edgar_emissions |>
  dplyr::select(
    -c(
      IPCC_annex,
      C_group_IM24_sh,
      Country_code_A3,
      Name,
      ipcc_code_2006_for_standard_report,
      Substance,
      fossil_bio
    )
  ) |>
  dplyr::rename(sector = ipcc_code_2006_for_standard_report_name) |>
  tidyr::pivot_longer(
    cols = -c(sector),
    names_to = "year",
    values_to = "emissions_co2_kt"
  ) |>
  dplyr::mutate(
    year = stringr::str_remove_all(year, "Y_") |>
      as.numeric(),
    emissions_co2_mt = emissions_co2_kt * 1e3
  ) |>
  dplyr::select(-emissions_co2_kt) |>
  dplyr::mutate(
    land_based_tranpsportation = ifelse(
      sector %in%
        c(
          "Civil Aviation",
          "Road Transportation no resuspension",
          "Other Transportation",
          "Railways"
        ),
      TRUE,
      FALSE
    )
  ) |>
  # Only need years that correspond to our dataset
  dplyr::filter(year >= analysis_start_year, year <= analysis_end_year)

# Fraction of global 2023 CO2 emissions that came from transportation sectors
fraction_co2_from_transportation <- annual_edgar_tranportation_emissions |>
  dplyr::mutate(
    transportation = ifelse(
      land_based_tranpsportation | sector == "Water-borne Navigation",
      TRUE,
      FALSE
    )
  ) |>
  dplyr::filter(year == 2023) |>
  dplyr::group_by(transportation) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(
    names_from = "transportation",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::mutate(fraction_from_transportation = `TRUE` / (`TRUE` + `FALSE`)) |>
  dplyr::pull(fraction_from_transportation) |>
  scales::percent()
```

```{r}
#| label: wrangle-gfw-time-series-emissions-data

annual_emissions_all_pollutants <- annual_emissions_all_pollutants |>
  tidyr::pivot_longer(
    -c(year, fishing),
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    fleet = ifelse(
      stringr::str_detect(pollutant, "dark"),
      "Non-broadcasting",
      "AIS-broadcasting"
    )
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_mt") |>
      stringr::str_remove_all("_dark")
  )

annual_emissions_all_pollutants_with_total <- annual_emissions_all_pollutants |>
  # Add total, that aggregates across broadcasting and non-broadcasting
  dplyr::bind_rows(
    annual_emissions_all_pollutants |>
      dplyr::group_by(year, fishing, pollutant) |>
      dplyr::summarize(emissions_mt = sum(emissions_mt)) |>
      dplyr::ungroup() |>
      dplyr::mutate(fleet = "Total")
  )

# Aggregate emissions across fishing/non-fishing
annual_emissions <- annual_emissions_all_pollutants_with_total |>
  dplyr::group_by(year, fleet, pollutant) |>
  dplyr::summarise(emissions_mt = sum(emissions_mt)) |>
  dplyr::ungroup()

# Calculate the total percent change in emissions by fleet
total_percent_change_by_fleet <- annual_emissions |>
  dplyr::filter(
    year %in% c(analysis_start_year, analysis_end_year),
    fleet != "Total"
  ) |>
  tidyr::pivot_wider(names_from = year, values_from = emissions_mt) |>
  dplyr::arrange(pollutant) |>
  dplyr::group_by(pollutant) |>
  dplyr::mutate(percent_total_2024 = `2024` / sum(`2024`)) |>
  dplyr::ungroup() %>%
  split(.[, c("pollutant")]) %>%
  purrr::map_df(janitor::adorn_totals, fill = NA) |>
  tidyr::fill(pollutant, .direction = "down") |>
  # janitor::adorn_totals("row") |>
  dplyr::mutate(percent_total_2024 = scales::percent(percent_total_2024)) |>
  dplyr::mutate(
    percent_change = 100 *
      (`2024` - `2017`) /
      `2017`
  ) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS (broadcasting)",
      TRUE ~ "S1 (non-broadcasting)"
    )
  )
```

```{r}
# For each pollutant, estimate how much emissions are underestimated by just using AIS
# As compared to using AIS + S1
ais_scale_underestimation_2024 <- total_percent_change_by_fleet |>
  dplyr::filter(fleet != "S1 (non-broadcasting)") |>
  dplyr::select(fleet, pollutant, `2024`) |>
  tidyr::pivot_wider(names_from = "fleet", values_from = `2024`) |>
  dplyr::mutate(
    ais_underestimation_2024 = (1 - `AIS (broadcasting)` / `AIS + S1`) |>
      scales::percent()
  ) |>
  dplyr::rename(
    emissions_mt_2024_ais = `AIS (broadcasting)`,
    emissions_mt_2024_ais_s1 = `AIS + S1`
  )

# For each pollutant, estimate how much the 2017->2024 trend in emissions is underestimated by just using AIS
# As compared to using AIS + S1
ais_trend_overestimation_2017_2024 <- total_percent_change_by_fleet |>
  dplyr::filter(fleet != "S1 (non-broadcasting)") |>
  dplyr::select(fleet, pollutant, percent_change) |>
  tidyr::pivot_wider(names_from = "fleet", values_from = percent_change) |>
  dplyr::mutate(
    trend_overestimation_2017_2024 = (`AIS (broadcasting)` /
      `AIS + S1` -
      1) |>
      scales::percent()
  ) |>
  dplyr::rename(
    trend_2017_2024_ais = `AIS (broadcasting)`,
    trend_2017_2024_ais_s1 = `AIS + S1`
  )

# Put all this information together
pollutant_ais_underestimation_overestimation_summary <- ais_scale_underestimation_2024 |>
  dplyr::left_join(ais_trend_overestimation_2017_2024, by = "pollutant") |>
  dplyr::select(
    pollutant,
    ais_underestimation_2024,
    trend_overestimation_2017_2024
  )
```

```{r}
#| label: wrangel-edgar-data
# Calculate annual terrestrial transportation emissions from EDGAR
land_based_transportation_emissions_edgar <- annual_edgar_tranportation_emissions |>
  dplyr::filter(land_based_tranpsportation) |>
  dplyr::group_by(year) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::mutate(data_source = "Terrestrial transportation (EDGAR)")

all_tranportation_data <- land_based_transportation_emissions_edgar |>
  dplyr::bind_rows(
    annual_emissions |>
      dplyr::filter(pollutant == "co2", fleet == "Total") |>
      dplyr::rename(emissions_co2_mt = emissions_mt) |>
      dplyr::mutate(data_source = "Marine vessels (AIS + S1)") |>
      dplyr::select(-c(fleet, pollutant))
  )


# Get analysis_start_year values for time-series normalization
all_tranportation_data_starting <- all_tranportation_data |>
  dplyr::filter(year == analysis_start_year) |>
  dplyr::mutate(emissions_co2_mt_starting = emissions_co2_mt) |>
  dplyr::select(-c(year, emissions_co2_mt))

all_tranportation_data_normalized <- all_tranportation_data |>
  dplyr::inner_join(all_tranportation_data_starting, by = "data_source") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_co2_mt /
      emissions_co2_mt_starting -
      1
  )


percent_transportation_emmissions_from_marine_2023 <- all_tranportation_data |>
  dplyr::filter(year == 2023) |>
  dplyr::select(-year) |>
  tidyr::pivot_wider(
    names_from = "data_source",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::mutate(
    percent = scales::percent(
      `Marine vessels (AIS + S1)` /
        (`Marine vessels (AIS + S1)` + `Terrestrial transportation (EDGAR)`)
    )
  ) |>
  dplyr::pull(percent)

percent_increase_from_2017_to_2023 <- all_tranportation_data_normalized |>
  dplyr::filter(year == 2023) |>
  dplyr::mutate(
    emissions_co2_mt_normalized_percent = scales::percent(
      emissions_co2_mt_normalized
    )
  )
```

```{r}
#| label: fig-inventory-comparison
#| #| fig-cap: Comparison of marine vessel CO$_{2}$ emissions inventory estimates, by year.
#| fig-width: 9
#| fig-height: 6
marine_based_transportation_emissions_edgar <- annual_edgar_tranportation_emissions |>
  dplyr::filter(sector == "Water-borne Navigation") |>
  dplyr::group_by(year) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::mutate(data_source = "EDGAR")

# Downloaded from here on August 27, 2025: https://data-explorer.oecd.org/vis?fs[0]=Topic%2C1%7CEnvironment%20and%20climate%20change%23ENV%23%7CAir%20and%20climate%23ENV_AC%23&pg=0&fc=Topic&bp=true&snb=17&df[ds]=dsDisseminateFinalDMZ&df[id]=DSD_MARITIME_TRANSPORT%40DF_MARITIME_TRANSPORT&df[ag]=OECD.SDD.NAD.SEEA&df[vs]=1.0&dq=W.A......ALL_VESSELS&pd=2019%2C2024&to[TIME_PERIOD]=false&vw=tb&isAvailabilityDisabled=false
oecd_data <- readr::read_csv(here::here(
  "data/oecd/annual_oecd_experimental_data.csv"
)) |>
  dplyr::select(year = TIME_PERIOD, emissions_co2_mt = OBS_VALUE) |>
  dplyr::mutate(data_source = "OECD")

imo_data <- tibble::tibble(
  year = c(2016, 2017, 2018),
  emissions_co2_mt = c(1.026e9, 1.064e9, 1.056e9)
) |>
  dplyr::mutate(data_source = "IMO")

all_inventory_data <- all_tranportation_data |>
  dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>
  dplyr::mutate(data_source = "GFW (AIS + S1)") |>
  dplyr::bind_rows(marine_based_transportation_emissions_edgar) |>
  dplyr::bind_rows(oecd_data) |>
  dplyr::bind_rows(imo_data) |>
  dplyr::bind_rows(
    annual_emissions |>
      dplyr::filter(pollutant == "co2", fleet == "AIS-broadcasting") |>
      dplyr::select(year, emissions_co2_mt = emissions_mt) |>
      dplyr::mutate(data_source = "GFW (AIS)")
  )

all_inventory_data |>
  ggplot(aes(x = year, y = emissions_co2_mt, color = data_source)) +
  geom_line(linewidth = 1) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "B", scale = 1e-9),
    limits = c(0, 1.5e9),
    breaks = seq(0, 1.5e9, 0.25e9)
  ) +
  theme_plot() +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(analysis_start_year, analysis_end_year)
  ) +
  labs(
    x = "",
    y = expression(Annual ~ CO[2] ~ emissions ~ (MT))
  ) +
  theme(panel.grid.minor = element_blank()) +
  scale_color_manual(
    name = "Emissions inventory",
    values = all_data_source_color_palette,
    breaks = c("GFW (AIS + S1)", "IMO", "GFW (AIS)", "OECD", "EDGAR")
  )

all_inventory_data |>
  dplyr::mutate(
    emissions_co2_mt = as.character(signif(emissions_co2_mt, 3) / 1e9)
  ) |>
  tidyr::pivot_wider(
    names_from = "data_source",
    values_from = "emissions_co2_mt"
  ) |>
  dplyr::rename(Year = year) |>
  dplyr::relocate(Year, `GFW (AIS + S1)`, `GFW (AIS)`) |>
  dplyr::mutate(across(everything(), ~ tidyr::replace_na(., "-"))) |>
  knitr::kable(
    format = "latex",
    label = "inventory_comparison",
    caption = "Comparison of annual marine vessel CO$_{2}$ emissions inventory estimates (billions of metric tonnes)."
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position") |>
  writeLines(
    here::here("tables/inventory_comparison.tex")
  )
```

```{r}
#| label: wrangle-gfw-spatials-emissions-data
# Total spatial emissions by pollutant for analysis_start_year and 2024
total_spatial_emissions_by_pollutant <- total_spatial_emissions_by_pollutant |>
  tidyr::pivot_longer(
    -c(lat_bin, lon_bin, year),
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_total_mt") |>
      stringr::str_to_upper() |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(emissions_mt_per_km2 = emissions_mt / lat_to_area(lat_bin)) |>
  dplyr::select(-emissions_mt) |>
  tidyr::pivot_wider(
    names_from = "year",
    values_from = "emissions_mt_per_km2",
    names_prefix = "emissions_mt_per_km2_"
  )
```

```{r}
# Summarize fraction of all estimate dark emissions that occurred inside the S1 footprint and in months that were imaged; inside the S1 footprint and in months that weren't imaged; and outside the S1 footprint entirely
dark_emissions_by_area <- monthly_aggregated_time_series |>
  dplyr::group_by(inside_footprint, s1_imaged) |>
  dplyr::summarise(emissions_co2_dark_mt = sum(emissions_co2_dark_mt)) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    percent_total = scales::percent(
      emissions_co2_dark_mt / sum(emissions_co2_dark_mt),
      accuracy = 0.1
    )
  )

# Summarize fraction of 2024 CO2 emissions that came from non-broadcasting vessels,
# disaggregated by fishing and non-fishing
dark_emissions_by_fishing <- monthly_aggregated_time_series |>
  dplyr::filter(lubridate::year(time) == 2024) |>
  dplyr::group_by(fishing) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    percent_dark = scales::percent(
      emissions_co2_dark_mt / (emissions_co2_dark_mt + emissions_co2_mt),
      accuracy = 0.1
    )
  )

monthly_fraction_fig <- monthly_aggregated_time_series |>
  dplyr::group_by(fishing, year = lubridate::year(time)) |>
  dplyr::summarise(
    emissions_co2_mt = sum(emissions_co2_mt),
    emissions_co2_dark_mt = sum(emissions_co2_dark_mt)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    fraction = emissions_co2_dark_mt /
      (emissions_co2_dark_mt + emissions_co2_mt)
  ) |>
  ggplot(aes(x = year, y = fraction, color = fishing)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA))
```

\# In-line manuscript statistics

-   The transportation sector is a major source of global greenhouse gas
    (GHG) and local air pollution emissions, responsible for
    `r fraction_co2_from_transportation` of all CO$_{2}$ emissions
    globally in 2023

-   For over
    `r scales::comma(signif(n_unique_vessels$n_unique_vessels, 2))`
    AIS-broadcasting vessels, we quantify point-estimate emissions for
    each vessel using precise latitude/longitude/timestamp information
    for over
    `r scales::comma(signif(n_ais_messages$n_billion_ais_messages, 3))`
    billion individual AIS messages

-   For non-broadcasting vessels, we leverage over
    `r floor(n_s1_detections$n_million_s1_detections)` million vessel
    detections from across more than
    `r floor(n_s1_detections$n_thousand_s1_scenes)` thousand S1 scenes.

-   Because many vessels do not broadcast AIS, current approaches solely
    based on AIS \textit{underestimate} current global emissions by
    `r pollutant_ais_underestimation_overestimation_summary |>   dplyr::filter(pollutant == "co2") |>   dplyr::pull(ais_underestimation_2024)`
    (in 2024), and \textit{overestimate} the increasing trend by
    `r pollutant_ais_underestimation_overestimation_summary |>   dplyr::filter(pollutant == "co2") |>   dplyr::pull(trend_overestimation_2017_2024)`
    (from `r analysis_start_year` to `r analysis_end_year`).

-   Using the latest data available from the EU's Emissions Database for
    Global Atmospheric Research (EDGAR), we estimate that in 2023 marine
    vessels accounted for
    `r percent_transportation_emmissions_from_marine_2023` of all
    transportation-related CO$_{2}$ emissions. Marine vessel CO$_{2}$
    emissions increased by
    `r percent_increase_from_2017_to_2023 |>   dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>   dplyr::pull(emissions_co2_mt_normalized_percent)`
    between `r analysis_start_year` and 2023,
    `r round(   (percent_increase_from_2017_to_2023 |>     dplyr::filter(data_source == "Marine vessels (AIS + S1)") |>     dplyr::pull(emissions_co2_mt_normalized)) /     (percent_increase_from_2017_to_2023 |>       dplyr::filter(data_source == "Terrestrial transportation (EDGAR)") |>       dplyr::pull(emissions_co2_mt_normalized)) )`
    times faster than the growth rate of terrestrial transportation
    emissions

-   Across the entire `r analysis_start_year` to `r analysis_end_year`
    time series,
    `r dark_emissions_by_area |>   dplyr::filter(inside_footprint & s1_imaged) |>   dplyr::pull(percent_total)`
    of non-broadcasting emissions occurred inside the S1 footprint and
    in pixel-months that were imaged;
    `r dark_emissions_by_area |>   dplyr::filter(inside_footprint & !s1_imaged) |>   dplyr::pull(percent_total)`
    occurred inside the S1 footprint but in pixel-months that weren't
    imaged; and
    `r dark_emissions_by_area |>   dplyr::filter(!inside_footprint) |>   dplyr::pull(percent_total)`
    occurred outside the S1 footprint entirely.

-   In `r analysis_end_year`,
    `r dark_emissions_by_fishing |> dplyr::filter(fishing) |> dplyr::pull(percent_dark)`
    of CO$_{2}$ emissions from fishing vessels were from
    non-broadcasting vessels, and
    `r dark_emissions_by_fishing |> dplyr::filter(!fishing) |> dplyr::pull(percent_dark)`
    of CO$_{2}$ emissions from non-fishing vessels were from
    non-broadcasting vessels. For both types of vessels, this fraction
    of emissions from non-broadcasting vessels has been declining
    steadily since 2020.

    # Figures

    ## Figure 1

    ### Figure 1, panel A

```{r}
total_annual_emissions_by_fleet_fig <- annual_emissions |>
  dplyr::filter(pollutant == "co2") |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS",
      TRUE ~ "S1"
    )
  ) |>
  ggplot(aes(x = year, y = emissions_mt, color = fleet)) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "B", scale = 1e-9),
    limits = c(0, 1.5e9),
    breaks = seq(0, 1.5e9, 0.25e9)
  ) +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(analysis_start_year, analysis_end_year)
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS + S1", "AIS", "S1")
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(Annual ~ CO[2] ~ emissions ~ (MT))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.75, 0.4),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  )
```

### Figure 2, panel B

```{r}
# Get analysis_start_year values for time-series normalization
annual_co2_emissions_starting <- annual_emissions |>
  dplyr::filter(pollutant == "co2", year == analysis_start_year) |>
  dplyr::mutate(emissions_co2_mt_starting = emissions_mt) |>
  dplyr::select(-c(year, emissions_mt))

normalized_annual_emissions_by_fleet_fig <- annual_emissions |>
  dplyr::filter(pollutant == "co2") |>
  dplyr::left_join(annual_co2_emissions_starting, by = "fleet") |>
  dplyr::mutate(
    emissions_co2_mt_normalized = emissions_mt / emissions_co2_mt_starting - 1
  ) |>
  dplyr::mutate(
    fleet = dplyr::case_when(
      fleet == "Total" ~ "AIS + S1",
      fleet == "AIS-broadcasting" ~ "AIS",
      TRUE ~ "S1"
    )
  ) |>
  ggplot(aes(x = year, y = emissions_co2_mt_normalized, color = fleet)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  scale_x_continuous(
    breaks = unique(annual_emissions$year),
    limits = c(analysis_start_year, analysis_end_year)
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c("AIS", "AIS + S1", "S1")
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of ~ CO[2],
      emissions ~ from ~ 2017 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.25, 0.85),
    legend.background = element_rect(fill = "white", color = "gray"),
    legend.title = element_blank()
  )
```

### Figure 1, panel A

```{r}
pollutant_map_data_co2 <- total_spatial_emissions_by_pollutant |>
  dplyr::filter(pollutant %in% c("CO2")) |>
  dplyr::mutate(
    emissions_mt_per_km2_delta_abs = emissions_mt_per_km2_2024 -
      emissions_mt_per_km2_2017
  ) |>
  dplyr::select(
    lon_bin,
    lat_bin,
    emissions_mt_per_km2_delta_abs,
    emissions_mt_per_km2_2024
  ) |>
  tidyr::pivot_longer(
    cols = c(emissions_mt_per_km2_delta_abs, emissions_mt_per_km2_2024)
  )

co2_map <- pollutant_map_data_co2 |>
  dplyr::mutate(
    name = ifelse(
      name == "emissions_mt_per_km2_delta_abs",
      "Change in emissions from 2017 to 2024",
      "2024 emissions"
    )
  ) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  ggspatial::geom_spatial_tile(
    aes(
      x = lon_bin + 0.5,
      y = lat_bin + 0.5,
      color = value,
      fill = value
    )
  ) +
  facet_wrap(name ~ ., ncol = 1) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  scico::scale_fill_scico(
    name = expression(CO[2] ~ emissions ~ (MT / km^2)),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  scico::scale_color_scico(
    name = expression(CO[2] ~ emissions ~ (MT / km^2)),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme_void() +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black"
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(2, "cm")
  )
```

### Figure 1, combined

```{r}
#| label: fig-emissions-by-data-source-and-maps
#| #| fig-cap: (A) Annual marine vessel CO2 emissions, with each line colored based on the data source. (B) Relative change in annual marine vessel CO2 emissions from a 2017 baseline, with each line colored based on the data source. (C) Spatial map of CO2 emissions in 2024 (top) and change in CO2 emissions between 2017 and 2024 (bottom), aggregated across AIS-broadcasting vessels and non-broadcasting vessels, and shown at a 1x1 degree spatial resolution; emissions are area-normalized for each pixel.
#| fig-width: 9
#| fig-height: 6
cowplot::plot_grid(
  cowplot::plot_grid(
    total_annual_emissions_by_fleet_fig,
    normalized_annual_emissions_by_fleet_fig,
    ncol = 1,
    labels = c("A", "B")
  ),
  co2_map,
  ncol = 2,
  labels = c("", "C"),
  rel_widths = c(0.4, 0.6),
  align = "vh"
)
```

### Figure 2

```{r}
#| label: fig-emissions-marine-terrestrial
#| fig-cap: Comparison of emissions data between marine transportation emissions from GFW and terrestrial transportation emissions from EDGAR.
#| fig-width: 7
#| fig-height: 6

# Summarize total emissions by data source for 2023, for putting on figure
emissions_comparison_figure <- all_tranportation_data |>
  ggplot(aes(x = year, y = emissions_co2_mt, color = data_source)) +
  geom_line(linewidth = 1.05) +
  scale_x_continuous(
    breaks = unique(all_tranportation_data_normalized$year),
    limits = c(analysis_start_year, 2023)
  ) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "B", scale = 1e-9),
    limits = c(0, NA)
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(
      atop(Annual ~ transportation, ~ CO[2] ~ emissions ~ (MT))
    ),
    color = "Data source"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.65, 0.5),
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", color = "gray")
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c(
      "Terrestrial transportation (EDGAR)",
      "Marine vessels (AIS + S1)"
    )
  )

emissions_comparison_figure_normalized <- all_tranportation_data_normalized |>
  dplyr::filter(
    data_source %in%
      c("Terrestrial transportation (EDGAR)", "Marine vessels (AIS + S1)")
  ) |>
  ggplot(aes(x = year, y = emissions_co2_mt_normalized, color = data_source)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(linewidth = 1.1) +
  scale_x_continuous(
    breaks = unique(all_tranportation_data_normalized$year),
    limits = c(analysis_start_year, 2023)
  ) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  theme_plot() +
  labs(
    x = "",
    y = expression(atop(
      Relative ~ change ~ of,
      CO[2] ~ emissions ~ from ~ 2017 ~ baseline
    ))
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.25, 0.875),
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", color = "gray")
  ) +
  scale_color_manual(
    values = all_data_source_color_palette,
    breaks = c(
      "Marine vessels (AIS + S1)",
      "Terrestrial transportation (EDGAR)"
    )
  )

cowplot::plot_grid(
  emissions_comparison_figure,
  emissions_comparison_figure_normalized,
  ncol = 1,
  labels = c("A", "B")
)
```

## Figure 3

```{r}
#| label: build-decomposition-model

# For aggregate annual extrapolation dataset, calculated total emissions, kw-hours,distance, hours
annual_extrapolation_dataset <- annual_ais_to_dark_activity_extrapolation |>
  dplyr::mutate(
    emissions_co2_mt_total = emissions_co2_mt + emissions_co2_mt_dark,
    kw_hours_total = kw_hours + kw_hours_dark,
    distance_nm_total = distance_nm + distance_nm_dark,
    hours_total = hours + hours_dark
  ) |>
  dplyr::select(
    -c(
      emissions_co2_mt,
      emissions_co2_mt_dark,
      kw_hours,
      kw_hours_dark,
      distance_nm,
      distance_nm_dark,
      hours,
      hours_dark,
      avg_speed_knots
    )
  ) |>
  # Calculate average aggregate speed as total distance traveled divided by total hours
  dplyr::mutate(avg_speed_knots = distance_nm_total / hours_total)

# Using annual_extrapolation_dataset, use tidymodels to train a random forest model
# to predict emissions
set.seed(123)

# Set initial split of data
inital_split <- annual_extrapolation_dataset |>
  rsample::initial_split(prop = 0.8)

training_data <- rsample::training(inital_split)
testing_data <- rsample::testing(inital_split)

# Specify recipe for model
model_recipe <- recipes::recipe(
  emissions_co2_mt_total ~
    fishing + length_size_class_percentile + kw_hours_total + avg_speed_knots,
  data = head(rsample::training(inital_split))
) |>
  recipes::step_num2factor(
    length_size_class_percentile,
    levels = as.character(unique(
      annual_extrapolation_dataset$length_size_class_percentile
    ))
  )

# Specify model type
model_type <- parsnip::rand_forest(
  trees = 1000,
  # Default of ranger is squareroot of number of model features. We have 4 features, so use a value of 2.
  mtry = 2,
  # Default of ranger is 5 for regression
  min_n = 5
) |>
  parsnip::set_engine("ranger") |>
  parsnip::set_mode("regression")

# Initialize model workflow
model_workflow <- workflows::workflow() |>
  workflows::add_recipe(model_recipe) |>
  workflows::add_model(model_type)

# Fit model on training data
workflow_fit <- parsnip::fit(model_workflow, training_data)

# Make predictions on testing data, and augment this to testing data frame
worflow_predictions <- parsnip::augment(workflow_fit, testing_data)


# Calculate performance metrics

# Multiple regression metrics
multi_metric <- yardstick::metric_set(
  yardstick::rsq,
  yardstick::mape,
  yardstick::rsq_trad
)

performance_metrics <- multi_metric(
  data = worflow_predictions,
  truth = emissions_co2_mt_total,
  estimate = .pred
) |>
  dplyr::select(-.estimator)

# Final model on all data
# Fit final model on all data
workflow_fit_final <- parsnip::fit(model_workflow, annual_extrapolation_dataset)
```

```{r}
#| label: tbl-decomposition-model-performance
#| tbl-cap: Performance metrics for random forest model that predicts annual CO2 emissions, by vessel size class and type, using hours and average speed.

performance_metrics |>
  knitr::kable(digits = 3)
```

```{r}
#| label: predict-using-decomposition-model
# Find initial/starting 2017 values of kw-hours and speed, by fishing and length_size_class_percentile, for decomposition
starting_values <- annual_extrapolation_dataset |>
  dplyr::filter(year == analysis_start_year) |>
  dplyr::select(
    fishing,
    length_size_class_percentile,
    kw_hours_total_starting = kw_hours_total,
    hours_total_starting = hours_total,
    avg_speed_knots_starting = avg_speed_knots
  )

# Add starting values to predictions dataset
predictions_dataset <- annual_extrapolation_dataset |>
  dplyr::left_join(
    starting_values,
    by = c("fishing", "length_size_class_percentile")
  )

final_model_fit <- workflows::extract_fit_parsnip(workflow_fit_final)

# Now run predictions for 4 different scenarios
final_prediction_dataset <- predictions_dataset |>
  # Add model predictions using observed kw-hours and observed speed
  dplyr::bind_cols(
    observed_kw_hours_observed_speed = predict(
      final_model_fit,
      predictions_dataset
    )$.pred
  ) |>
  # Add model predictions using observed kw-hours and constant speed (from 2017)
  dplyr::bind_cols(
    observed_kw_hours_constant_speed = predict(
      final_model_fit,
      predictions_dataset |>
        dplyr::select(-avg_speed_knots) |>
        dplyr::rename(avg_speed_knots = avg_speed_knots_starting)
    )$.pred
  ) |>
  # Add model predictions using constant kw-hours (from 2017) and observed speed
  dplyr::bind_cols(
    constant_kw_hours_observed_speed = predict(
      final_model_fit,
      predictions_dataset |>
        dplyr::select(-kw_hours_total) |>
        dplyr::rename(kw_hours_total = kw_hours_total_starting)
    )$.pred
  ) |>
  # Add model predictions using constant kw-hours (from 2017) and constant speed (from 2017)
  dplyr::bind_cols(
    constant_kw_hours_constant_speed = predict(
      final_model_fit,
      predictions_dataset |>
        dplyr::select(-avg_speed_knots) |>
        dplyr::rename(avg_speed_knots = avg_speed_knots_starting) |>
        dplyr::select(-kw_hours_total) |>
        dplyr::rename(kw_hours_total = kw_hours_total_starting)
    )$.pred
  )
```

## Figure 3

```{r}
#| label: fig-hours-speed-decomposition
#| fig-cap: Estimated decomposition of change in total CO2 emissions from 2017 baseline (aggregating across both broadcasting and non-broadcasting vessels), using the random forest model outputs of four different scenarios that decompose the relative impact of changes in total kW-hours and changes in average speed/
#| fig-width: 7.5
#| fig-height: 3
# Aggregate prediction values by year (across vessel size classes and type)
final_annual_prediction_dataset <- final_prediction_dataset |>
  dplyr::group_by(year) |>
  dplyr::summarise(across(
    c(
      observed_kw_hours_observed_speed,
      observed_kw_hours_constant_speed,
      constant_kw_hours_observed_speed,
      constant_kw_hours_constant_speed
    ),
    ~ sum(., na.rm = TRUE)
  )) |>
  dplyr::rename(
    `Observed kW-hours, observed speed` = observed_kw_hours_observed_speed,
    `Observed kW-hours, constant speed` = observed_kw_hours_constant_speed,
    `Constant kW-hours, observed speed` = constant_kw_hours_observed_speed,
    `Constant kW-hours, constant speed` = constant_kw_hours_constant_speed
  ) |>
  tidyr::pivot_longer(-year)

# We will normalize all of these predictions to the 2017 starting prediction
starting_value_from_prediction <- final_annual_prediction_dataset |>
  dplyr::filter(year == analysis_start_year) |>
  dplyr::rename(starting_value = value) |>
  dplyr::select(-year)

time_series_predicted_decomposition_df <- final_annual_prediction_dataset |>
  dplyr::inner_join(starting_value_from_prediction, by = "name") |>
  dplyr::mutate(normalized_value = value / starting_value - 1)

time_series_predicted_decomposition_df |>
  ggplot(aes(x = year, y = normalized_value, color = name)) +
  geom_line(linewidth = 1.1) +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_color_manual(
    "Prediction scenario",
    values = c(
      "coral3",
      "coral4",
      paletteer::paletteer_d("colorblindr::OkabeIto")[[2]],
      paletteer::paletteer_d("colorblindr::OkabeIto")[[5]]
    )
  ) +
  scale_x_continuous(breaks = unique(final_prediction_dataset$year)) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "",
    y = expression(atop(
      Percent ~ change ~ of ~ total ~ CO[2],
      emissions ~ from ~ 2017 ~ basline
    )),
  ) +
  theme_plot() +
  theme(panel.grid.minor.x = element_blank())
```

# Supplementary figures

### Relationship between length and engine power

```{r}
#| label: fig-ais-length-power-relationship
#| fig-cap: The relationship between main engine power (kilowatts) and length (meters), where each point is a specific AIS-broadcasting vessel. The relationship is disaggregated for non-fishing vessels and fishing vessels.
#| #| fig-height: 10
#| #| fig-width: 6
vessel_size_info |>
  dplyr::mutate(
    fishing = ifelse(
      fishing,
      "Fishing",
      "Non-fishing"
    ) |>
      forcats::fct_rev()
  ) |>
  ggplot(aes(x = length_m, y = main_engine_power_kw)) +
  geom_point(size = 0.1, alpha = 0.5) +
  theme_plot() +
  facet_wrap(fishing ~ ., ncol = 1, scales = "free") +
  labs(x = "Length (m)", y = "Main engine power (kW)") +
  scale_y_continuous(labels = scales::comma)
```

### Size distribution of vessel class for S1 and AIS data

```{r}
#| label: fig-s1-ais-size-class-comparison
#| #| fig-cap: Figure summarizing the length distributions for each vessel type (fishing and non-fishing) and size class, separately for AIS-broadcasting vessels and S1 detections. The point represents the median value, the vertical bars represent the interquartile range (i.e., 25th percentile and 75th percentile), and the lightly colored horizonal lines represents the full range of values.
#| fig-height: 6
#| fig-width: 8
#|
s1_ais_vessel_size_class_comparison |>
  dplyr::mutate(
    fishing = ifelse(fishing, "Fishing", "Non-fishing") |>
      forcats::fct_rev()
  ) |>
  dplyr::mutate(
    SOURCE = dplyr::case_when(
      source == "s1" ~ "S1 unmatched detections",
      source == "ais" ~ "AIS-broadcasting"
    ),
    length_size_class_percentile = glue::glue(
      "Length bin: {length_size_class_percentile}"
    ) %>%
      forcats::fct_relevel("Length bin: 10", after = Inf)
  ) |>
  ggplot(aes(x = median_length_m, y = SOURCE)) +
  geom_point(size = 1) +
  geom_errorbar(
    aes(xmin = min_length_m, xmax = max_length_m),
    width = 0,
    alpha = 0.25
  ) +
  geom_errorbar(aes(xmin = iqr_lower_length_m, xmax = iqr_upper_length_m)) +
  facet_grid(length_size_class_percentile ~ fishing, scales = "free_x") +
  theme_plot() +
  labs(x = "Vessel length (m)", y = "") +
  scale_x_continuous(limits = c(0, NA))
```

### Fraction matched detections, by vessel size and type

```{r}
#| label: fig-dark-vessel-fraction-detected-time-series
#| fig-cap: Fraction of S1-detected vessels matched AIS-broadcasting vessels over time, disaggregated by vessel type and length size class
#| fig-height: 6
#| fig-width: 8

summarized_s1_dark_fleet_data <- annual_non_broadcasting_detections_emissions |>
  dplyr::group_by(year, fishing, length_size_class_percentile) %>%
  dplyr::summarize(across(
    c(number_dark_detections, number_ais_detections, emissions_co2_dark_mt),
    ~ sum(., na.rm = TRUE)
  )) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    ratio_dark_to_ais_detections = number_dark_detections /
      number_ais_detections,
    fraction_tracked = number_ais_detections /
      (number_ais_detections + number_dark_detections)
  ) |>
  dplyr::mutate(
    fishing = ifelse(fishing, "Fishing", "Non-fishing") |>
      forcats::fct_rev()
  )

summarized_s1_dark_fleet_data |>
  ggplot(aes(
    x = year,
    y = ratio_dark_to_ais_detections,
    color = factor(length_size_class_percentile),
    group = length_size_class_percentile
  )) +
  geom_line(linewidth = 1.1) +
  paletteer::scale_color_paletteer_d(palette = "jcolors::pal10") +
  theme_plot() +
  facet_grid(. ~ fishing) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(
    color = "Length bin",
    x = "",
    y = "Ratio of S1-detected dark vessels\nto matched AIS-broadcasting vessels"
  ) +
  guides(
    linetype = guide_legend(order = 2, reverse = TRUE),
    color = guide_legend(order = 1)
  )
```

### Dark fleet emissions, by vessel size and type

```{r}
#| label: fig-global-annual-co2-dark-emissions-by-type
#| fig-cap: Summary of global annual CO2 dark emissions over time, disaggregated by vessel type.
#| fig-height: 8
#| fig-width: 8
#|
summarized_s1_dark_fleet_data |>
  ggplot(aes(
    x = year,
    y = emissions_co2_dark_mt,
    fill = factor(length_size_class_percentile)
  )) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_continuous(breaks = unique(summarized_s1_dark_fleet_data$year)) +
  labs(x = "", y = "CO2 dark emissions (MT)", fill = "Length bin") +
  theme_plot() +
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  facet_grid(fishing ~ ., scales = "free_y") +
  paletteer::scale_fill_paletteer_d(palette = "jcolors::pal10")
```

```{r}
#| label: fig-pollutant-maps
#| fig-cap: Spatial map of 2024 emissions for each pollutant, aggregated across the AIS-broadcasting fleet and non-broadcasting, and shown at a 1x1 degree spatial resolution
#| fig-width: 8
#| fig-height: 8
pollutant_map_data_all <- total_spatial_emissions_by_pollutant |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

pollutant_map_list <- purrr::map(
  unique(pollutant_map_data_all$pollutant),
  function(pollutant_tmp) {
    pollutant_tmp <- as.character(pollutant_tmp)

    pollutant_map_data_all |>
      dplyr::filter(pollutant == pollutant_tmp) |>
      ggplot() +
      geom_sf(data = world_bbox_sf, fill = "grey90") +
      ggspatial::geom_spatial_tile(
        aes(
          x = lon_bin + 0.5,
          y = lat_bin + 0.5,
          color = emissions_mt_per_km2_2024,
          fill = emissions_mt_per_km2_2024
        )
      ) +
      geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
      scico::scale_fill_scico(
        name = expression(Emissions ~ (MT / km^2)),
        palette = "vik",
        midpoint = 0,
        trans = ggallin::pseudolog10_trans,
        na.value = "grey90",
        labels = scales::comma
      ) +
      scico::scale_color_scico(
        name = expression(Emissions ~ (MT / km^2)),
        palette = "vik",
        midpoint = 0,
        trans = ggallin::pseudolog10_trans,
        na.value = "grey90",
        labels = scales::comma
      ) +
      labs(x = "", y = "") +
      theme_void() +
      guides(
        fill = guide_colorbar(
          title.position = "top",
          title.hjust = 0.5,
          frame.colour = "black",
          ticks.colour = "black"
        )
      ) +
      scale_x_continuous(expand = c(0, 1)) +
      scale_y_continuous(expand = c(0, 1)) +
      theme(
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.key.width = unit(1, "cm")
      ) +
      facet_wrap(pollutant ~ ., labeller = label_parsed)
  }
)

cowplot::plot_grid(plotlist = pollutant_map_list, ncol = 3)
```

```{r}
#| label: make-fig-co2-emissions-time-series

# Total monthly non-spatial emissions by pollutant
total_monthly_emissions_by_pollutant <- total_monthly_emissions_by_pollutant |>
  dplyr::mutate(month = lubridate::ymd(month)) |>
  tidyr::pivot_longer(
    -month,
    names_to = "pollutant",
    values_to = "emissions_mt"
  ) |>
  dplyr::mutate(
    pollutant = stringr::str_remove_all(pollutant, "emissions_") |>
      stringr::str_remove_all("_total_mt") |>
      stringr::str_to_upper() |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::filter(
    lubridate::year(month) <= analysis_end_year,
    lubridate::year(month) >= analysis_start_year
  )

pollutant_time_series_data_co2 <- total_monthly_emissions_by_pollutant |>
  dplyr::filter(pollutant == "CO2")

```

```{r}
#| label: fig-data-richness
#| fig-cap: (A) Time series of monthly global emissions for each pollutant, aggregated across AIS-broadcasting vessels and non-broadcasting vessels. (B) CO2 emissions during 2024 port stays, aggregated by country. (C) Total trip-level CO2 emissions in 2024 for domestic (within-country) trips for top 25 countries, and total trip-level CO2 emissions in 2024 for international trips for top 25 country-to-country paris. (D)  Total 2024 CO2 emissions by vessel type for AIS-broadcasting vessels.
#| fig-width: 9
#| fig-height: 8

make_pollutant_time_series_figure <- function(
  pollutant_time_series_data,
  y_axis_label
) {
  {
    ggplot(
      data = pollutant_time_series_data,
      aes(x = month, y = emissions_mt)
    ) +
      geom_line(linewidth = 1.1) +
      scale_y_continuous(labels = scales::comma) +
      theme_plot() +
      labs(x = "", y = y_axis_label) +
      theme(panel.grid.minor.x = element_blank()) +
      scale_x_continuous(
        breaks = seq(
          lubridate::ymd(glue::glue("{analysis_start_year}-01-01")),
          lubridate::ymd(glue::glue("{analysis_end_year+1}-01-01")),
          by = "1 year"
        ),
        labels = seq(analysis_start_year, analysis_end_year + 1)
      ) +
      scale_y_continuous(limits = c(0, NA))
  }
}

time_series_fig_co2 <- make_pollutant_time_series_figure(
  pollutant_time_series_data_co2,
  y_axis_label = expression(Monthly ~ CO[2] ~ emissions ~ (MT))
) +
  scale_y_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  )

# Divde international trip emissions equally to from and to countries
international_trip_emissions_by_country <- trip_co2_emissions_by_from_to_countries |>
  dplyr::filter(from_country_iso3 != to_country_iso3) |>
  dplyr::mutate(emissions_co2_mt = emissions_co2_mt / 2) |>
  dplyr::select(country_iso3 = from_country_iso3, emissions_co2_mt) |>
  dplyr::bind_rows(
    trip_co2_emissions_by_from_to_countries |>
      dplyr::mutate(emissions_co2_mt = emissions_co2_mt / 2) |>
      dplyr::select(country_iso3 = to_country_iso3, emissions_co2_mt)
  ) |>
  dplyr::group_by(country_iso3) |>
  dplyr::summarise(emissions_co2_mt = sum(emissions_co2_mt)) |>
  dplyr::ungroup()

domestic_trip_emissions_by_country <- trip_co2_emissions_by_from_to_countries |>
  dplyr::filter(from_country_iso3 == to_country_iso3) |>
  dplyr::select(country_iso3 = from_country_iso3, emissions_co2_mt)

all_port_and_trip_emissions <- international_trip_emissions_by_country |>
  dplyr::mutate(type = "International trips") |>
  dplyr::bind_rows(
    domestic_trip_emissions_by_country |>
      dplyr::mutate(type = "Domestic trips")
  ) |>
  dplyr::bind_rows(
    port_visit_co2_emissions_by_country |>
      dplyr::rename(country_iso3 = port_country_iso3) |>
      dplyr::mutate(type = "Port stays")
  )

all_countries_for_plot <- land |>
  sf::st_drop_geometry() |>
  dplyr::select(country_iso3) |>
  tidyr::crossing(
    all_port_and_trip_emissions |>
      dplyr::distinct(type)
  )

port_trip_emissions_fig <- ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  geom_sf(
    data = all_countries_for_plot |>
      dplyr::left_join(land, by = ) |>
      dplyr::left_join(
        all_port_and_trip_emissions,
        by = c("country_iso3", "type")
      ) |>
      sf::st_as_sf(),
    aes(fill = emissions_co2_mt),
    color = "white",
    linewidth = 0.01
  ) +
  facet_wrap(type ~ ., ncol = 1) +
  coord_sf(crs = map_projection) +
  theme_void() +
  scale_fill_viridis_c(
    option = "A",
    begin = 0.2,
    end = 0.7,
    na.value = "black",
    labels = scales::unit_format(unit = "M", scale = 1e-6)
  ) +
  labs(
    x = "",
    fill = expression(2024 ~ CO[2] ~ emissions ~ (MT))
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black"
    )
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(1.75, "cm")
  )

port_emissions_fig <- ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  geom_sf(
    data = land |>
      dplyr::rename(port_country_iso3 = country_iso3) |>
      dplyr::left_join(
        port_visit_co2_emissions_by_country,
        by = c("port_country_iso3")
      ),
    aes(fill = emissions_co2_mt),
    color = "white",
    linewidth = 0.01
  ) +
  coord_sf(crs = map_projection) +
  theme_void() +
  scale_fill_viridis_c(
    option = "A",
    begin = 0.2,
    end = 0.7,
    na.value = "black",
    labels = scales::unit_format(unit = "M", scale = 1e-6)
  ) +
  labs(
    x = "",
    fill = expression(2024 ~ port ~ stay ~ CO[2] ~ emissions ~ (MT))
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black"
    )
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(1, "cm")
  )

domestic_trip_emissions_fig <- ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  geom_sf(
    data = land |>
      dplyr::rename(from_country_iso3 = country_iso3) |>
      dplyr::left_join(
        trip_co2_emissions_by_from_to_countries |>
          dplyr::filter(from_country_iso3 == to_country_iso3),
        by = c("from_country_iso3")
      ),
    aes(fill = emissions_co2_mt),
    color = "white",
    linewidth = 0.01
  ) +
  coord_sf(crs = map_projection) +
  theme_void() +
  scale_fill_viridis_c(
    option = "A",
    begin = 0.2,
    end = 0.7,
    na.value = "black",
    labels = scales::unit_format(unit = "M", scale = 1e-6)
  ) +
  labs(
    x = "",
    fill = expression(2024 ~ domestic ~ trip ~ CO[2] ~ emissions ~ (MT))
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black"
    )
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(1, "cm")
  )

trip_emissions_fig <- trip_co2_emissions_by_from_to_countries |>
  dplyr::mutate(
    domestic = ifelse(from_country_iso3 == to_country_iso3, TRUE, FALSE)
  ) |>
  dplyr::group_by(domestic) |>
  dplyr::slice_max(n = 25, order_by = emissions_co2_mt) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    from_country = countrycode::countrycode(
      from_country_iso3,
      "iso3c",
      "country.name"
    )
  ) |>
  dplyr::mutate(
    to_country = countrycode::countrycode(
      to_country_iso3,
      "iso3c",
      "country.name"
    )
  ) |>
  dplyr::mutate(
    label = ifelse(
      domestic,
      from_country,
      glue::glue("{from_country} to {to_country}")
    )
  ) |>
  dplyr::mutate(
    domestic = ifelse(domestic, "Domestic trips", "International trips")
  ) |>
  ggplot(aes(x = emissions_co2_mt, y = reorder(label, emissions_co2_mt))) +
  geom_bar(stat = "identity") +
  facet_wrap(. ~ domestic, scales = "free_y") +
  theme_plot() +
  scale_x_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  labs(x = expression(2024 ~ CO[2] ~ emissions ~ (MT)), y = "")

emissions_by_vessel_class_barplot <- annual_ais_co2_emissions_by_vessel_type |>
  dplyr::filter(year == 2024) |>
  dplyr::mutate(
    fishing = ifelse(
      vessel_class %in%
        c(
          "trawlers",
          "squid_jigger",
          "tuna_purse_seines",
          "set_longlines",
          "pole_and_line",
          "set_gillnets",
          "pots_and_traps",
          "dredge_fishing",
          "other_seines",
          "other_purse_seines",
          "trollers"
        ),
      "Fishing",
      "Non-fishing"
    ) |>
      forcats::fct_rev()
  ) |>
  dplyr::mutate(
    vessel_class = stringr::str_replace_all(
      vessel_class,
      "\\.",
      ": "
    ) |>
      stringr::str_replace_all("_", " ") |>
      stringr::str_to_sentence() |>
      stringr::str_replace_all(
        "chemical oil",
        "chemical or oil"
      )
  ) |>
  ggplot(aes(
    x = emissions_co2_mt,
    y = reorder(vessel_class, emissions_co2_mt)
  )) +
  geom_col(position = position_dodge2(preserve = "single")) +
  ggforce::facet_col(~fishing, space = "free", scales = "free") +
  theme_plot() +
  labs(
    x = "2024 CO2 emissions (MT)",
    y = ""
  ) +
  scale_x_continuous(
    labels = scales::unit_format(unit = "M", scale = 1e-6),
    limits = c(0, NA)
  )

cowplot::plot_grid(
  cowplot::plot_grid(
    time_series_fig_co2,
    emissions_by_vessel_class_barplot,
    ncol = 1,
    rel_heights = c(0.4, 0.6),
    labels = c("A", "B")
  ),
  port_trip_emissions_fig,
  labels = c("", "C"),
  rel_widths = c(0.52, 0.48)
)
```

```{r}
#| label: tbl-top-port-stay-emissions-by-country
#| tbl-cap: Top 10 countries in terms of 2024 port stay emissions.
all_port_and_trip_emissions |>
  dplyr::mutate(
    country = countrycode::countrycode(
      country_iso3,
      "iso3c",
      "country.name"
    )
  ) |>
  dplyr::group_by(type) |>
  dplyr::slice_max(n = 3, order_by = emissions_co2_mt) |>
  knitr::kable()
```

```{r}
#| label: fig-pollutant-time-series
#| fig-cap: Time series of monthly global emissions for each pollutant, aggregated across AIS-broadcasting vessels and non-broadcasting vessels.
#| fig-width: 8
#| fig-height: 8

pollutant_time_series_data_all <- total_monthly_emissions_by_pollutant |>
  dplyr::mutate(
    pollutant = dplyr::case_when(
      pollutant == "CO2" ~ "CO[2]",
      pollutant == "N2O" ~ "N[2]*O",
      pollutant == "CH4" ~ "CH[4]",
      pollutant == "PM2.5" ~ "PM[2.5]",
      pollutant == "PM10" ~ "PM[10]",
      pollutant == "SOX" ~ "SO[X]",
      pollutant == "NOX" ~ "NO[X]",
      TRUE ~ pollutant
    )
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels)
  )

make_pollutant_time_series_figure(
  pollutant_time_series_data_all,
  y_axis_label = "Monthly emissions (MT)"
) +
  facet_wrap(
    pollutant ~ .,
    ncol = 3,
    scales = "free_y",
    labeller = label_parsed
  )
```

### Registered data validation

```{r}
#| label: fig-registered-data-performance
#| fig-cap: Comparison of registered and simulated CO$_{2}$ emissions from main engine fuel consumption over 24 hours. Estimates are shown using main engine power and design speed from IMO (left), our machine learning random forest models (center), and the same registered dataset (right). The solid red line shows a linear regression fit, while the dashed line represents the 1:1 line. Performance values for traditional R$^{2}$ are shown for each set of vessel characteristics. Blue shading indicates point density. All data are rule-of-three compliant, meaning each cell (20×20 mt) displayed contains at least three vessels.
#| fig-width: 8
#| fig-height: 4
model_levels <- c(
  "IMO report lookup table",
  "ML-derived engine power\nand design speed",
  "Registered data"
)

vessel_long <- registered_validation_data |>
  dplyr::select(
    co2_emissions_tonnes_registered,
    "IMO report lookup table" = co2_emissions_tonnes_estimate_imo,
    "ML-derived engine power\nand design speed" = co2_emissions_tonnes_estimate_rf,
    "Registered data" = co2_emissions_tonnes_estimate_registered
  ) |>
  tidyr::pivot_longer(
    cols = -co2_emissions_tonnes_registered,
    names_to = "Model",
    values_to = "Estimate"
  ) |>
  dplyr::mutate(Model = factor(Model, levels = model_levels)) # Order levels


r2_labels <- vessel_long |>
  dplyr::group_by(Model) |>
  yardstick::rsq_trad(
    truth = co2_emissions_tonnes_registered,
    estimate = Estimate
  ) |>
  dplyr::mutate(label = paste0("Traditional R² = ", round(.estimate, 2)))

ggplot(
  vessel_long,
  aes(
    x = Estimate,
    y = co2_emissions_tonnes_registered
  )
) +
  geom_bin2d(
    bins = 30,
    aes(
      fill = after_stat(count),
      alpha = after_stat(ifelse(count < 3, 0, 1)) # Full transparency for counts < 3
    )
  ) +
  scale_fill_gradientn(
    colours = colorRampPalette(RColorBrewer::brewer.pal(9, "Blues"))(256),
    trans = "log10",
    name = "Count",
    guide = guide_colorbar(
      ticks.linewidth = 0.3,
      frame.colour = "black"
    )
  ) +
  geom_smooth(
    method = "lm",
    linewidth = 0.5,
    color = "red",
    linetype = "solid",
    se = FALSE
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_text(
    data = r2_labels |>
      dplyr::filter(!Model %in% c("RF design speed", "RF engine power")),
    aes(label = label),
    x = 450,
    y = 50,
    inherit.aes = FALSE,
    size = 3,
    fontface = "bold"
  ) +
  coord_fixed(ratio = 1, clip = "on") +
  facet_wrap(Model ~ ., nrow = 1, scales = "fixed") +
  labs(
    x = "Simulated CO₂ Emissions (mt)",
    y = "Registered CO₂ Emissions (mt)"
  ) +
  scale_x_continuous(limits = c(0, 600)) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    axis.text.y = element_text(size = 7),
    axis.title = element_text(size = 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.ticks.length = unit(0.25, "cm"),
    panel.spacing = unit(1.5, "lines"),
    strip.placement = "outside",
    legend.position = "none", # Remove legend
  )
```

### MRV validation

```{r}
#| label: fig-mrv-performance
#| fig-cap: Comparison of published MRV emissions and simulated CO$_{2}$ emissions. The sample includes 2,276 observations from yearly aggregated vessel trips where hours at sea differ by no more than 15\% between our data and the MRV dataset. The solid red line shows the linear regression fit, and the dashed line represents the 1:1 line.}.
#| fig-width: 8
#| fig-height: 5

## Emissions validation ----
validation_performance_metrics <- yardstick::metric_set(
  yardstick::rsq,
  yardstick::rsq_trad
)

# EU dataset
eu_validation_data_updated <- mrv_data_validation |>
  dplyr::filter(total_time_spent_at_sea_hours != 0) |>
  dplyr::mutate(
    co2_trip_emissions_eu = co2_emissions_from_all_voyages_between_ports_under_a_ms_jurisdiction_m_tonnes +
      co2_emissions_from_all_voyages_which_departed_from_ports_under_a_ms_jurisdiction_m_tonnes +
      co2_emissions_from_all_voyages_to_ports_under_a_ms_jurisdiction_m_tonnes
  ) |>
  dplyr::rename(
    total_time_spent_at_sea_hours_eu = total_time_spent_at_sea_hours
  )
# Emission results by trip
eu_validation_trip_updated <- trip_emissions_for_mrv_validation |>
  dplyr::filter(total_time_spent_at_sea_hours != 0) |>
  dplyr::rename(
    co2_trip_emissions_gfw = total_emissions_co2_mt,
    total_time_spent_at_sea_hours_gfw = total_time_spent_at_sea_hours
  )

find_duplicates <- function(data, key_columns) {
  duplicate_indices <- duplicated(data[key_columns]) |
    duplicated(data[key_columns], fromLast = TRUE)
  duplicates <- data[duplicate_indices, ]
  return(duplicates)
}

aggregate_duplicates <- function(data, group_by_cols, sum_cols, concat_cols) {
  aggregated_data <- data |>
    dplyr::group_by(dplyr::across(dplyr::all_of(group_by_cols))) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(sum_cols), sum, .names = "{.col}"),
      dplyr::across(
        dplyr::all_of(concat_cols),
        ~ paste(unique(.x), collapse = "-"),
        .names = "{.col}"
      ),
      .groups = 'drop'
    )

  return(aggregated_data)
}

group_by_columns <- c("imo_number", "year")
sum_columns <- c(
  "total_time_spent_at_sea_hours_gfw",
  "total_distance_nm",
  "co2_trip_emissions_gfw"
)

concat_columns <- c("ssvid")

eu_validation_trip_updated <- aggregate_duplicates(
  eu_validation_trip_updated,
  group_by_columns,
  sum_columns,
  concat_columns
)

# Data selection and filtering
merged_df <- merge(
  eu_validation_data_updated,
  eu_validation_trip_updated,
  by.x = c("imo_number", "reporting_period"),
  by.y = c("imo_number", "year")
)

merged_df$diff <- abs(
  merged_df$total_time_spent_at_sea_hours_eu -
    merged_df$total_time_spent_at_sea_hours_gfw
)
merged_df$max <- pmax(
  merged_df$total_time_spent_at_sea_hours_eu,
  merged_df$total_time_spent_at_sea_hours_gfw
)

# Function to filter data and calculate performance metrics based on a threshold
calculate_metrics <- function(threshold, merged_df) {
  filtered_hours_df <- subset(merged_df, diff / max < threshold)
  filtered_df <- filtered_hours_df |>
    dplyr::select(co2_trip_emissions_eu, co2_trip_emissions_gfw)

  # Calculate performance metrics
  comparison_results <- filtered_df |>
    tidyr::pivot_longer(
      -co2_trip_emissions_eu,
      names_to = "model",
      values_to = "estimate"
    ) |>
    dplyr::group_by(model) |>
    validation_performance_metrics(
      truth = co2_trip_emissions_eu,
      estimate = estimate
    ) |>
    dplyr::select(-.estimator) |>
    tidyr::pivot_wider(names_from = .metric, values_from = .estimate)

  # Return results with added threshold information
  comparison_results <- comparison_results |>
    dplyr::mutate(threshold = threshold, n_observations = nrow(filtered_df))

  return(comparison_results)
}

# Range of thresholds to test
thresholds <- seq(0.05, 0.4, 0.05)
# Apply calculate_metrics function over range of thresholds
performance_results <- purrr::map_df(
  thresholds,
  ~ calculate_metrics(.x, merged_df)
)

best_performance <- performance_results |>
  dplyr::filter(rsq_trad == max(rsq_trad))

# Select data with threshold of 0.15 the one showing best performance
filtered_hours_df <- subset(merged_df, diff / max < best_performance$threshold)
filtered_df <- filtered_hours_df |>
  dplyr::select(co2_trip_emissions_eu, co2_trip_emissions_gfw)

ggplot(
  filtered_df,
  aes(
    x = co2_trip_emissions_gfw,
    y = co2_trip_emissions_eu
  )
) +
  geom_point(
    size = 1.5,
    alpha = 0.5,
    color = "#104E8B"
  ) +
  geom_smooth(
    method = "lm",
    linewidth = 0.5,
    color = "red",
    linetype = "solid",
    se = FALSE
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  annotate(
    geom = "text",
    label = paste0(
      "Traditional R² = ",
      format(best_performance$rsq_trad, digits = 2)
    ),
    x = 130000,
    y = 10000,
    size = 3,
    fontface = "bold"
  ) +
  coord_fixed(ratio = 1, clip = "on") +
  scale_x_continuous(limits = c(0, max(filtered_df))) +
  scale_y_continuous(limits = c(0, max(filtered_df))) +
  labs(
    x = "GFW CO2 emissions (mt)",
    y = "MRV CO2 emissions (mt)"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    axis.text.y = element_text(size = 7),
    axis.title = element_text(size = 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.ticks.length = unit(0.25, "cm"),
    panel.spacing = unit(1.5, "lines"),
    strip.placement = "outside"
  )
```

# Supplementary tables

## Dark fleet model performance metrics 

```{r}
#| results = 'asis'
all_performance_metrics_table <- all_performance_metrics |>
  dplyr::mutate(
    model = stringr::str_replace_all(model, "_", "; ") |>
      stringr::str_to_sentence(),
    test = stringr::str_replace_all(test, "_", " ") |>
      stringr::str_to_sentence()
  ) |>
  dplyr::select(
    Model = model,
    Test = test,
    Metric = metric,
    Estimate = estimate
  ) |>
  dplyr::filter(
    !(Metric %in%
      c(
        "rsq_trad_backtransformed_naive",
        "rsq_trad_backtransformed_regression",
        "rsq_trad_backtransformed_rmse",
        "rsq_trad_untransformed_log_scale"
      ))
  ) |>
  dplyr::mutate(
    Metric = dplyr::case_when(
      Metric == "accuracy" ~ "Accuracy",
      Metric == "f_meas" ~ "F1 score",
      Metric == "pr_auc" ~ "Precision-recall AUC",
      Metric == "roc_auc" ~ "ROC AUC",
      Metric == "precision" ~ "Precision",
      Metric == "sensitivity" ~ "Sensitivity",
      Metric == "specificity" ~ "Specificity",
      Metric ==
        "rsq_trad_backtransformed_residuals" ~ "Traditional R² (backtransformed)",
      Metric == "rsq_trad" ~ "Traditional R²",
      Metric == "mae" ~ "MAE",
      TRUE ~ Metric
    )
  ) |>
  dplyr::mutate(Model = forcats::fct_relevel(Model, "Emissions; regression")) |>
  dplyr::arrange(Model) |>
  tidyr::pivot_wider(values_from = Estimate, names_from = Test) |>
  knitr::kable(
    format = "latex",
    label = "all_performance_metrics_table",
    caption = "Summary of dark fleet model performance metrics for the: 1) emissions regression model; 2) detections classification mode; and 3) detections regression model. Performance estimates for each metric are shown for both the inside S1 footprint and outside S1 footprint tests.",
    digits = 3
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

all_performance_metrics_table

writeLines(
  all_performance_metrics_table,
  here::here("tables/all_performance_metrics_table.tex")
)
```

## Summary of 2017 and 2024 emissions, by pollutant and data source
  
```{r}
total_percent_change_by_fleet_tbl <- total_percent_change_by_fleet |>
  dplyr::mutate(across(
    c(`2017`, `2024`, percent_change),
    ~ signif(., 3) |>
      prettyNum(big.mark = ",")
  )) |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::arrange(pollutant) |>
  dplyr::rename(
    `Data source` = fleet,
    Pollutant = pollutant,
    `2017 emissions (MT)` = `2017`,
    `2024 emissions (MT)` = `2024`,
    `Percent 2024 total` = percent_total_2024,
    `Percent change from 2017 to 2024` = percent_change
  ) |>
  knitr::kable(
    format = "latex",
    label = "total_percent_change_by_fleet",
    caption = "Summary of 2017 and 2024 emissions, by pollutant and data source."
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

writeLines(
  total_percent_change_by_fleet_tbl,
  here::here("tables/total_percent_change_by_fleet.tex")
)
```

## Underestimation of 2024 emissions using only AIS data, and overestimation of 2017 to 2024 emissions trend using only AIS data, by pollutant.

```{r}
pollutant_ais_underestimation_overestimation_summary_tbl <- pollutant_ais_underestimation_overestimation_summary |>
  dplyr::mutate(
    pollutant = stringr::str_to_upper(pollutant) |>
      stringr::str_replace_all("_", "."),
    pollutant = ifelse(pollutant == "VOCS", "VOCs", pollutant)
  ) |>
  dplyr::mutate(
    pollutant = forcats::fct_relevel(pollutant, pollutant_factor_levels_tables)
  ) |>
  dplyr::arrange(pollutant) |>
  dplyr::rename(
    Pollutant = pollutant,
    "AIS-only underestimation of 2024 emissions" = ais_underestimation_2024,
    "AIS-only overestimation of 2017-2024 emissions trend" = trend_overestimation_2017_2024
  ) |>
  knitr::kable(
    format = "latex",
    label = "pollutant_ais_underestimation_overestimation_summary",
    caption = "Underestimation of 2024 emissions using only AIS data, and overestimation of 2017 to 2024 emissions trend using only AIS data, by pollutant."
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

writeLines(
  pollutant_ais_underestimation_overestimation_summary_tbl,
  here::here(
    "tables/pollutant_ais_underestimation_overestimation_summary.tex"
  )
)
```

### Feature importance

```{r}
#| label: fig-feature-importance
#| fig-cap: Feature importance plots for the three final model fit random forests
#| fig-height: 9

all_varimp_data |>
  dplyr::mutate(
    model = stringr::str_replace_all(model, "_", "; ") |>
      stringr::str_to_sentence() |>
      forcats::fct_relevel(
        "Emissions; regression",
        "Detections; classification",
        "Detections; regression"
      ),
  ) |>
    ggplot(aes(x = importance, y = tidytext::reorder_within(model_feature, importance, model))) +
    geom_bar(stat = "identity") +
    labs(
      x = "Feature importance score (Gini impurtity corrected)",
      y = "",
    ) +
    theme_plot() +
  tidytext::scale_y_reordered() +
  facet_wrap(model~.,ncol=1,scales="free")
```

### Classification model performance

```{r}
#| label: fig-pr-curves
#| fig-cap: Precision recall curves for the detection classification model, shown for the inside S1 footprint and outside S1 footprint performance tests

performance_detections_cls_roc_pr_curves |>
  tidyr::pivot_wider(names_from = "metric", values_from = "estimate") |>
  dplyr::mutate(test = stringr::str_replace_all(test, "_", " ")
                |> stringr::str_to_title()) |>
  ggplot(aes(x = recall, y = precision)) +
  geom_line() +
  facet_wrap(test~.,) +
    labs(
      x = "Recall",
      y = "Precision",
    ) +
    theme_plot()
```

```{r}
#| label: fig-roc-curves
#| fig-cap: ROC curves for the detection classification model, shown for the inside S1 footprint and outside S1 footprint performance tests

performance_detections_cls_roc_curve |>
  tidyr::pivot_wider(names_from = "metric", values_from = "estimate")|>
  dplyr::mutate(test = stringr::str_replace_all(test, "_", " ")
                |> stringr::str_to_title()) |>
  ggplot(aes(x = 1-specificity, y = sensitivity)) +
  geom_line() +
  facet_wrap(test~.,) +
    labs(
      x = "False positive rate (1 - specificity)",
      y = "True positive rate (sensitivity)",
    ) +
    theme_plot()
```


```{r}
#| label: fig-conf-mat
#| fig-cap: Confusion matricies for the detection classification model, shown for the inside S1 footprint and outside S1 footprint performance tests

performance_detections_cls_conf_mat |>
  dplyr::mutate(test = stringr::str_replace_all(test, "_", " ")
                |> stringr::str_to_title()) |>
  dplyr::mutate(prediction = forcats::fct_rev(prediction)) |>
  ggplot(aes(x = truth, y = prediction)) +
  geom_tile(aes(fill = count), color = "black") +
  geom_label(aes(label = prettyNum(count,big.mark=",")), vjust = 1) +
  scale_fill_gradient(low = "gray95", high = "gray10",guide="none") +
  facet_wrap(test~.) +
  coord_equal() +
  labs(x = "Truth",
       y = "Prediction") +
  theme_minimal()
```


### Dynamic AIS - annual emissions and vessels by AIS receiver type

```{r}
#| label: fig-annual-emissions-by-ais-receiver-type
#| fig-cap: Annual total CO2 emissions (metric tonnes) by the AIS-broadcasting fleet, disaggregated by AIS receiver type (satellite, terrestrial, and dynamic) and total aggregated across receiver types. A dashed vertical line is shown for 2022, the first year in which dynamic AIS was widely adopted.
#| #| fig-width: 7.5
#| fig-height: 6
annual_global_emissions_by_receiver_type <- annual_global_emissions_by_receiver_type |>
  # Don't include unknown receiver type
  dplyr::filter(receiver_type != "unknown") |>
  # Fill in missing values with 0s
  tidyr::complete(
    receiver_type,
    tidyr::nesting(year),
    fill = list(emissions_co2_mt = 0, n_distinct_ssvid = 0)
  ) |>
  dplyr::mutate(receiver_type = stringr::str_to_sentence(receiver_type)) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025)

annual_global_emissions_by_receiver_type |>
  dplyr::rename(
    `CO[2]~emissions~(MT)` = emissions_co2_mt,
    `Number~unique~vessels` = n_distinct_ssvid
  ) |>
  tidyr::pivot_longer(cols = -c(year, receiver_type)) |>
  ggplot(aes(x = year, y = value, color = receiver_type)) +
  facet_wrap(name ~ ., scales = "free_y", ncol = 1, labeller = label_parsed) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  scale_x_continuous(
    breaks = unique(annual_global_emissions_by_receiver_type$year)
  ) +
  scale_color_brewer(
    type = "qual",
    palette = "Set2",
    guide = guide_legend(reverse = TRUE)
  ) +
  theme_plot() +
  labs(x = "", y = "", color = "AIS receiver type") +
  theme(panel.grid.minor.x = element_blank()) +
  geom_vline(xintercept = 2022, linetype = 2)
```

### Dynamic AIS - annual emissions by AIS receiver type for top flags

```{r}
#| label: fig-annual-emissions-by-ais-receiver-type-top-flags
#| fig-cap: Annual total CO2 emissions (metric tonnes) by the AIS-broadcasting fleet, disaggregated by AIS receiver type (satellite, terrestrial, and dynamic) and total aggregated across receiver types, for the top 10 flags with the most emissions in 2024. A dashed vertical line is shown for 2022, the first year in which dynamic AIS was widely adopted.
#| fig-width: 7.5
#| fig-height: 6
annual_global_emissions_by_receiver_type_and_flag <- annual_global_emissions_by_receiver_type_and_flag |>
  # Add column name for country name based on flag

  dplyr::mutate(
    flag = countrycode::countrycode(
      flag,
      origin = "iso3c",
      destination = "country.name"
    )
  ) |>
  # Don't include unknown receiver type
  dplyr::filter(receiver_type != "unknown") |>
  dplyr::mutate(
    receiver_type = stringr::str_to_sentence(receiver_type) |>
      forcats::fct_relevel("Satellite", "Terrestrial", "Dynamic")
  ) |>
  # Fill in missing values with 0s
  tidyr::complete(
    receiver_type,
    tidyr::nesting(year, flag),
    fill = list(emissions_co2_mt = 0, n_distinct_ssvid = 0)
  ) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025)

top_n_flags_2024 <- annual_global_emissions_by_receiver_type_and_flag |>
  dplyr::filter(year == 2024, receiver_type == "Total") |>
  dplyr::slice_max(n = 10, order_by = emissions_co2_mt) |>
  dplyr::select(flag)

annual_global_emissions_by_receiver_type_and_flag |>
  dplyr::inner_join(top_n_flags_2024, by = "flag") |>
  dplyr::mutate(flag = forcats::fct_relevel(flag, top_n_flags_2024$flag)) |>
  ggplot(aes(x = year, y = emissions_co2_mt, color = flag)) +
  facet_grid(receiver_type ~ ., scales = "free_y") +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  scale_x_continuous(
    breaks = unique(annual_global_emissions_by_receiver_type_and_flag$year)
  ) +
  theme_plot() +
  labs(x = "", y = expression(CO[2] ~ emissions ~ (MT)), color = "Flag") +
  theme(panel.grid.minor.x = element_blank()) +
  geom_vline(xintercept = 2022, linetype = 2) +
  paletteer::scale_color_paletteer_d(palette = "rcartocolor::Safe")

```

### Dynamic AIS - map showing spatial change in emissions by AIS receiver type

```{r}
#| label: fig-co2-emissions-change-map-by-receiver-type
#| fig-cap: Spatial change in CO2 emissions by the AIS-broadcasting fleet between 2017 and 2024 (metric tonnes) by receiver type (satellite, terrestrial, and dynamic) and total across receiver types, at a 1x1 degree spatial resolution.
#| fig-width: 7.5
#| fig-height: 8
# Annual spatial CO2 emissions vt receiver type  (satellite; dynamic; or terrestrial)
annual_spatial_emissions_by_receiver_type <- annual_spatial_emissions_by_receiver_type |>
  dplyr::mutate(
    receiver_type = stringr::str_to_sentence(receiver_type) |>
      forcats::fct_relevel("Satellite", "Terrestrial", "Dynamic")
  ) |>
  # Don't use 2025 data, since the year is incomplete
  dplyr::filter(year < 2025) |>
  dplyr::mutate(
    emissions_mt_per_km2 = emissions_co2_mt / lat_to_area(lat_bin)
  ) |>
  dplyr::select(-emissions_co2_mt)

annual_spatial_emissions_by_receiver_type |>
  dplyr::filter(year %in% c(analysis_start_year, analysis_end_year)) |>
  tidyr::pivot_wider(names_from = year, values_from = emissions_mt_per_km2) |>
  dplyr::mutate(across(c(`2017`, `2024`), ~ tidyr::replace_na(., 0))) |>
  dplyr::mutate(emissions_mt_per_km2 = `2024` - `2017`) |>
  ggplot() +
  geom_sf(data = world_bbox_sf, fill = "grey90") +
  ggspatial::geom_spatial_tile(aes(
    x = lon_bin + 0.5,
    y = lat_bin + 0.5,
    fill = emissions_mt_per_km2,
    color = emissions_mt_per_km2
  )) +
  geom_sf(data = land, color = "#374a6d", fill = "#374a6d") +
  facet_wrap(receiver_type ~ ., ncol = 1) +
  scico::scale_fill_scico(
    expression(atop(
      Change ~ of ~ CO[2] ~ emissions,
      from ~ 2017 ~ to ~ 2024 ~ (MT / km^2)
    )),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  scico::scale_color_scico(
    expression(atop(
      Change ~ of ~ CO[2] ~ emissions,
      from ~ 2017 ~ to ~ 2024 ~ (MT / km^2)
    )),
    palette = "vik",
    midpoint = 0,
    trans = ggallin::pseudolog10_trans,
    na.value = "grey90",
    breaks = c(-1e3, -1e2, -1e1, 0, 1e1, 1e2, 1e3),
    labels = scales::comma
  ) +
  labs(x = "", y = "") +
  theme_void() +
  theme(legend.position = "right", legend.direction = "vertical") +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black",
      barwidth = unit(1, "cm"),
      barheight = unit(5, "cm")
    )
  ) +
  scale_x_continuous(expand = c(0, 1)) +
  scale_y_continuous(expand = c(0, 1)) +
  coord_sf(crs = map_projection)
```

## MRV performance validation table

```{r}
performance_results |>
  dplyr::mutate(
    rsq = signif(rsq, 2),
    rsq_trad = signif(rsq_trad, 2),
    n_observations = prettyNum(n_observations, big.mark = ",")
  ) |>
  dplyr::select(
    Tolerance = threshold,
    `Number of vessels` = n_observations,
    "R$^{2}$" = rsq,
    "Traditional R$^{2}$" = rsq_trad
  ) |>
  knitr::kable(
    format = "latex",
    label = "tab-performance-tolerance-table",
    caption = "Performance summary across different thresholds for the hours at sea tolerance between datasets. Results are shown using the R$^{2}$ (measuring linear association between observed and predicted values) and traditional R$^{2}$ (proportion of variance explained compared to the mean-only baseline).",
    escape = FALSE
  ) |>
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position") |>
  writeLines(
    here::here("tables/mrv_performance_results.tex")
  )
```